### 2018-11-28 siam_main.tex
Initial commit

% SIAM Article Template
\documentclass[review,onefignum,onetabnum]{siamonline171218}

% Information that is shared between the article and the supplement
% (title and author information, macros, packages, etc.) goes into
% ex_shared.tex. If there is no supplement, this file can be included
% directly.

\input{siam_shared}

% Optional PDF information
%\ifpdf
%\hypersetup{
%  pdftitle={An Example Article},
%  pdfauthor={D. Doe, P. T. Frank, and J. E. Smith}
%}
%\fi

% The next statement enables references to information in the
% supplement. See the xr-hyperref package for details.

%\externaldocument{ex_supplement}

% FundRef data to be entered by SIAM
%<funding-group>
%<award-group>
%<funding-source>
%<named-content content-type="funder-name"> 
%</named-content> 
%<named-content content-type="funder-identifier"> 
%</named-content>
%</funding-source>
%<award-id> </award-id>
%</award-group>
%</funding-group>

\begin{document}

\maketitle

% REQUIRED
\begin{abstract}
  This is an example SIAM \LaTeX\ article. This can be used as a
  template for new articles.  Abstracts must be able to stand alone
  and so cannot contain citations to the paper's references,
  equations, etc.  An abstract must consist of a single paragraph and
  be concise. Because of online formatting, abstracts must appear as
  plain as possible. Any equations should be inline.
\end{abstract}

% REQUIRED
\begin{keywords}
  example, \LaTeX
\end{keywords}

% REQUIRED
\begin{AMS}
  68Q25, 68R10, 68U05
\end{AMS}

\section{Introduction}

\section*{Acknowledgments}
We would like to acknowledge the assistance of volunteers in putting
together this example manuscript and supplement.

\bibliographystyle{siamplain}
\bibliography{references}

\end{document}

### 2018-11-28 siam_main.tex
Abstract and main body of paper included from aistats. Some formatting and spacing issues.

@@
 % REQUIRED
-\begin{abstract}
-  This is an example SIAM \LaTeX\ article. This can be used as a
-  template for new articles.  Abstracts must be able to stand alone
-  and so cannot contain citations to the paper's references,
-  equations, etc.  An abstract must consist of a single paragraph and
-  be concise. Because of online formatting, abstracts must appear as
-  plain as possible. Any equations should be inline.
-\end{abstract}
-
+\input{tex/abstract}
 % REQUIRED
@@
 
-% REQUIRED
-\begin{AMS}
-  68Q25, 68R10, 68U05
-\end{AMS}
-
-\section{Introduction}
-
-\section*{Acknowledgments}
-We would like to acknowledge the assistance of volunteers in putting
-together this example manuscript and supplement.
+\input{tex/paper}
 

### 2018-11-30 tex/abstract.tex
Adding abstract and main body of paper. Main body is similar to aistats except that the proposition proofs for Potts model have been added.

\begin{abstract}
    Submodular extensions of an energy function can be used to efficiently
    compute approximate marginals via variational inference. The accuracy of
    the marginals depends crucially on the quality of the submodular extension.
    To identify the best possible extension, we show an equivalence between the
    submodular extensions of the energy and the objective functions of linear
    programming (LP) relaxations for the corresponding MAP estimation problem.
    This allows us to (i) establish the worst-case optimality of the submodular
    extension for Potts model used in the literature; (ii) identify the
    worst-case optimal submodular extension for the more general class of
    metric labeling; and (iii) efficiently compute the marginals for the widely
    used dense CRF model with the help of a recently proposed Gaussian filtering method. Using synthetic and real data, we show that our approach provides comparable upper bounds on the log-partition function to those obtained using tree-reweighted message passing (TRW) in cases where the latter is computationally feasible. Importantly, unlike TRW, our approach provides the first practical algorithm to compute an upper bound on the dense CRF model.
\end{abstract}

### 2018-11-30 tex/paper.tex
Adding abstract and main body of paper. Main body is similar to aistats except that the proposition proofs for Potts model have been added.

\mysection{Introduction}
The desirable optimization properties of submodular set functions have been
widely exploited in the design of approximate MAP estimation algorithms for
discrete conditional random fields (CRFs) \citep{boykov2001fast,
kumar2011improved}. Submodularity has also been recently used to design an
elegant variational inference algorithm to compute the marginals of a discrete
CRF by minimising an upper-bound on the log-partition function. In the initial
work of \citep{djolonga2014map}, the energy of the CRF was restricted to be submodular. In a later work ~\citep{zhang2015higher}, the algorithm was extended to handle more general Potts energy functions. The key idea here is to define a large ground set such that its subsets represent valid labelings, sublabelings or even incorrect labelings (these may assign two separate labels to a random variable and hence be invalid). Given the large ground set, it is possible to define a submodular set function whose value is equal to the energy of the CRF for subsets that specify a valid labeling of the model. We refer to such a set function as a {\em submodular extension} of the energy.

For a given energy function, there exists a large number of possible submodular extensions. The accuracy of the variational inference algorithm depends crucially on the choice of the submodular extension. Yet, previous work has largely ignored the question of identifying the best extension. Indeed, the difficulty of identifying submodular extensions of general energy functions could be a major reason why the experiments of~\citep{zhang2015higher} were restricted to the special case of models specified by the Potts energy functions.

In this work, we establish a hitherto unknown connection between the submodular
extension of the Potts model proposed by Zhang et al.\citep{zhang2015higher}, and the
objective function of an accurate linear programming (LP) relaxation of the
corresponding MAP estimation problem~\citep{kleinberg2002approximation}. This
connection has three important practical consequences. First, it establishes
the accuracy of the submodular extension of the Potts model, via the
UGC-hardness worst-case optimality of the LP relaxation. Second, it provides an
accurate submodular extension of the hierarchical Potts model, via the LP
relaxation of the corresponding MAP estimation problem proposed
by Kleinberg and Tardos \citep{kleinberg2002approximation}. Since any metric can be accurately
approximated as a mixture of hierarchical Potts
models~\citep{bartal1996probabilistic, bartal1998approximating}, this result
also provides a computationally feasible algorithm for estimating the marginals
for metric labeling. Third, it establishes the equivalence between the
subgradient of the LP relaxation and the conditional gradient of the problem of
minimising the upper bound of the log-partition. This allows us to employ the
widely used dense CRF, since the subgradient of its LP relaxation can be
efficiently computed using a recently proposed modified Gaussian filtering
algorithm~\citep{ajanthan2017efficient}. As a consequence, we provide the first
efficient algorithm to compute an upper bound of the log-partition function of
dense CRFs. This provides complementary information to the popular mean-field
inference algorithm for dense CRFs, which computes a lower bound on the
log-partition ~\citep{koltun2011efficient}. We show that the quality of our
solution is comparable to tree reweighted message passing (TRW)
~\citep{wainwright2005new} for the case of sparse CRFs. Unlike our approach,
TRW is computationally infeasible for dense CRFs, thereby limiting its use in
practice. Using dense CRF models, we perform stereo matching on standard
data sets and obtain better results than \citep{koltun2011efficient}. The complete code is available at \url{https://github.com/pankajpansari/denseCRF}.

\mysection{Preliminaries}
\label{sec:prelim}

We now introduce the notation and definitions that we will make use of in the remainder of the paper.

\myparagraph{Submodular Functions} Given a ground set $U = \left\{1, \dots, N \right\}$, denote by $2^U$ its power set. A set function $F: 2^U \to \mathbb{R}$ is {\it submodular} if, for all subsets $A, B \subseteq U$, we have
\begin{equation}
    F(A \cup B) + F(A \cap B) \leq F(A) + F(B).
\end{equation}
The set function $F$ is {\it modular} if there exists ${\bf s} \in \mathbb{R}^N$ such that $F(A) = \sum_{k \in A} s_k \ \forall \ A \subseteq 2^U$. Henceforth, we will use the shorthand $s(A)$ to denote ${\sum_{k \in A}} s_k$.

\myparagraph{Extended Polymatroid} Associated with any submodular function $F$ is a special polytope known as the {\it extended polymatroid} defined as
%
\begin{align}
EP(F) &= \{{\bf s} \in \mathbb{R}^N | \enskip \forall A \subseteq U: s(A) \leq F(A)\},
\label{eq:epf}
\end{align}
where $\bf s$ denotes the modular function $s(.)$ considered as a vector.

\myparagraph{Lovasz Extension} For a given set function $F$ with $F(\emptyset) = 0$, the value of its Lovasz extension $f({\bf w}): \mathbb{R}^N \to \mathbb{R}$ is defined as follows: order the components of $\bf w$ in decreasing order such that $w_{j_1} \geq w_{j_2} \geq \dots \geq w_{j_N}$, where $(j_1, j_2, \dots, j_N)$ is the corresponding permutation of the indices. Then,
%
\begin{align}
    f({\bf w}) = \sum_{k = 1}^{N} w_{j_k} \left(F(j_1, \dots, j_k) - F(j_1,
    \dots, j_{k - 1})\right).
    \label{eq:lovasz_def}
\end{align}
%
The function $f$ is an extension because it equals $F$ on the vertices of the
unit cube. That is, for any $A \subseteq V$, $f({\bf 1}_A) = F(A)$ where ${\bf
1}_A$ is the $0\mhyphen{1}$ indicator vector corresponding to the elements of $A$.

{\prop By Edmond's greedy algorithm \citep{edmonds1970submodular}, if ${\bf w} \geq 0$ (non-negative elements), 
\begin{equation}
    f({\bf w}) = \max_{{\bf s} \in EP(F)} \langle {\bf w}, {\bf s} \rangle.
    \label{eq:lovasz_max}
\end{equation}
\label{prop:greedy}}

Property \ref{prop:greedy} implies that an LP over $EP(F)$ can be solved by computing the value of the Lovasz extension using equation (\ref{eq:lovasz_def}).

{\prop The Lovasz extension $f$ of a submodular function $F$ is a convex piecewise linear function.\label{prop:lovasz_convex}}

Property \ref{prop:lovasz_convex} holds since $f({\bf w})$ is the pointwise maximum of linear functions according to equation (\ref{eq:lovasz_max}).

\myparagraph{CRF and Energy Functions}
A CRF is defined as a graph on a set of random variables ${\mathcal X} = \{X_1,\dots, X_N\}$ related by a set of edges $\mathcal N$. We wish to assign every variable $X_a$ one of the labels from the set $\mathcal{L} = \{1, 2, \dots, L\}$. The quality of a labeling $\bf x$ is given by an energy function defined as 
%
\begin{equation}
    E({\bf x}) = \sum_{a \in {\mathcal X}} \phi_a(x_a) + \sum_{(a, b) \in {\mathcal N}} \phi_{ab}(x_a, x_b),
    \label{eq:energy_min}
  \end{equation}
%
where $\phi_a$ and $\phi_{ab}$ are the unary and pairwise potentials respectively. In computer vision, we often think of $\mathcal X$ as arranged on a grid. A \emph{sparse CRF} has $\mathcal N$ defined by 4-connected or 8-connected neighbourhood relationships. In a \emph{dense CRF}, on the other hand, every variable is connected to every other variable.

The energy function also defines a probability distribution $P({\bf x})$ as follows:
\begin{equation}
    P({\bf x}) =
    \begin{cases}
        \frac{1}{\mathcal{Z}} \exp(-E({\bf x})) & \quad \text{if} \enskip {\bf x} \in {\mathcal L}^N,\\
       0  & \quad \text{otherwise}. \\
    \end{cases}
\label{eq:crf_prob_distrib}
\end{equation}
The normalization factor $Z = \sum_{{\bf x} \in {\mathcal L}^N} \exp(-E({\bf x}))$ is known as the {\it partition function}.

\myparagraph{Inference} There are two inference problems in CRFs:

(i) Marginal inference: We want to compute the marginal probabilities $P(X_a = i)$ for every $a = 1, 2, \dots, N$ and $i = 1, 2, \dots, L$.

(ii) MAP inference: We want to find a labeling with the minimum energy, that is, $\min_{{\bf x} \in {\mathcal L}^N} E({\bf x})$. Equivalently, MAP inference finds the mode of $P({\bf x})$.

\mysection{Review: Variational Inference Using Submodular Extensions}
\label{sec:review}
We now summarise the marginal inference method of Zhang et al. \citep{zhang2015higher} which
uses submodular extensions.

\begin{figure}
\centering
\includegraphics[scale = 0.30]{./figures/1_of_L.png}
\mycaption{\footnotesize \em Illustration of $1$-of-$L$ encoding used in \citep{zhang2015higher} with 4 variables and 3 labels. The blue labeling, corresponding to $X_1 = 1, X_2 = 3, X_3 = 2, X_4 = 1$, is valid. The yellow labeling, corresponding to $X_2 = 2, X_3 = 1, 3, X_4 = 3$, is invalid since $X_3$ has been assigned multiple labels and $X_1$ has been assigned none.} 
\label{fig:encoding}
\end{figure}

\myparagraph{Submodular Extensions} A submodular extension is defined using
a ground set such that some of its subsets correspond to valid CRF labelings.
To such an extension, we need an encoding scheme which gives the sets corresponding to valid CRF labelings.

One example of an encoding scheme is the $1$-of-$L$ encoding, illustrated in figure \ref{fig:encoding}. Let each variable $X_a$ take one of $L$ possible labels. In this scheme, we represent the set of possible assignments for $X_a$ by the set $V_a = \{v_{a1}, v_{a2}, \dots, v_{aL}\}$. If $X_a$ is assigned label $i$, then we select the element $v_{ai}$. Extending to all variables, our ground set becomes $V = \cup_{a = 1}^N V_a$. A valid assignment $A \subseteq V$ assigns each variable exactly one label, that is, $|A \cap V_a| = 1$ for all $V_a$. We denote the set of valid assignments by $\mathcal M$ where $\mathcal{M} = \cap_{a = 1}^{N} \mathcal{M}_a$ and $\mathcal{M}_a = \{A: |A \cap V_a| = 1\}$.

Using our ground set $V$, we can define a submodular function $F$ which equals $E(\bf x)$ for all sets corresponding to valid labelings, that is, $F(A_{\bf x}) = E({\bf x}),\ {\bf x} \in {\mathcal L}^N$ where $A_{\bf x}$ is the set encoding of $\bf x$. We call such a function $F$ a {\it submodular extension} of $E({\bf x})$.

\myparagraph{Upper-Bound on Log-Partition} Using a submodular extension $F$ and given any ${\bf s} \in EP(F)$, we can obtain an upper-bound on the partition function as
\begin{equation}
    {\mathcal Z} = \sum_{A \in {\mathcal M}} \exp(-F(A)) \leq \sum_{A \in {\mathcal M}} \exp(-s(A)),
\end{equation}
where $\mathcal M$ is the set of valid labelings. The upper-bound is the partition function of the distribution $Q(A) \propto \exp(-s(A))$, which factorises fully because $s(.)$ is modular. Since ${\bf s} \in EP(F)$ is a free parameter, we can obtain good approximate marginals of the distribution $P(\cdot)$ by minimising the upper-bound. After taking logs, we can equivalently write our optimisation problem as
\begin{equation}
  \min_{{\bf s} \in EP(F)} g({\bf s}), \text{where } g({\bf s}) = \log \sum_{A \in {\mathcal M}} \exp(-s(A)).
\label{eq:min_upper_bound}
\end{equation}

\myparagraph{Conditional Gradient Algorithm} The conditional gradient algorithm
(algorithm \ref{algo:fw_inference}) \citep{frank1956algorithm} is a good
candidate for solving problem (\ref{eq:min_upper_bound}) due to two reasons.
First, problem \eqref{eq:min_upper_bound} is convex. Second, as solving an LP over EP(F) is computationally tractable (property \ref{prop:greedy}), the conditional gradient can be found efficiently. The algorithm starts with an initial solution $\bf s_0$ (line \ref{line:initial}). At each iteration, we compute the conditional gradient $\bf s^*$ (line \ref{line:cond_grad}), which minimises the linear approximation $g({\bf s}_k) + \nabla g({\bf s}_k)^T ({\bf s} - {\bf s}_k)$ of the objective function. Finally, $\bf s$ is updated by either (i) fixed step size schedule, as in line \ref{line:step_size} of algorithm \ref{algo:fw_inference}, or (ii) by doing line search ${\bf s}_{k+1} = \min_{0 \leq \gamma \leq 1} g(\gamma {\bf s}^* + (1 - \gamma) {\bf s}_{k})$.

\begin{algorithm}
\mycaption{Upper Bound Minimisation using Conditional Gradient Descent}
\label{algo:fw_inference}
\begin{algorithmic}[1]
    \STATE Initialize ${\bf s} = {\bf s}_0 \in EP(F)$ \label{line:initial}
\FOR{k = 1 to \texttt{MAX\_ITER}}
\STATE ${\bf s^*} = \argmin_{{\bf s} \in EP(F)} \langle \nabla g({\bf s}_k), {\bf s} \rangle$ \label{line:cond_grad}
\IF{$\langle {\bf s^* - s}_k, \nabla g({\bf s}_k) \rangle \leq \epsilon$}
\STATE {\bf break}
\ENDIF
\STATE ${\bf s}_{k + 1} = {\bf s}_k + \gamma({\bf s^*} - {\bf s}_k)$ \text{with} $\gamma = 2/(k + 2)$ \label{line:step_size}
\ENDFOR   
\RETURN $\bf s$
\end{algorithmic}
\end{algorithm}

\mysection{Worst-case Optimal Submodular Extensions via LP Relaxations}

\myparagraph{Worst-case Optimal Submodular Extensions} Different choices of extensions $F$ change the domain in problem \eqref{eq:min_upper_bound}, leading to different upper bounds on the log-partition function. How does one come up with an extension which yields the tightest bound?

In this paper, we focus on submodular extension families $\mathcal F(.)$ which
for each instance of the energy function $E(.)$ belonging to a given class $\mathcal E$ gives a corresponding
submodular extension ${\mathcal F}(E)$. We find the extension family ${\mathcal
F}_{opt}$ that is \emph{worst-case optimal}. This implies that there does not
exist another submodular extension family ${\mathcal F}$ that gives a tighter
upper bound for problem \eqref{eq:min_upper_bound} than ${\mathcal F}_{opt}$
for all instances of the energy function in $\mathcal E$. Formally, 
\begin{equation}
    \nexists {\mathcal F}: \min_{{\bf s} \in EP({\mathcal F}(E))} g({\bf s})
    \leq \min_{{\bf s} \in EP({\mathcal F}_{opt}(E))} g({\bf s}) \enspace
    \forall \; E(.) \in \mathcal E.
    \label{eq:best_submod}
\end{equation}
Note that our problem is different from taking a given energy model and
obtaining a submodular extension which is optimal for that model. Also, we seek
a closed-form analytical expression for $\mathcal F$. For the sake of clarity,
in the analysis that follows we use $F$ to represent ${\mathcal F}(E)$ where
the meaning is clear from context. The two classes of energy functions we
consider in this paper are Potts and hierarchical Potts families.
%
\myparagraph{Using LP Relaxations} If we introduce a temperature parameter in $P({\bf x})$ (equation \eqref{eq:crf_prob_distrib}) by using $E({\bf x})/T$ and decrease $T$, the resulting distribution starts to peak more sharply around its mode. As $T \to 0$, marginal estimation becomes the same as MAP inference since the resulting distribution $P^0({\bf x})$ has mass 1 at its mode ${\bf x}^*$ and is 0 everywhere else. Given the MAP solution ${\bf x}^*$, one can compute the marginals as $P(X_i = j) = [x_i^* = j]$, where [.] is the Iverson bracket.

Motivated by this connection, we ask if one can introduce a temperature parameter to our problem \eqref{eq:min_upper_bound} and transform it to an LP relaxation in the limit $T \to 0$? We can then hope to use the tightest LP relaxations of MAP problems known in literature to find worst-case optimal submodular extensions. We answer this question in affirmative. Specifically, in the following two sections we show how one can select the set encoding and submodular extension to convert problem \eqref{eq:min_upper_bound} to the tightest known LP relaxations for Potts and hierarchical Potts models. Importantly, we prove the worst-case optimality of the extensions thus obtained.

\begin{figure}
\centering
\includegraphics[scale = 0.30]{./figures/potts_chain_extension.png}
\mycaption{\footnotesize \em An illustration of the worst-case optimal
submodular extension for Potts model for a chain graph of 4 variables, each of
which can take 3 labels. The figure shows the way to compute the extension
values of the set $A = \{v_{1, 1}, v_{4, 1}, v_{3, 2}$\}.}
\label{fig:alternate}
\end{figure}
\begin{figure}
\centering
\resizebox{150pt}{120pt}{
\begin{tikzpicture}[-latex ,auto, node distance =1.5 cm and 1.5cm ,on grid
  , semithick , state/.style ={ circle , color = black , draw,black , text=blue , minimum width =1 cm}, cross/.style={color=red, very thick, dashed}, unary/.style={color=blue, very thick}]

  \node[state] (n11) at (1, 1) {\Large $1\mhyphen1$};
  \node[state] (n12) at (3, 1) {\Large $1\mhyphen2$};
  \node[state] (n13) at (5, 1) {\Large $1\mhyphen3$};
  \node[state] (n21) at (7, 1) {\Large $2\mhyphen1$};
  \node[state] (n22) at (9, 1) {\Large $2\mhyphen2$};
  \node[state] (n23) at (11, 1) {\Large $2\mhyphen3$};
  \node[state] (source) at (6, 3) {\Large s};
  \node[state] (sink) at (6, -3) {\Large t};

% \begin{scope}[ every node/.style={fill=white,circle}]
\path[unary] (n11) edge [bend right =25] node [pos=0.20, anchor=west] {\Large$\phi_1(1)$} (sink);
\path[unary] (n12) edge [bend right =25] node [pos=0.20, anchor=west]{\Large$\phi_1(2)$} (sink);
\path[unary] (n13) edge [bend right =25] node [pos=0.20, anchor=west]{\Large$\phi_1(3)$} (sink);
\path[unary] (n21) edge [bend left =25] node [pos=0.20, anchor=west]{\Large$\phi_2(1)$} (sink);
\path[unary] (n22) edge [bend left =25] node [pos=0.20, anchor=west]{\Large$\phi_2(2)$} (sink);
\path[unary] (n23) edge [bend left =25] node [pos=0.20, anchor=west]{\Large$\phi_2(3)$} (sink);
%\end{scope}

\path[cross] (n11) edge [bend left =25] node[above] {} (n21);
\path[cross] (n21) edge [bend left =25] node[below] {} (n11);

\path[cross] (n12) edge [bend left =25] node[above] {} (n22);
\path[cross] (n22) edge [bend left =25] node[below] {} (n12);

\path[cross] (n13) edge [bend left =25] node[above] {} (n23);
\path[cross] (n23) edge [bend left =25] node[below] {} (n13);
\end{tikzpicture}
}
\mycaption{\footnotesize \em An st-graph specifying the worst-case optimal submodular
extension for Potts model for 2 variables with 3 labels each and connected to
each other. There is a node for each variable and each label, that is, for all
elements of the ground set. The nodes have been labeled as `variable-label',
hence node 1-1 represents the element $v_{11}$ and so on. The solid blue arcs
model the unary potentials, and the dotted red arcs represent the pairwise
potentials. Each dotted red arc has weight $w_{12}/2$.}
\label{fig:optimal_extension_graph}
\end{figure}

\mysection{Potts Model}

The Potts model, also known as the uniform metric, specifies the pairwise potentials $\phi_{ab}(x_a, x_b)$ in equation (\ref{eq:energy_min}) as follows:
\begin{equation}
    \phi_{ab}(x_a, x_b) = w_{ab} \cdot [x_a \neq x_b],
\end{equation}
where $w_{ab}$ is the weight associated with edge $(a, b)$. 
%
\myparagraph{Tightest LP Relaxation} Before describing our set encoding and submodular extension, we briefly outline the LP relaxation of the corresponding MAP estimation problem. To this end, we define indicator variables $y_{ai}$ which equal 1 if $X_a = i$, and 0 otherwise. The following LP relaxation is the tightest known for Potts model in the worst-case, assuming the Unique Games Conjecture to be true \citep{manokaran2008sdp}
\begin{align}
    \text{(P-LP)} \quad \min_{\bf y} \enskip & E({\bf y}) =  \sum_{a} \sum_{i} \phi_a(i)y_{ai} + \nonumber\\
     &\sum_{(a, b) \in {\mathcal N}} \sum_{i} \frac{w_{ab}}{2} \cdot |y_{ai} - y_{bi}| \nonumber\\
    \text{s.t} \quad &{\bf y} \in \Delta.
\label{eq:p-lp}
\end{align}
The set $\Delta$ is the union of $N$ probability simplices:
\begin{equation}
    \Delta = \{{\bf y}_a \in \mathbb{R}^L | {\bf y}_a \geq 0  \textrm{ and } \langle {\bf 1}, {\bf y}_a \rangle = 1\},
    \label{eq:delta}
\end{equation}
where $\bf y$ is the vector of all variables and ${\bf y}_a$ is the component
of $\bf y$ corresponding to $X_a$.

\myparagraph{Set Encoding} We choose to use the $1$-of-$L$ encoding for Potts model as described in section \ref{sec:review}. With the encoding scheme for Potts model above, $g({\bf s})$ can be factorised and problem (\ref{eq:min_upper_bound}) can be rewritten as:
%
\begin{equation}
\min_{{\bf s} \in EP(F)} \quad \sum_{a = 1}^{N} \log \sum_{i = 1}^L \exp(-s_{ai}).
\label{eq:potts_upper_bound}
\end{equation}
(See Remark 1 in appendix)
\myparagraph{Marginal Estimation with Temperature} We now introduce
a temperature parameter $T > 0$ to problem \eqref{eq:potts_upper_bound} which
divides $E({\bf x})$, or equivalently divides $\bf s$ belonging to $EP(F)$. Also, since $T > 0$, we can multiply the objective by $T$ leaving the problem unchanged. Without changing the solution, we can transform problem \eqref{eq:potts_upper_bound} as follows
\begin{equation}
    \min_{{\bf s} \in EP(F)} \quad g_T ({\bf s}) = \sum_{a = 1}^{N} T \cdot \log \sum_{i = 1}^L \exp(-\frac{s_{ai}}{T}).
\label{eq:potts_temp}
\end{equation}
\myparagraph{Worst-case Optimal Submodular Extension} 
We now connect our marginal estimation problem (\ref{eq:min_upper_bound}) with
LP relaxations using the following proposition.
{\proposition Using the $1$-of-$L$ encoding scheme, in the limit $T \to 0$,
  problem (\ref{eq:potts_temp}) for Potts model becomes:
\begin{align}
    - \min_{{\bf y} \in \Delta} f({\bf y}) 
\end{align}
where $f(.)$ is the Lovasz extension of $F(.)$.
\label{proposition:potts_equiv}}

\begin{proof}
In the limit of $T \to 0$, we can rewrite the above problem as
\begin{equation}
    \min_{{\bf s} \in EP(F)} \quad \sum_{a = 1}^{N} \max_{i} (-s_{ai}).
\end{equation}
In vector form, the problem becomes
\begin{align}
    &\min_{{\bf s} \in EP(F)} \max_{{\bf y} \in \Delta} - \langle {\bf y}, {\bf s} \rangle \\
    & = - \max_{{\bf s} \in EP(F)} \min_{{\bf y} \in \Delta} \langle {\bf y}, {\bf s} \rangle
    \label{eq:maxmin}
\end{align}
$\Delta$ is the union of $N$ probability simplices:
\begin{equation}
    \Delta = \{{\bf y}_a \in \mathbb{R}^L | {\bf y}_a \succeq 0  \textrm{ and } \langle {\bf 1}, {\bf y}_a \rangle = 1\}
    \label{eq:delta2}
\end{equation}
where ${\bf y}_a$ is the component of $\bf y$ corresponding to the $a$-th variable. By the minimax theorem for LP, we can reorder the terms:
\begin{equation}
    - \min_{{\bf y} \in \Delta} \max_{{\bf s} \in EP(F)} \langle {\bf y}, {\bf s} \rangle 
    \label{eq:minmax}
\end{equation}
Recall that $\max_{{\bf s} \in EP(F)} \langle {\bf y}, {\bf s} \rangle$ is the value of the Lovasz extension of $F$ at $\bf y$, that is, $f({\bf y})$. Hence, as $T \to 0$, the marginal inference problem converts to minimising the Lovasz extension under the simplices constraint:
\begin{equation}
    - \min_{{\bf y} \in \Delta} f({\bf y}) 
    \label{eq:minLovasz}
\end{equation}
\end{proof}

The above problem is equivalent to an LP relaxation of the corresponding MAP estimtation problem (see Remark 2 in appendix). We note that $g_T ({\bf s})$ in problem \eqref{eq:potts_temp} becomes the objective function of an LP relaxation in the limit $T \to 0$. We seek to obtain the worst-case optimal submodular extension by making $g_T ({\bf s})$ same as the objective of (P-LP) as $T \to 0$. Since at $T = 1$, problems \eqref{eq:potts_temp} and \eqref{eq:potts_upper_bound} are equivalent, this gives us the worst-case optimal extension for our problem \eqref{eq:potts_upper_bound}.

The question now becomes how to recover the worst-case optimal submodular extension using $E({\bf y})$. The following propositions answers this question.
{\proposition The worst-case optimal  submodular extension for Potts model is given by $F_{Potts}(A) = \sum_{i = 1}^L F_i(A)$, where
\begin{align}
    F_i(A) &= \sum_a \phi_{a}(i) [|A \cap \{v_{ai}\}| = 1] + \nonumber \\
           &\sum_{(a, b) \in {\mathcal N}} \frac{w_{ab}}{2} \cdot [|A \cap \{v_{ai}, v_{bi}\}| = 1]
\end{align}
Also, $E({\bf y})$ in (P-LP) is the Lovasz extension of $F_{Potts}$.
\label{proposition:potts_worst-case optimal}}
\begin{proof}
Since $F_{Potts}$ is sum of Ising models $F_i$, we first focus on a particular label $i$ and then generalize. Consider a graph with only two variables $X_a$ and $X_b$ with an edge between them. The ground set in this case is $\{v_{ai}, v_{bi}\}$. Let the corresponding relaxed indicator variables be ${\bf y} = \{y_{aj}, y_{bj}\}$, such that $y_{ai}, y_{bi} \in [0, 1]$ and assume $y_{ai} > y_{bi}$. The Lovasz extension in this case is:

\begin{align}
f({\bf y}) &= y_{ai} \cdot [F_i(\{v_{ai}\}) - F_i(\{\})] + y_{bi} \cdot [F_i(\{v_{ai}, v_{bi}\}) - F_i(\{v_{ai}\})] \nonumber \\
&= y_{ai} \cdot [\left(\phi_{a}\left(j\right) + \frac{w_{ab}}{2}\right) - 0] + y_{bi} \cdot [\left(\phi_{a}\left(j\right) + \phi_{b}\left(j\right)\right) - \left(\phi_{a}\left(j\right) + \frac{w_{ab}}{2}\right)] \nonumber \\
&=  \phi_{a}\left(j\right) \cdot y_{ai} + \phi_{b}\left(j\right) \cdot y_{bi} + \frac{w_{ab}}{2} \cdot \left(y_{ai} - y_{bi}\right)
\end{align}

In general for both orderings of $y_{ab}$ and $y_{bi}$, we can write
%
\begin{equation}
f({\bf y}) = \phi_{a}(j) \cdot y_{ai} + \phi_{b}(j) \cdot y_{bi} + \frac{w_{ab}}{2} \cdot |y_{ai} - y_{bi}|
\label{eq:a-lovasz_ising}
\end{equation}
%
Extending the Lovasz extension of equation \eqref{eq:a-lovasz_ising} to all variables and all labels gives us $E({\bf y})$ in (P-LP).

\end{proof}
\vspace{-1mm}

Proposition \ref{proposition:potts_worst-case optimal} paves the way for us to
identify the worst-case optimal extension for hierarchical Potts model, which
we discuss in the following section.

\mysection{Hierarchical Potts Model}
\label{sec:rhst_worst-case optimal}

\begin{figure}
\centering
\includegraphics[scale = 0.30]{./figures/rhst_notation.png}
\mycaption{\footnotesize \em A hierarchical Potts model instance illustrating notations with 2 meta-labels (blue) and 4 labels (yellow). All labels are at the same level. $r = 2$, that is, edge-length decreases by 2 at each level. Also, distance between labels 1 and 3 is $d(1, 3) = 0.5 + 1 + 1 + 0.5 = 3$.} 
\label{fig:rhst}
\end{figure}
Potts model imposes the same penalty for unequal assignment of labels to neighbouring variables, regardless of the label dissimilarity. A more natural approach is to vary the penalty based on how different the labels are. A hierarchical Potts model permits this by specifying the distance between labels using a tree with the following properties:
\vspace{2mm}
\begin{enumerate}
  \itemsep+1em 
\item The vertices are of two types: (i) the leaf nodes representing labels, and (ii) the non-leaf nodes, except the root, representing meta-labels.
\item The lengths of all the edges from a parent to its children are the same.
\item The lengths of the edges along any path from the root to a leaf decreases by a factor of at least $r \geq 2$ at each step.
\item The metric distance between nodes of the tree is the sum of the edge lengths on the unique path between them. 
\end{enumerate}
\vspace{2mm}
%
A subtree $T$ of an hierarchical Potts model is a tree comprising all the descendants of some node $v$ (not root). Given a subtree $T$, $l_T$ denotes the length of the tree-edge leading upward from the root of $T$ and $L(T)$ denotes the set of leaves of $T$. We call the leaves of the tree as labels and all other nodes of the tree expect the root as \emph{meta-labels}. Figure \ref{fig:rhst} illustrates the notations in the context of a hierarchical Potts model.
%
\myparagraph{Tightest LP Relaxation} We use the same indicator variables $y_{ai}$ that were employed in the LP relaxation of Potts model. Let $y_a(T) = \sum_{i \in L(T)} y_{ai}$. The following LP relaxation is the tightest known for hierarchical Potts model in the worst-case, assuming the Unique Games Conjecture to be true \citep{manokaran2008sdp}
\begin{align}
\text{(T-LP)} \quad \min_{\bf y} \enskip & \widetilde{E}({\bf y}) =  \sum_{a} \sum_{i} \phi_a(i)y_{ai} + \nonumber\\
                                         &\sum_{(a, b) \in {\mathcal N}} w_{ab} \sum_T l_T \cdot |y_{a}(T) - y_{b}(T)| \nonumber\\
    \text{such that} \quad &{\bf y} \in \Delta.
\label{eq:t-lp}
\end{align}
The set $\Delta$ is the same domain as defined in equation (\ref{eq:delta}). We
rewrite this LP relaxation using indicator variables $z_{ai}$ for all labels
and meta-labels as
\begin{align}
    \text{(T-LP-FULL)} &\quad \min \widetilde{E}({\bf z}) \nonumber \\
    {} &\text{such that} \quad {\bf z} \in \Delta'
\label{eq:t-lp-full}
\end{align}
   where $\Delta'$ is the convex hull of the vectors satisfying
\begin{align}
    \enskip &\sum_{i \in \mathcal{L}} z_{ai} = 1, \enskip z_{ai} \in \{0, 1\} \enskip \forall a \in {\mathcal X}, i \in \mathcal{L} \\
    \text{and} \enskip &z_{ai} = \sum_{j \in L(T_i) } z_{aj}.\enskip \forall a \in {\mathcal X}, i \in \mathcal{R - L} \label{eq:consistency_constraint} 
\end{align}
The details of the new relaxation (T-LP-FULL) can be found in the appendix.
%
\myparagraph{Set Encoding} For any variable $X_a$, let the set of possible assignment of labels and meta-labels be the set $V_a = \{v_{a1}, \dots, v_{aM}\}$, where $M$ is the total number of nodes in the tree except the root. Our ground set is $V = \cup_{a = 1}^{N} V_a$ of size $N \cdot M$.

A consistent labeling of a variable assigns it one label, and all meta-labels on the path from root to the label. Let us represent the set of consistent assignments for $X_a$ by the set $P_a = \{p_{a1}, \dots, p_{aL}\}$, where $p_{ai}$ is the collection of elements from $V_a$ for label $i$ and all meta-labels on the path from root to label $i$. The set of valid labelings $A \subseteq V$ assigns each variable exactly one consistent label. This constraint can be formally written as $\mathcal{M} = \cap_{a = 1}^{N} \mathcal{M}_a$ where $\mathcal{M}_a$ has exactly one element from $P_a$. Let $s'_{ai}$ be the sum of the components of $\bf s$ corresponding to the elements
of $p_{ai}$, that is, 
\begin{equation}
s'_{ai} = \sum_{t \in p_{ai}} s_t.
\label{eq:tree_s}
\end{equation}
Using our encoding scheme, we rewrite problem (\ref{eq:min_upper_bound}) as:
\begin{equation}
\min_{{\bf s} \in EP(F)} \quad \sum_{a = 1}^{N} \log \sum_{i = 1}^L \exp(-s'_{ai}).
\label{eq:metric_upper_bound}
\end{equation}
%
\myparagraph{Marginal Estimation with Temperature} Similar to Potts model, we now introduce a temperature parameter $T > 0$ to problem \eqref{eq:metric_upper_bound}. The transformed problem becomes
\begin{equation}
    \min_{{\bf s} \in EP(F)} \quad g_T ({\bf s}) = \sum_{a = 1}^{N} T \cdot \log \sum_{i = 1}^L \exp(-\frac{s'_{ai}}{T}).
\label{eq:metric_temp}
\end{equation}
\myparagraph{Worst-case Optimal Submodular Extension} The following proposition connects the marginal estimation problem (\ref{eq:min_upper_bound}) with LP relaxations:

{\proposition In the limit $T \to 0$, problem (\ref{eq:metric_temp}) for hierarchical Potts energies becomes:
\begin{align}
    - \min_{{\bf z} \in \Delta'} f({\bf z}) 
\end{align}
\label{proposition:metric_equiv}}
(Proof in appendix).

The above problem is equivalent to an LP relaxation of the corresponding MAP
estimtation problem (see Remark 3 in appendix). Hence, $g_T ({\bf s})$ becomes the objective function of an LP relaxation in the limit $T \to 0$. We seek to make this objective same as $\widetilde{E}({\bf z})$ of (T-LP-FULL) in the limit $T \to 0$. The question now becomes how to recover the  worst-case optimal submodular extension from $\widetilde{E}({\bf z})$.
%
{\proposition The  worst-case optimal submodular extension for hierarchical Potts model is given by $F_{hier}(A) = \sum_{i = 1}^M F_i(A)$, where
\begin{align}
    F_i(A) &= \sum_a \phi'_{a}(i) [|A \cap \{v_{ai}\}| = 1] + \nonumber \\
           &\sum_{(a, b) \in {\mathcal N}} {w_{ab}} \cdot l_{T_i} \cdot [|A \cap \{v_{ai}, v_{bi}\}| = 1]
\end{align}
Also, $\widetilde{E}({\bf z})$ in (T-LP-FULL) is the Lovasz extension of $F_{hier}$.
\label{proposition:rhst_worst-case optimal}}
(Proof in appendix)

Since any finite metric space can be probabilistically approximated by mixture
of tree metric \citep{bartal1996probabilistic}, the worst-case optimal
submodular extension for metric energies can be obtained using $F_{hier}$. Note
that $F_{hier}$ reduces to $F_{Potts}$ for Potts model. One can see this by
considering the Potts model as a star-shaped tree with edge weights as 0.5.
\begin{figure}
\centering
\begin{tabular}{cc}
\includegraphics[scale = 0.16]{./figures/synthetic_tree.png} &
\includegraphics[scale = 0.16]{./figures/synthetic_tree_2_new.png} \\
  \scriptsize(a) Tree A & \scriptsize(b) Tree B \\
\end{tabular}
\mycaption{\footnotesize \em The hierarchical Potts models defining pairwise
distance among 20 labels used for upper-bound comparison with TRW. Blue nodes
are the meta-labels and yellow nodes are labels. All the edges at a particular
level have the same edge weights. The sequence of weights from root level to
leaf level is 1, 0.5 for tree A and 1, 1, 0.5 for tree B. The yellow node is
shown to clump together 4 and 5 leaf nodes for tree A and B respectively.} 
\label{fig:syn_rhst}
\end{figure}

\mysection{Fast Conditional Gradient Computation for Dense CRFs}
\label{sec:dense_crf}

\myparagraph{Dense CRF Energy Function} A dense CRF is specified by the following energy function 
\begin{equation}
    E({\bf x}) = \sum_{a \in {\mathcal X}} \phi_a(x_a) + \sum_{a \in {\mathcal X}} \sum_{b \in {\mathcal X}, b \neq a} \phi_{ab}(x_a, x_b).
\end{equation}
Note that every random variable is a neighbour of every other random variable in a dense CRF. Similar to previous work \citep{koltun2011efficient}, we consider the pairwise potentials to be to be Gaussian, that is, 
\begin{align}
    &\phi(i, j) = \mu(i, j) \sum_{m} w^{(m)} k({\bf f}_a^{(m)}, {\bf f}_b^{(m)}), \\
    &k({\bf f}_a^{(m)}, {\bf f}_b^{(m)}) = \exp\left( \frac{- ||{\bf f}_a - {\bf f}_b ||^2}{2} \right).
\end{align}
The term $\mu(i,j)$ is known as {\it label compatibility} function between
labels $i$ and $j$. Potts model and hierarchical Potts models are examples of
$\mu(i, j)$. The other term is a mixture of Gaussian kernels $k(.,.)$ and is
called the {\it pixel compatibility} function. The terms ${\bf f}_a^{(m)}$ are
features that describe the random variable $X_a$. In practice, similar to
\citep{koltun2011efficient}, we use $x, y$ coordinates and RGB values
associated to a pixel as its features.

Algorithm \ref{algo:fw_inference} assumes that the conditional gradient $\bf s^*$ in step {\ref{line:cond_grad}} can be computed efficiently. This is certainly not the case for dense CRFs, since computing $s^*$ involves $NL$ function evaluations of the submodular extension $F$, where $N$ is the number of variables, and $L$ is the number of labels. Each $F$ evaluation has complexity ${\mathcal O}(N)$ using the efficient Gaussian filtering algorithm of \citep{koltun2011efficient}. However, computation of $s^*$ would still be ${\mathcal O} (N^2)$ this way, which is clearly impractical for computer-vision applications where $N \sim 10^5 - 10^6$.

However, using the equivalence of relaxed LP objectives and the Lovasz
extension of submodular extensions in proposition
\ref{proposition:potts_equiv}, we are able to compute $s^*$ in ${\mathcal
O}(NL)$ time. Specifically, we use the algorithm of
Ajanthan et al. \citep{ajanthan2017efficient}, which provides an efficient filtering procedure to compute the subgradient of the LP relaxation objective $E({\bf y})$ of (P-LP).\\
%
\vspace{-0.5cm}
{\proposition Computing the subgradient of $E({\bf y})$ in (P-LP) is equivalent to computing the conditional gradient for the submodular function $F_{Potts}$. \label{proposition:subgrad}} 
\newline(Proof in appendix).
%\begin{proof}
%From lemma \ref{lemma:lovasz_lp}, 
%\begin{align}
%    E({\bf y}) &= f({\bf y})  \nonumber \\
%    {} &=  \max_{{\bf s} \in EP(F)} \langle {\bf y}, {\bf s} \rangle.\nonumber
%\end{align}
%
%The subdifferential of $E({\bf y})$ at $\bf y_0$ is the `active' linear function at $\bf y_0$. Hence,
%\begin{align}
%    \partial E({\bf y})|_{\bf y = y_0} &=  \argmax_{{\bf s} \in EP(F)} \langle {\bf y_0}, {\bf s} \rangle
%    \label{eq:subgrad}
%\end{align}
%Equation (\ref{eq:subgrad}) is equivalent to ${\bf s}^*$ computation in line \ref{line:cond_grad} of algorithm \ref{algo:fw_inference}, with ${\bf y_0} = - \nabla g({\bf s}_k)$.
%\end{proof}
%

\begin{figure}
\centering
\resizebox{150pt}{120pt}{
\begin{tikzpicture}[-latex ,auto ,node distance =2 cm and 2cm ,on grid
  , semithick , state/.style ={ circle , color = black , draw,black , text=blue , minimum width =1 cm}, cross/.style={color=red, very thick, dashed}, unary/.style={color=blue, very thick}]

  \node[state] (n11) at (1, 1) {\Large $1\mhyphen1$};
  \node[state] (n12) at (5, 1) {\Large $1\mhyphen2$};
  \node[state] (n13) at (3, -1) {\Large $1\mhyphen3$};
  \node[state] (n21) at (7, 1) {\Large $2\mhyphen1$};
  \node[state] (n22) at (11, 1) {\Large $2\mhyphen2$};
  \node[state] (n23) at (9, -1) {\Large $2\mhyphen3$};
  \node[state] (source) at (6, 3) {\Large s};
  \node[state] (sink) at (6, -3) {\Large t};

  \path[unary] (n11) edge [bend left =25] node[below] {\Large$\phi_1(1)$} (n12);
\path[unary] (n12) edge [bend left =25] node[left] {\Large$\phi_1(2)$} (n13);
\path[unary] (n13) edge [bend left =25] node[left] {\Large$\phi_1(3)$} (n11);
\path[unary] (n21) edge [bend left =25] node[below] {\Large$\phi_2(1)$} (n22);
\path[unary] (n22) edge [bend left =25] node[right] {\Large$\phi_2(2)$} (n23);
\path[unary] (n23) edge [bend left =25] node[right] {\Large$\phi_2(3)$} (n21);

\path[cross] (n11) edge [bend left =25] node[above] {} (n21);
\path[cross] (n21) edge [bend left =25] node[below] {} (n11);

\path[cross] (n12) edge [bend left =25] node[above] {} (n22);
\path[cross] (n22) edge [bend left =25] node[below] {} (n12);

\path[cross] (n13) edge [bend left =25] node[above] {} (n23);
\path[cross] (n23) edge [bend left =25] node[below] {} (n13);
\end{tikzpicture}
}
\mycaption{\footnotesize \em An st-graph specifying the alternate submodular
extension for Potts model for 2 variables with 3 labels each and connected to
each other. The convention used is same as in figure \ref{fig:optimal_extension_graph}. Each dotted red arc has weight $w_{12}/2$.  This alternate extension was also used to derive the extension for hierarchical Potts model.}
\label{fig:alternate_extension}
\end{figure}

A similar observation can be made in case of hierarchical Potts model. Hence we have the first practical algorithm to compute upper bound of log-partition function of a dense CRF for Potts and metric energies.

\mysection{Experiments}
\label{sec:exp}
Using synthetic data, we show that our upper-bound compares favorably with TRW
for both Potts and hierarchical Potts models. For comparison, we restrict
ourselves to sparse CRFs, as the code available for TRW does not scale well to
dense CRFs. We also perform stereo matching using dense CRF models and compare
our results with the mean-field-based approach of \citep{koltun2011efficient}.
All experiments were run on a x86-64, 3.8GHz machine with 16GB RAM. In this
section, we refer to our algorithm as \emph{Submod} and mean field as \emph{MF}.

%\begin{figure}
%       \Tree [.root [.{m0} 0-3 ] [.{m1} 4-7 ] [.{m2} 8-11 ] [.{m3} 12-15 ] [.{m4} 16-19 ]]
%   %    \Tree [.root [.{m0} 0 1 2 3 ] [.{m1} 4 5 6 7 ] [.{m2} 8 9 10 11 ] [.{m3} 12 13 14 15 ] [.{m4} 16 17 18 19 ]]
%%    \Tree[. [. [.1 .2 .3]]]
%\mycaption{\footnotesize \em An $r$-HST defining pairwise distance among 20 labels used for upper-bound comparison with TRW. $m0$ stands for meta-label 0, and so on.}
%\label{fig:syn_rhst}
%\end{figure}
\begin{figure}
\centering
\includegraphics[scale = 0.50]{./figures/synthetic_upper_bound_comparison.pdf}
\mycaption{\footnotesize \em Upper-bound comparison using synthetic data. The
plot shows the ratio (Submod bound - TRW bound)/|TRW bound| averaged over 100
unary instances as a function of pairwise weights using the worst-case optimal and
alternate extension for Potts and hierarchical Potts models. We observe that
the worst-case optimal extension
(solid) results in tighter bounds as compared to the respective alternate
extensions (dotted). Also, the worst-case optimal extension bounds are in
similar range as the TRW bounds.}
\label{fig:syn_plot}
\end{figure}
\begin{figure*}
    \centering
\begin{tabular}{cccccc}
        \includegraphics[scale = 0.14]{\imagePath/stereo/venus_GT} &
        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/venus_left_mf} &
        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/venus_left_submod} &
    %
        \includegraphics[scale = 0.19]{\imagePath/stereo/tsukuba_GT} &
        \includegraphics[scale = 0.19]{\imagePath/stereo/potts/tsukuba_left_mf} &
        \includegraphics[scale = 0.19]{\imagePath/stereo/potts/tsukuba_left_submod} \\
        \scriptsize(a) Venus GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution & \scriptsize(a) Tsukuba GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
        {} & \scriptsize  60.32s, 1.83e+07 &  \scriptsize 469.75s, {\bf 1.55e+07} & {} & \scriptsize  14.93s, 8.21e+06 & \scriptsize 215.22s, {\bf 4.12e+06} \\
         \includegraphics[scale = 0.14]{\imagePath/stereo/cones_GT} &
        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/cones_left_mf} &
        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/cones_left_submod} &
        %
       %
         \includegraphics[scale = 0.15]{\imagePath/stereo/teddy_GT} &
        \includegraphics[scale = 0.15]{\imagePath/stereo/potts/teddy_left_mf} &
        \includegraphics[scale = 0.15]{\imagePath/stereo/potts/teddy_left_submod} \\
    \scriptsize(a) Cones GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution & \scriptsize(a) Teddy GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
        {} & \scriptsize 239.14s, 2.68e+07 &  \scriptsize 1082.72s, {\bf 1.27e+07} & {} & \scriptsize 555.30s, 2.36e+07 & \scriptsize 1257.86s, {\bf 1.58e+07}\\
\end{tabular}
%\vspace{2mm}
\mycaption{\footnotesize \em Stereo matching using dense CRFs with Potts
compatibility and Gaussian pairwise potentials. We compare our solution with
the mean-field algorithm of Koltun and Krahenbuhl \citep{koltun2011efficient}. We observe that our method gives better-looking solutions with lower energy value at the cost of higher computational time.}
\label{fig:stereo}
\end{figure*}

\mysubsection{Upper-bound Comparison using Synthetic Data}
\label{subsec:synthetic}
\myparagraph{Data} We generate lattices of size 100 $\times$ 100, where each
lattice point represents a variable taking one of 20 labels. The pairwise
relations of the sparse CRFs are defined by 4-connected neighbourhoods. The
unary potentials are uniformly sampled in the range [0, 10]. We consider (a)
Potts model and (b) hierarchical Potts models with pairwise distance between
labels given by the trees of figure \ref{fig:syn_rhst}. The pairwise weights
are varied in the range $\{1, 2, 5, 10\}$. We compare the results of
our worst-case optimal submodular extension with an alternate submodular
extension as given in figure \ref{fig:alternate_extension}.

\myparagraph{Method} For our algorithm, we use the standard schedule $\gamma = 2/(k + 2)$ to obtain step size $\gamma$ at iteration $k$. We run our algorithm till convergence - 100 iterations suffice for this. The experiments are repeated for 100 randomly generated unaries for each model and each weight. For TRW, we used the MATLAB toolbox of \citep{domke2013learning}. The baseline code does not optimise over tree distributions. We varied the edge-appearance probability in trees over the range [0.1 - 0.5] and found 0.5 to give tightest upper bound.

\myparagraph{Results} We plot the ratio of the normalised difference of the upper
bound values of our method with TRW as a function of pairwise weights. The
ratios are averaged over 100 instances of unaries. Figure \ref{fig:syn_plot}
shows the plots for Potts and hierarchical Potts models for the worst-case
optimal and alternate extension. We find that the optimal extension (solid) results 
in tighter upper-bounds than the alternate extension (dotted) for both models.
This is because the representation of the submodular function using figure \ref{fig:alternate_extension} necessitates that $\phi_a(i)$ be non-negative. This implies that $F(A)$ values are larger
for the worst-case optimal extension of figure \ref{fig:optimal_extension_graph} as compared to the
alternate extension. Hence the minimisation problem \ref{eq:min_upper_bound}
has larger domain $EP(F)$ for the optimal extension, thereby resulting in better minima. Figure \ref{fig:syn_plot} also indicates that our algorithm with optimal extension provides similar range of upper bound as TRW, thereby providing empirical justification of our
method. TRW makes use of the standard LP relaxation \citep{chekuri2004linear} which is tighter than Kleinberg-Tardos relaxation, resulting in better approximation. However, TRW does not scale well with neighborhood size, thereby prohibiting its use in dense CRFs.

%==> Stereo_special_mf_1.640000_12.990000_50.840000_1.510000_1.100000/cones_left_mf_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
%239.143 2.67669e+07
%
%==> Stereo_special_mf_1.640000_12.990000_50.840000_1.510000_1.100000/teddy_left_mf_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
%555.304 2.35632e+07
%
%==> Stereo_special_mf_1.640000_12.990000_50.840000_1.510000_1.100000/tsukuba_left_mf_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
%14.9264 8.21469e+06
%
%==> Stereo_special_mf_1.640000_12.990000_50.840000_1.510000_1.100000/venus_left_mf_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
%60.3171 1.83186e+07
%
%==> Stereo_special_submod_1.640000_12.990000_50.840000_1.510000_1.100000/cones_left_submod_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
%1082.72 1.26817e+07
%
%==> Stereo_special_submod_1.640000_12.990000_50.840000_1.510000_1.100000/teddy_left_submod_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
%1257.86 1.57924e+07
%
%==> Stereo_special_submod_1.640000_12.990000_50.840000_1.510000_1.100000/tsukuba_left_submod_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
%215.219 4.12202e+06
%
%==> Stereo_special_submod_1.640000_12.990000_50.840000_1.510000_1.100000/venus_left_submod_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
%469.752 1.55122e+07
%\begin{minipage}
%\end{minipage}

\mysubsection{Stereo Matching using Dense CRFs}
\label{subsec:stereo}
\myparagraph{Data} We demonstrate the benefit our algorithm for stereo matching on images extracted from the Middlebury stereo matching dataset \citep{scharstein2001taxonomy}. We use dense CRF models with Potts compatibility term and Gaussian pairwise potentials. The unary terms are obtained using the absolute difference matching function of \citep{scharstein2001taxonomy}.

\myparagraph{Method} We use the implementation of mean-field algorithm for
dense CRFs of \citep{koltun2011efficient} as our baseline. For our algorithm,
we make use of the modified Gaussian filtering implementation for dense CRFs by
\citep{ajanthan2017efficient} to compute the conditional gradient at each step.
The step size $\gamma$ at each iteration is selected by doing line search. We
run our algorithm till 100 iterations, since the visual quality of the solution
does not show much improvement beyond this point. We run mean-field up to convergence, with a threshold of 0.001 for change in KL-divergence.

\myparagraph{Results}

Figure \ref{fig:stereo} shows some example solutions obtained by picking the
label with maximum marginal probability for each variable for mean-field and
for our algorithm. We also report the time and energy values of the solution
for both methods. Though we are not performing MAP estimation, energy values
give us a quantitative indication of the quality of solutions. For the full set of
21 image pairs (2006 dataset), the average ratio of the energies of the solutions from our method 
compared to mean-field is 0.943. The avearge time ratio is 10.66. We observe that our algorithm results in more natural looking stereo matching results with lower energy values for all images. However, mean-field runs faster than our method for each instance.

\mysection{Discussion}
\label{sec:conclude}
We have established the relation between submodular extension for the Potts
model and the LP relaxation for MAP estimation using Lovasz extension. This allowed us to identify the worst-case optimal submodular extension for Potts as well as the general metric labeling problems. It is worth noting that it might still be possible to obtain an improved submodular extension for a given problem instance. The design of a computationally feasible algorithm for this task is an interesting direction of future research. While our current work has focused on pairwise graphical models, it can be readily applied to high-order potentials by considering the corresponding LP relaxation objective as the Lovasz extension of a submodular extension. The identification of such extensions for popular high-order potentials such as the $P^n$ Potts model or its robust version could further improve the accuracy of important computer vision applications such as semantic segmentation.

\newpage

### 2018-11-30 tex/paper.tex
Details of transformed LP for hiearchical Potts added.

@@
 \end{align}
-The set $\Delta$ is the same domain as defined in equation (\ref{eq:delta}). We
-rewrite this LP relaxation using indicator variables $z_{ai}$ for all labels
-and meta-labels as
+The set $\Delta$ is the same domain as defined in equation (\ref{eq:delta}).
+
+\myparagraph{Transformed Tightest LP Relaxation} We take (T-LP) and rewrite it using indicator variables for all labels and meta-labels.
+Let $\mathcal R$ denote the set of all labels and meta-labels, that is, all nodes in the tree apart from the root.
+Also, let $\mathcal L$ denote the set of labels, that is, the leaves of the tree.
+Let $T_i$ denote the subtree which is rooted at the $i$-th node.
+We introduce an indicator variable $z_{ai} \in \{0, 1\}$, where
 \begin{align}
-    \text{(T-LP-FULL)} &\quad \min \widetilde{E}({\bf z}) \nonumber \\
-    {} &\text{such that} \quad {\bf z} \in \Delta'
+    \enskip  z_{ai} =  \begin{cases} 
+        y_{ai} \quad &\text{if} \enskip i \in {\mathcal L} \\
+        y_{a}(T_i) \quad &\text{if} \enskip i \in {\mathcal R - \mathcal L} \\
+    \end{cases}
+\end{align}
+
+We need to extend the definition of unary potentials to the expanded label space as follows:
+\begin{align}
+    \text{where} \enskip    \phi'_{a}(i) =  \begin{cases} 
+        \phi_{a}(i) \quad &\text{if} \enskip i \in {\mathcal L} \\
+        0  \quad &\text{if} \enskip i \in {\mathcal R - L} \\
+    \end{cases}
+\end{align}
+%
+We can now rewrite problem \eqref{eq:t-lp} in terms of new indicator variables $z_{ai}$:
+%
+\begin{align}
+    \text{(T-LP-FULL)} \quad &\min \widetilde{E}({\bf z}) = \sum_{i \in \mathcal{R}} \sum_{a \in {\mathcal X}} \phi_a'(i) \cdot z_{ai} + \nonumber \\
+            &\sum_{i \in \mathcal{R}} \sum_{(a, b) \in {\mathcal N}} w_{ab} \cdot l_{T_i} \cdot |z_{ai} - z_{bi}| \nonumber \\
+    \text{such that} \quad {\bf z} \in \Delta'
 \label{eq:t-lp-full}
@@
 \end{align}
-The details of the new relaxation (T-LP-FULL) can be found in the appendix.
+%
+Constraint (\ref{eq:consistency_constraint}) ensures consistency among labels and meta-labels, that is, if a label is assigned then all the meta-labels which lie on the path from the root to the label should be assigned as well.
 %

### 2018-11-30 tex/paper.tex
New section: higher-order Potts model described and LP relaxation specified.

@@
 
+\mysection{Higher-order Potts Model} The pairwise Potts model often fails
+to capture useful image statistics because it restricts the order of the
+potentials to be at most two. Higher order clique potentials can model complex
+interactions of random variables, and thereby overcome this difficulty.
+
+The simplest generalisation of pairwise Potts model to higher-order results in
+\textit{higher-order Potts model}. The clique potentials for these models are
+defined as
+
+\begin{equation}
+  \phi_c({\bf x}_c) = 
+  \begin{cases}
+    \gamma_0, &{\text{if}} \; \forall i, j \in c: x_i = x_j, \\
+    \gamma_1, &{\text{else}}. 
+  \end{cases} 
+\end{equation}
+
+\myparagraph{LP Relaxation}
+\begin{align}
+    \text{(HOP-LP)} \quad \min_{\bf y} \enskip & E({\bf y}) =  \sum_{a = 1}^N
+    \sum_{i = 1}^L \phi_a(i)y_{ai} + \nonumber\\
+    &\sum_{c \in {\mathcal C}} \max_{(a, b) \in c} \frac{w_c}{2} \sum_{i = 1}^L \cdot |y_{ai} - y_{bi}| \nonumber\\
+    \text{s.t} \quad &{\bf y} \in \Delta.
+\label{eq:hop-lp}
+\end{align}
+The constraints for (HOP-LP) are the same as for (P-LP).
+
 \mysection{Fast Conditional Gradient Computation for Dense CRFs}

### 2018-12-04 tex/paper.tex
Included discussion about extension and proof for higher-order potts model

@@
 
+\myparagraph{LP-based Submodular Extension} We use the same $1-of-L$ encoding
+scheme as for Potts model. In a similar way, problem can be rewritten as
+
+%Same as eq 5.4
+\begin{equation}
+\min_{{\bf s} \in EP(F)} \quad \sum_{a = 1}^{N} \log \sum_{i = 1}^L \exp(-s_{ai}).
+\label{eq:hop_upper_bound}
+\end{equation}
+
+Introducing a temperature parameter $T$ to problem (),
+
+%Same as eq 5.5
+\begin{equation}
+    \min_{{\bf s} \in EP(F)} \quad g_T ({\bf s}) = \sum_{a = 1}^{N} T \cdot \log \sum_{i = 1}^L \exp(-\frac{s_{ai}}{T}).
+\label{eq:hop_temp}
+\end{equation}
+
+and considering the limit $T \to 0$, problem () becomes
+
+%Same as eq 5.6
+\begin{align}
+    - \min_{{\bf y} \in \Delta} f({\bf y}) 
+\end{align}
+where $f(.)$ is the Lovasz extension of $F(.)$. We are interested in making
+this LP same as the (), thereby enabling us to recover the submodular extension
+for higher-order Potts. The following proposition gives us the submodular
+extension.
+{\proposition The Lovasz extension of the following submodular extension for
+  higher-order Potts model:
+  \begin{align}
+      F_{HOPotts}(A) &= \sum_a \sum_i \phi_{a}(i) [|A \cap \{v_{ai}\}| = 1] + \nonumber \\
+        &\sum_{c \in {\mathcal C}} \max_{(a, b) \in c} \frac{w_c}{2} \sum_{i = 1}^L [|A \cap \{v_{ai}, v_{bi}\}| = 1]
+\end{align}
+is the objective function $E({\bf y})$ of the LP relaxation (HOP-LP).
+\label{proposition:hop_extension}}
+\begin{proof}
+  We first observe that $\forall {\bf y} \in {\mathcal L}^N$, that is, for all
+  valid labelings, $F_{HOPotts}(A_{\bf y})$ equals $E({\bf y}$.  Hence, $F_{HOPotts}$
+    is an extension.
+
+Consider a clique with 3 variables $X_a, X_b, X_c$
+each of which can take one of three labels $i = 1, 2, 3$. Then ${\bf x}
+= [x_{a1}, x_{a2},x_{a3},x_{b1},x_{b2},x_{b3},x_{c1},x_{c2},x_{c3}]$ is the
+vector of relaxed variables. To compute $E({\bf x})$ and $f({\bf x})$, it will
+suffice to specify per label ordering. So let us assume $x_{a1} > x_{b1}
+> x_{c1}$, $x_{a2} > x_{b2} > x_{c2}$ and $x_{a3} > x_{b3} > x_{c3}$.
+
+Then, the relaxed objective is
+
+\begin{equation}
+  E({\bf x}) = \frac{w_c}{2} \cdot [(x_{a1} - x_{c1}) + (x_{a2} - x_{c2}) +(x_{a3} - x_{c3})  ]
+\end{equation}
+
+It can be shown that regardless of inter-labeling rankings, the L.E is given
+by:
+
+\begin{equation}
+  f({\bf x}) = \frac{w_c}{2} \cdot [(x_{a1} - x_{c1}) + (x_{a2} - x_{c2}) +(x_{a3} - x_{c3})  ]
+\end{equation}
+
+One can generalise this over more cliques, variables, and labels to show the equivalence of $E({\bf x})$ and $f({\bf x})$.
+The Lovasz extension of a set function is convex iff it is submodular. Since $E({\bf x})$
+is convex, this implies that $F_{HOPotts}$ is submodular. 
+\end{proof}
+
 \mysection{Fast Conditional Gradient Computation for Dense CRFs}

### 2018-12-04 tex/paper.tex
Proof for equivalence of conditional gradient and subgradient for dense crf included.

@@
 
-Introducing a temperature parameter $T$ to problem (),
+Introducing a temperature parameter $T$ to problem (\ref{eq:hop_upper_bound}),
 
@@
 
-and considering the limit $T \to 0$, problem () becomes
+and considering the limit $T \to 0$, problem (\ref{eq:hop_temp}) becomes
 
@@
 \newline(Proof in appendix).
-%\begin{proof}
-%From lemma \ref{lemma:lovasz_lp}, 
-%\begin{align}
-%    E({\bf y}) &= f({\bf y})  \nonumber \\
-%    {} &=  \max_{{\bf s} \in EP(F)} \langle {\bf y}, {\bf s} \rangle.\nonumber
-%\end{align}
-%
-%The subdifferential of $E({\bf y})$ at $\bf y_0$ is the `active' linear function at $\bf y_0$. Hence,
-%\begin{align}
-%    \partial E({\bf y})|_{\bf y = y_0} &=  \argmax_{{\bf s} \in EP(F)} \langle {\bf y_0}, {\bf s} \rangle
-%    \label{eq:subgrad}
-%\end{align}
-%Equation (\ref{eq:subgrad}) is equivalent to ${\bf s}^*$ computation in line \ref{line:cond_grad} of algorithm \ref{algo:fw_inference}, with ${\bf y_0} = - \nabla g({\bf s}_k)$.
-%\end{proof}
-%
+\begin{proof}
+  For the Potts model, we derived the worst-case optimal extension $F_{Potts}$ by
+  making its Lovasz extension $f({\bf y})$ same as the objective function
+    $E({\bf y})$ of the worst-case optimal LP relaxation. Hence, we have
+\begin{align}
+    E({\bf y}) &= f({\bf y})  \nonumber \\
+    {} &=  \max_{{\bf s} \in EP(F)} \langle {\bf y}, {\bf s} \rangle.\nonumber
+\end{align}
+
+The subdifferential of $E({\bf y})$ at $\bf y_0$ is the `active' linear function at $\bf y_0$. Hence,
+\begin{align}
+    \partial E({\bf y})|_{\bf y = y_0} &=  \argmax_{{\bf s} \in EP(F)} \langle {\bf y_0}, {\bf s} \rangle
+    \label{eq:subgrad}
+\end{align}
+Equation (\ref{eq:subgrad}) is equivalent to ${\bf s}^*$ computation in line \ref{line:cond_grad} of algorithm \ref{algo:fw_inference}, with ${\bf y_0} = - \nabla g({\bf s}_k)$.
+\end{proof}
 

### 2018-12-04 tex/appendix.tex
Including appendix.

%\title{Appendix: Worst-case Optimal Submodular Extensions for Marginal Estimation}
\appendix
\mysection{Proofs for Potts Model Extension}

\paragraph{Remark 1} We show using induction over the number of variables that with $1$-of-$L$ encoding for Potts,

\begin{equation}
    \sum_{A \in {\mathcal M}} \exp(-s(A)) = \prod_{a = 1}^{N} \sum_{i = 1}^L \exp(-s_{ai})
    %%\label{eq:ind}
\end{equation}

Let $t$ be the number of variables, $V^t$ be the corresponding ground set and ${\mathcal M}^t$ be the sets corresponding to valid labelings. Equation \eqref{eq:ind} clearly holds for $t = 1$.

Let us assume that the relation holds for $t = N$, that is,

\begin{equation}
    \sum_{A^N \in {\mathcal M}^N} \exp(-s(A^N)) = \prod_{a = 1}^{N} \sum_{i = 1}^L \exp(-s_{ai})
\end{equation}

For $t = N + 1$,

\begin{align}
    \sum_{A^{N + 1} \in {\mathcal M}^{N + 1}} \exp(-s(A^{N + 1})) &=  \sum_{i = 1}^L \sum_{A^N \in {\mathcal M}^{N}} \exp(-s(A^N) - s_{N+1, i}) \nonumber \\
                                                                  &= \sum_{i = 1}^L \exp(-s_{N+1, i}) \sum_{A^N \in {\mathcal M}^{N}} \exp(-s(A^N) ) \nonumber \\
                                                                  &= \sum_{i = 1}^L \exp(-s_{N+1, i}) \prod_{a = 1}^{N} \sum_{i = 1}^L \exp(-s_{ai}) \nonumber \\
                                                                  &= \prod_{a = 1}^{N + 1} \sum_{i = 1}^L \exp(-s_{ai})
\end{align}

{\lemma Given any submodular extension $F(.)$ of a Potts energy function $E(.)$, its Lovasz extension $f(.)$ defines an LP relaxation of the MAP problem for $E(.)$ as 
\begin{align}
    \min_{{\bf y} \in \Delta} f({\bf y}).
\end{align}
%\label{lemma:lovasz_lp}}
\begin{proof}
    By definition of a submodular extension and the Lovasz extension, $E({\bf x}) = F(A_{\bf x}) = f(1_{A_{\bf x}})$ for all valid labelings ${\bf x}$. Also, from property 1, $f({\bf y})$ is maximum of linear functions. Hence, $f({\bf y})$ is a piecewise linear relaxation of $E({\bf x})$.

    The domain $\Delta$ is a polytope formed by union of $N$ probability simplices
\begin{equation}
    \Delta = \{{\bf y}_a \in \mathbb{R}^L | {\bf y}_a \succeq 0  \textrm{ and } \langle {\bf 1}, {\bf y}_a \rangle = 1\}
\end{equation}
With objective as maximum of linear functions and domain as a polytope, we have an LP relaxation of the corresponding MAP problem.
\end{proof}

{\proposition In the limit $T \to 0$, the following problem for Potts energies 
\begin{equation}
    \min_{{\bf s} \in EP(F)} \quad g_T ({\bf s}) = \sum_{a = 1}^{N} T \cdot \log \sum_{i = 1}^L \exp(-\frac{s_{ai}}{T}).
%\label{eq:potts_temp}
\end{equation}

becomes
\begin{align}
    - \min_{{\bf y} \in \Delta} f({\bf y}) 
\end{align}
%\label{proposition:potts_equiv}}

{\proposition The objective function $E({\bf y})$ of the LP relaxation (P-LP) is the Lovasz extension of $F_{Potts}(A) = \sum_{i = 1}^L F_i(A)$, where
\begin{align}
    F_i(A) &= \sum_a \phi_{a}(i) [|A \cap \{v_{ai}\}| = 1] + \nonumber \\
           &\sum_{(a, b) \in {\mathcal N}} \frac{w_{ab}}{2} \cdot [|A \cap \{v_{ai}, v_{bi}\}| = 1]
\end{align}
%\label{proposition:a-potts_optimal}}

\begin{proof}
Since $F_{Potts}$ is sum of Ising models $F_i$, we first focus on a particular label $i$ and then generalize. Consider a graph with only two variables $X_a$ and $X_b$ with an edge between them. The ground set in this case is $\{v_{ai}, v_{bi}\}$. Let the corresponding relaxed indicator variables be ${\bf y} = \{y_{aj}, y_{bj}\}$, such that $y_{ai}, y_{bi} \in [0, 1]$ and assume $y_{ai} > y_{bi}$. The Lovasz extension in this case is:

\begin{align}
f({\bf y}) &= y_{ai} \cdot [F_i(\{v_{ai}\}) - F_i(\{\})] + y_{bi} \cdot [F_i(\{v_{ai}, v_{bi}\}) - F_i(\{v_{ai}\})] \nonumber \\
&= y_{ai} \cdot [\left(\phi_{a}\left(j\right) + \frac{w_{ab}}{2}\right) - 0] + y_{bi} \cdot [\left(\phi_{a}\left(j\right) + \phi_{b}\left(j\right)\right) - \left(\phi_{a}\left(j\right) + \frac{w_{ab}}{2}\right)] \nonumber \\
&=  \phi_{a}\left(j\right) \cdot y_{ai} + \phi_{b}\left(j\right) \cdot y_{bi} + \frac{w_{ab}}{2} \cdot \left(y_{ai} - y_{bi}\right)
\end{align}

In general for both orderings of $y_{ab}$ and $y_{bi}$, we can write
%
\begin{equation}
f({\bf y}) = \phi_{a}(j) \cdot y_{ai} + \phi_{b}(j) \cdot y_{bi} + \frac{w_{ab}}{2} \cdot |y_{ai} - y_{bi}|
%\label{eq:a-lovasz_ising}
\end{equation}
%
Extending the Lovasz extension of equation \eqref{eq:a-lovasz_ising} to all variables and all labels gives us $E({\bf y})$ in (P-LP). 
\end{proof}

\mysection{Proofs for Hierarchical Potts Model Extension} 
\myparagraph{Transformed Tightest LP Relaxation} We take (T-LP) and rewrite it using indicator variables for all labels and meta-labels. Let $\mathcal R$ denote the set of all labels and meta-labels, that is, all nodes in the tree apart from the root. Also, let $\mathcal L$ denote the set of labels, that is, the leaves of the tree. Let $T_i$ denote the subtree which is rooted at the $i$-th node. We introduce an indicator variable $z_{ai} \in \{0, 1\}$, where
\begin{align}
    \enskip  z_{ai} =  \begin{cases} 
        y_{ai} \quad &\text{if} \enskip i \in {\mathcal L} \\
        y_{a}(T_i) \quad &\text{if} \enskip i \in {\mathcal R - \mathcal L} \\
    \end{cases}
\end{align}

We need to extend the definition of unary potentials to the expanded label space as follows:
\begin{align}
    \text{where} \enskip    \phi'_{a}(i) =  \begin{cases} 
        \phi_{a}(i) \quad &\text{if} \enskip i \in {\mathcal L} \\
        0  \quad &\text{if} \enskip i \in {\mathcal R - L} \\
    \end{cases}
\end{align}
%
We can now rewrite problem \eqref{eq:t-lp} in terms of new indicator variables $z_{ai}$:
%
\begin{align}
    \text{(T-LP-FULL)} \quad &\min \widetilde{E}({\bf z}) = \sum_{i \in \mathcal{R}} \sum_{a \in {\mathcal X}} \phi_a'(i) \cdot z_{ai} + \nonumber \\
            &\sum_{i \in \mathcal{R}} \sum_{(a, b) \in {\mathcal N}} w_{ab} \cdot l_{T_i} \cdot |z_{ai} - z_{bi}| \nonumber \\
    \text{such that} \quad {\bf z} \in \Delta'
%\label{eq:t-lp-full}
\end{align}
   where $\Delta'$ is the convex hull of the vectors satisfying
\begin{align}
    \enskip &\sum_{i \in \mathcal{L}} z_{ai} = 1, \enskip z_{ai} \in \{0, 1\} \enskip \forall a \in {\mathcal X}, i \in \mathcal{L} \\
    \text{and} \enskip &z_{ai} = \sum_{j \in L(T_i) } z_{aj}.\enskip \forall a \in {\mathcal X}, i \in \mathcal{R - L} %\label{eq:consistency_constraint} 
\end{align}
%
Constraint (\ref{eq:consistency_constraint}) ensures consistency among labels and meta-labels, that is, if a label is assigned then all the meta-labels which lie on the path from the root to the label should be assigned as well. We are now going to identify a suitable set encoding and the worst-case optimal submodular extension using (T-LP-FULL).

{\lemma Given any submodular extension $F(.)$ of an hierachical Potts energy function $E(.)$, its Lovasz extension $f(.)$ defines an LP relaxation of the corresponding MAP estimation problem as
\begin{align}
    \min_{{\bf z} \in \Delta'} f({\bf z}) 
\end{align}
%\label{lemma:lovasz_lp_metric}}

\begin{proof}
    By definition of a submodular extension and the Lovasz extension, $E({\bf x}) = F(A_{\bf x}) = f(1_{A_{\bf x}})$ for all valid labelings ${\bf x}$. Also, from property 1, $f({\bf y})$ is maximum of linear functions. Hence, $f({\bf y})$ is a piecewise linear relaxation of $E({\bf x})$.

    We can write the domain $\Delta'$ as 
    \begin{equation}
    \Delta' = \{ {\bf y}_a \in \mathbb{R}^M | {\bf y}_a \succeq 0, \enskip \langle {\bf 1}, {\bf y}_a^{label} \rangle = 1, \enskip {\bf y}_a(p_{ai}) = {\bf 1} \textrm{ or } {\bf y}_a(p_{ai}) = {\bf 0} \forall i \in [1, L]\}
    \end{equation} 
where ${\bf y}_a$ is the component of $\bf y$ corresponding to the $a$-th variable, ${\bf y}_a^{label}$  is the component of ${\bf y}_a$ corresponding to the $L$ labels, and ${\bf y}_a(p_{ai})$ is the component of ${\bf y}_a$ corresponding to the elements of $p_{ai}$.

    Since $\Delta'$ is defined by linear equalities and inequalities, it is a polytope. With objective as maximum of linear functions and domain as a polytope, we have an LP relaxation of the corresponding MAP problem.
\end{proof}

{\proposition In the limit $T \to 0$, the following problem for hierarchical Potts energies

\begin{equation}
\min_{{\bf s} \in EP(F)} \quad g_T ({\bf s}) = \sum_{a = 1}^{N} T \cdot \log \sum_{i = 1}^L \exp(-\frac{s'_{ai}}{T}).
%\label{eq:metric_temp}
\end{equation}
becomes:
\begin{align}
    - \min_{{\bf z} \in \Delta'} f({\bf z}) 
\end{align}
%\label{proposition:metric_equiv}}

\begin{proof}
In the limit of $T \to 0$, we can rewrite the above problem as
\begin{equation}
    \min_{{\bf s} \in EP(F)} \quad \sum_{a = 1}^{N} \max_{i} (-s'_{ai}).
\end{equation}
In vector form, the problem becomes
\begin{align}
    &\min_{{\bf s} \in EP(F)} \max_{{\bf z} \in \Delta} - \langle {\bf z}, {\bf s'} \rangle \\
    & = - \max_{{\bf s} \in EP(F)} \min_{{\bf z} \in \Delta} \langle {\bf z}, {\bf s'} \rangle
    %\label{eq:metric_maxmin}
\end{align}
\begin{equation}
    \text{where } \Delta = \{{\bf z}_a \in \mathbb{R}^L | {\bf z}_a \succeq 0  \textrm{ and } \langle {\bf 1}, {\bf z}_a \rangle = 1\}
\end{equation}
where ${\bf z}_a$ is the component of $\bf z$ corresponding to the $a$-th variable. We can unpack $\bf s'$ using 
\begin{equation}
s'_{ai} = \sum_{t \in p_{ai}} s_t.
%\label{eq:tree_s}
\end{equation}
and rewrite problem \eqref{eq:metric_maxmin} as
\begin{align}
    - \max_{{\bf s} \in EP(F)} \min_{{\bf y} \in \Delta'} \langle {\bf y}, {\bf s} \rangle
    %\label{eq:metric_maxmin_unpack}
\end{align}
The new constraint set $\Delta'$ ensures that the binary entries of labels and meta-labels is consistent:
\begin{align}
    \textrm{where } \Delta' &= \{ {\bf y}_a \in \mathbb{R}^M | {\bf y}_a \succeq 0, \enskip \langle {\bf 1}, {\bf y}_a^{label} \rangle = 1, \nonumber \\
    &\enskip {\bf y}_a(p_{ai}) = {\bf 1} \textrm{ or } {\bf y}_a(p_{ai}) = {\bf 0} \forall i \in [1, L]\}
\end{align} 
where ${\bf y}_a$ is the component of $\bf y$ corresponding to the $a$-th variable, ${\bf y}_a^{label}$  is the component of ${\bf y}_a$ corresponding to the $L$ labels, and ${\bf y}_a(p_{ai})$ is the component of ${\bf y}_a$ corresponding to the elements of $p_{ai}$.

By the minimax theorem for LP, we can reorder the terms:
\begin{equation}
    - \min_{{\bf y} \in \Delta'} \max_{{\bf s} \in EP(F)} \langle {\bf y}, {\bf s} \rangle 
    %\label{eq:metric_minmax}
\end{equation}
Recall that $\max_{{\bf s} \in EP(F)} \langle {\bf y}, {\bf s} \rangle$ is the value of the Lovasz extension of $F$ at $\bf y$, that is, $f({\bf y})$. Hence, as $T \to 0$, the marginal inference problem converts to minimising the Lovasz extension under the constraints $\Delta'$:
\begin{equation}
    - \min_{{\bf y} \in \Delta'} f({\bf y}) 
    %\label{eq:metric_minLovasz}
\end{equation}
\end{proof}

{\proposition The objective function $\widetilde{E}({\bf z})$ of (T-LP-FULL) is the Lovasz extension of $F_{r-\textrm{HST}}(A) = \sum_{i = 1}^M F_i(A)$, where
\begin{align}
    F_i(A) &= \sum_a \phi'_{a}(i) [|A \cap \{v_{ai}\}| = 1] + \nonumber \\
           &\sum_{(a, b) \in {\mathcal N}} {w_{ab}} \cdot l_{T_i} \cdot [|A \cap \{v_{ai}, v_{bi}\}| = 1]
\end{align}
%\label{proposition:rhst_optimal}}
\begin{proof}
    We observe that $F_{r-\textrm{HST}}$ is of exactly the same form as $F_{Potts}$, except that the Ising models $F_i$ are defined over not just labels, but meta-labels as well. Using the same logic as in the proof of proposition \ref{proposition:a-potts_optimal}, each $F_i$ is the Lovasz extension of 
        \begin{equation}
            \widetilde{E}_i({\bf z}) = \left( \sum_{a \in {\mathcal X}} \phi_a'(i) \cdot z_{ai} + \sum_{(a, b) \in {\mathcal N}} w_{ab} \cdot l_{T_i} \cdot |z_{ai} - z_{bi}| \right)
        \end{equation}
    and the results follows. 
\end{proof}

\section{Additional Experiment Results}

\begin{figure*}[!]
    \centering
\begin{tabular}{cc}
    \includegraphics[scale = 0.23]{./figures/synthetic_tree_2.png} &
\includegraphics[scale = 0.35]{./figures/synthetic_upper_bound_tree2.pdf} \\
    \scriptsize(a)  & \scriptsize(b) \\ 
\end{tabular}
\mycaption{\footnotesize \em (a) The hierarchical Potts model defining pairwise distance among 20 labels used for upper-bound comparison with TRW. Blue nodes are the meta-labels and yellow nodes are labels. (b) Upper-bound comparison using synthetic data. The plot shows the ratio of upper bound value of log partition function from our algorithm to that of TRW averaged over 100 unary instances as a function of pairwise weights. We observe that our method results in comparable upper bound values to TRW.}
%\label{fig:syn_plot}
\end{figure*}

\mysubsection{Upper-bound Comparison using Synthetic Data with Different Tree}
%\label{subsec:synthetic}
\myparagraph{Data}  We consider hierarchical Potts model with pairwise distance between labels given by the tree of figure \ref{fig:syn_plot} (a). We vary the pairwise weight in the range $\{1, 2, 5, 10\}$. The rest of the setting is same as in the sythetic results subsection of the main paper.

\myparagraph{Method} We use the same method as described in the sythetic results subsection of the main paper.

\myparagraph{Results} We plot the ratio of upper bound values of our method to that of TRW as a function of pairwise weights. The ratios are averaged over 100 instances of unaries. Figure \ref{fig:syn_plot} (b) shows the plots for hierarchical Potts models for the new tree. Similar to the first tree (in main paper), we find that our algorithm results in similar range of upper bound as TRW, thereby providing further empirical justification of our method.

\begin{figure*}[!]
	\centering
\begin{tabular}{ccc}
        \includegraphics[scale = 0.25]{\imagePath/stereo/venus_left} &
        \includegraphics[scale = 0.25]{\imagePath/stereo/potts/venus_left_mf_marginals} &
        \includegraphics[scale = 0.25]{\imagePath/stereo/potts/venus_left_submod_marginals} \\
        \scriptsize(a) Venus left image & \scriptsize(b) MF entropy & \scriptsize(c) Submod entropy \\
    %
        \includegraphics[scale = 0.30]{\imagePath/stereo/tsukuba_left} &
        \includegraphics[scale = 0.30]{\imagePath/stereo/potts/tsukuba_left_mf_marginals} &
        \includegraphics[scale = 0.30]{\imagePath/stereo/potts/tsukuba_left_submod_marginals} \\
        \scriptsize(a) Tsukuba left image & \scriptsize(b) MF entropy & \scriptsize(c) Submod entropy\\
         \includegraphics[scale = 0.25]{\imagePath/stereo/cones_left} &
        \includegraphics[scale = 0.25]{\imagePath/stereo/potts/cones_left_mf_marginals} &
        \includegraphics[scale = 0.25]{\imagePath/stereo/potts/cones_left_submod_marginals} \\
        \scriptsize(a) Cones left image & \scriptsize(b) MF entropy & \scriptsize(c) Submod entropy \\ 
        %
       %
         \includegraphics[scale = 0.26]{\imagePath/stereo/teddy_left} &
        \includegraphics[scale = 0.26]{\imagePath/stereo/potts/teddy_left_mf_marginals} &
        \includegraphics[scale = 0.26]{\imagePath/stereo/potts/teddy_left_submod_marginals} \\
        \scriptsize(a) Teddy left image & \scriptsize(b) MF entropy & \scriptsize(c) Submod entropy\\
\end{tabular}
\vspace{2mm}
\mycaption{\footnotesize \em Entropy plots for stereo matching using dense CRFs with Potts compatibility and Gaussian pairwise potentials for the mean-field algorithm of \citet{koltun2011efficient} and our method. We observe that our method gives more detailed uncertainty information.} 
%\label{fig:potts_entropy}
\end{figure*}

\mysubsection{Entropy Plots for Stereo Matching using Dense CRFs with Potts Model}
We plotted the entropy of the marginals obtained from mean-field and our inference algorithm for the Potts models for all the stereo matching instances. We observe that our mean-field does not give enough uncertainty information, whereas our method gives much more detailed information. However, we also observe that whereas mean-field is too certain (has less entropy) almost everywhere, our method results in uncertainty even in those regions where we expect good marginals to be certain, such as in the interior regions of objects.

\begin{figure*}[!]
    \centering
    \includegraphics[scale = 0.33]{./figures/synthetic_tree.png}
    \mycaption{\footnotesize \em Tree used to specify hierachical Potts potentials for stereo matching for `Venus' image having disparity values 1-20. Similar tree topology was used for the other images.}
    %\label{fig:stereo_tree}
\end{figure*}

%==> cones_left_cones_tree1_mf.txt <==
%46.3463 3.72918e+06
%
%==> teddy_left_teddy_tree1_mf.txt <==
%62.7843 2.45632e+06
%
%==> tsukuba_left_tsukuba_tree1_mf.txt <==
%5.59413 846057
%
%==> venus_left_venus_tree1_mf.txt <==
%14.3497 1.87249e+06
%
%==> cones_left_cones_tree1_submod_tree.txt <==
%2100.14 2.62802e+07
%
%==> teddy_left_teddy_tree1_submod_tree.txt <==
%2191.12 3.46133e+07
%
%==> tsukuba_left_tsukuba_tree1_submod_tree.txt <==
%359.85  6.60389e+06
%
%==> venus_left_venus_tree1_submod_tree.txt <==
%712.88  1.55201e+07

\begin{figure*}[!]
	\centering
\begin{tabular}{ccc}
        \includegraphics[scale = 0.25]{\imagePath/stereo/venus_GT} &
        \includegraphics[scale = 0.25]{\imagePath/stereo/tree/venus_left_venus_tree1_mf} &
        \includegraphics[scale = 0.25]{\imagePath/stereo/tree/venus_left_venus_tree1_submod} \\ 
        \scriptsize(a) Venus GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution \\ 
        {} & \scriptsize 14.35 {\bf 1.87e+06} &  \scriptsize 712.88  15.52e+07\\

        \includegraphics[scale = 0.30]{\imagePath/stereo/tsukuba_GT} &
        \includegraphics[scale = 0.30]{\imagePath/stereo/tree/tsukuba_left_tsukuba_tree1_mf} &
        \includegraphics[scale = 0.30]{\imagePath/stereo/tree/tsukuba_left_tsukuba_tree1_submod} \\
        \scriptsize(a) Tsukuba GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
        {} & \scriptsize 5.59 {\bf 0.85e+06} & \scriptsize 359.85  6.60e+06\\

        \includegraphics[scale = 0.25]{\imagePath/stereo/cones_GT} &
        \includegraphics[scale = 0.25]{\imagePath/stereo/tree/cones_left_cones_tree1_mf} &
        \includegraphics[scale = 0.25]{\imagePath/stereo/tree/cones_left_cones_tree1_submod} \\
        \scriptsize(a) Cones GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution \\
        {} & \scriptsize 46.34 {\bf 3.73e+06} &  \scriptsize 2100.14 26.28e+06\\

        \includegraphics[scale = 0.26]{\imagePath/stereo/teddy_GT} &
        \includegraphics[scale = 0.26]{\imagePath/stereo/tree/teddy_left_teddy_tree1_mf} &
        \includegraphics[scale = 0.26]{\imagePath/stereo/tree/teddy_left_teddy_tree1_submod} \\
        \scriptsize(a) Teddy GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
        {} & \scriptsize 62.78 {\bf 2.46e+06} & \scriptsize 2191.12 34.63e+06\\

\end{tabular}

\mycaption{\footnotesize \em Stereo matching using dense CRFs with hierarchical Potts compatibility given by trees similar to figure \ref{fig:stereo_tree} and Gaussian pairwise potentials. We compare our solution with the mean-field algorithm of \citet{koltun2011efficient}. We observe that our method gives better-looking solutions, though mean-field runs faster and results in lower energy solutions.}
%\label{fig:stereo_hier}
\end{figure*}

\begin{figure*}[!]
	\centering
\begin{tabular}{ccc}
        \includegraphics[scale = 0.25]{\imagePath/stereo/venus_left} &
        \includegraphics[scale = 0.25]{\imagePath/stereo/tree/venus_left_venus_tree1_mf_marginals} &
        \includegraphics[scale = 0.25]{\imagePath/stereo/tree/venus_left_venus_tree1_submod_marginals} \\ 
        \scriptsize(a) Venus left image & \scriptsize(b) MF entropy & \scriptsize(c) Submod entropy \\

        \includegraphics[scale = 0.30]{\imagePath/stereo/tsukuba_left} &
        \includegraphics[scale = 0.30]{\imagePath/stereo/tree/tsukuba_left_tsukuba_tree1_mf_marginals} &
        \includegraphics[scale = 0.30]{\imagePath/stereo/tree/tsukuba_left_tsukuba_tree1_submod_marginals} \\
        \scriptsize(a) Tsukuba left image & \scriptsize(b) MF entropy & \scriptsize(c) Submod entropy\\

        \includegraphics[scale = 0.25]{\imagePath/stereo/cones_left} &
        \includegraphics[scale = 0.25]{\imagePath/stereo/tree/cones_left_cones_tree1_mf_marginals} &
        \includegraphics[scale = 0.25]{\imagePath/stereo/tree/cones_left_cones_tree1_submod_marginals} \\
        \scriptsize(a) Cones left image & \scriptsize(b) MF entropy & \scriptsize(c) Submod entropy \\

        \includegraphics[scale = 0.26]{\imagePath/stereo/teddy_left} &
        \includegraphics[scale = 0.26]{\imagePath/stereo/tree/teddy_left_teddy_tree1_mf_marginals} &
        \includegraphics[scale = 0.26]{\imagePath/stereo/tree/teddy_left_teddy_tree1_submod_marginals} \\
        \scriptsize(a) Teddy left image & \scriptsize(b) MF entropy & \scriptsize(c) Submod entropy\\
\end{tabular}

\mycaption{\footnotesize \em Entropy plots for stereo matching using dense CRFs with hiearchical Potts compatibility given by trees similar to figure \ref{fig:stereo_tree} and Gaussian pairwise potentials for the mean-field algorithm of \citet{koltun2011efficient} and our method. We observe that our method gives more detailed uncertainty information.}
%\label{fig:stereo_hier}
\end{figure*}

\mysubsection{Stereo Matching using Dense CRFs with Hierarchical Potts}
%\label{subsec:stereo_hier}

\myparagraph{Data} We demonstrate the benefit our algorithm for stereo matching using hierarchical Potts model. Similar to Potts case, we use images extracted from the Middlebury stereo matching dataset \citep{scharstein2001taxonomy}. We use dense CRF models with hierarchical Potts compatibility term and Gaussian pairwise potentials. The hierarchical Potts was specified using tree similar to figure \ref{fig:stereo_tree} for all images. The unary terms are obtained using the absolute difference matching function of \citep{scharstein2001taxonomy}.

\myparagraph{Method} We used the implementation of mean-field algorithm for dense CRFs of \citep{koltun2011efficient} as our baseline. For our algorithm, we made use of the modified Gaussian filtering implementation for dense CRFs by \citep{ajanthan2017efficient} to compute the conditional gradient at each step. The step size $\gamma$ at each iteration was selected by doing line search. We ran our algorithm till 100 iterations, since the visual quality of the solution did not show much improvement beyond this point. We ran mean-field up to convergence, with a threshold of 0.001 for change in KL-divergence.

\myparagraph{Results}

Figure \ref{fig:stereo_hier} shows the solutions obtained by picking the label with maximum marginal probability for each variable for mean-field and for our algorithm. We also report the time and energy values of the solution for both methods. Though we are not performing MAP estimation, energy values give us a quantitative indication of the quality of solutions. We observe that our algorithm results in more natural looking stereo matching results. However, mean-field gives lower energy value solutions and runs faster than our method for each instance.
\clearpage

### 2018-12-04 tex/paper.tex
Including appendix.

@@
 
-%==> Stereo_special_mf_1.640000_12.990000_50.840000_1.510000_1.100000/cones_left_mf_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
-%239.143 2.67669e+07
-%
-%==> Stereo_special_mf_1.640000_12.990000_50.840000_1.510000_1.100000/teddy_left_mf_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
-%555.304 2.35632e+07
-%
-%==> Stereo_special_mf_1.640000_12.990000_50.840000_1.510000_1.100000/tsukuba_left_mf_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
-%14.9264 8.21469e+06
-%
-%==> Stereo_special_mf_1.640000_12.990000_50.840000_1.510000_1.100000/venus_left_mf_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
-%60.3171 1.83186e+07
-%
-%==> Stereo_special_submod_1.640000_12.990000_50.840000_1.510000_1.100000/cones_left_submod_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
-%1082.72 1.26817e+07
-%
-%==> Stereo_special_submod_1.640000_12.990000_50.840000_1.510000_1.100000/teddy_left_submod_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
-%1257.86 1.57924e+07
-%
-%==> Stereo_special_submod_1.640000_12.990000_50.840000_1.510000_1.100000/tsukuba_left_submod_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
-%215.219 4.12202e+06
-%
-%==> Stereo_special_submod_1.640000_12.990000_50.840000_1.510000_1.100000/venus_left_submod_1.640000_12.990000_50.840000_1.510000_1.100000.txt <==
-%469.752 1.55122e+07
-%\begin{minipage}
-%\end{minipage}
-
 \mysubsection{Stereo Matching using Dense CRFs}
@@
 
-\myparagraph{Results}
-
-Figure \ref{fig:stereo} shows some example solutions obtained by picking the
+\myparagraph{Results} Figure \ref{fig:stereo} shows some example solutions obtained by picking the
 label with maximum marginal probability for each variable for mean-field and
@@
 However, mean-field runs faster than our method for each instance.
+
+\mysubsection{Stereo Matching using Higher-order Potts Model}
+\label{subsec:stereo_hop}
+\myparagraph{Data} Next, we use the higher-order Potts model for stereo
+matching on the Middlebury dataset. For this
+purpose, we take a dense CRF model with Potts compatibility term and Gaussian pairwise
+potentials, and augment it with higher-order Potts term.
+
+\myparagraph{Method} We again use the modified Gaussian filtering implementation for dense CRFs by
+\citep{ajanthan2017efficient} to compute the conditional gradient at each step.
+To compute the contribution of the higher-order Potts, we make use of an
+efficient strategy to reuse computation. Recall that computing the conditional
+gradient coordinates $s^*_{\sigma(i)}$ requires us to obtain differences of the
+nested sets $F(S_i) - F(S_{i - 1})$. We store the clique state at each $i$, and
+this enables us to compute the difference for each coordinate in a constant
+amount of time. In this case, computation of conditional gradient takes
+$\mathcal{O}(cNL)$ where $c$ is the number of hihger-order cliques.
+
+\myparagraph{Results} Figure \ref{fig:stereo_hop} shows some example solutions obtained by picking the
+label with maximum marginal probability for each variable.
 

### 2018-12-04 tex/appendix.tex
Edited the appendix section for Potts model.

@@
     \sum_{A \in {\mathcal M}} \exp(-s(A)) = \prod_{a = 1}^{N} \sum_{i = 1}^L \exp(-s_{ai})
-    %%\label{eq:ind}
+    \label{eq:ind}
 \end{equation}
 
+\begin{proof}
 Let $t$ be the number of variables, $V^t$ be the corresponding ground set and ${\mathcal M}^t$ be the sets corresponding to valid labelings. Equation \eqref{eq:ind} clearly holds for $t = 1$.
@@
 \end{align}
-
+\end{proof}
 {\lemma Given any submodular extension $F(.)$ of a Potts energy function $E(.)$, its Lovasz extension $f(.)$ defines an LP relaxation of the MAP problem for $E(.)$ as 
@@
 With objective as maximum of linear functions and domain as a polytope, we have an LP relaxation of the corresponding MAP problem.
-\end{proof}
-
-{\proposition In the limit $T \to 0$, the following problem for Potts energies 
-\begin{equation}
-    \min_{{\bf s} \in EP(F)} \quad g_T ({\bf s}) = \sum_{a = 1}^{N} T \cdot \log \sum_{i = 1}^L \exp(-\frac{s_{ai}}{T}).
-%\label{eq:potts_temp}
-\end{equation}
-
-becomes
-\begin{align}
-    - \min_{{\bf y} \in \Delta} f({\bf y}) 
-\end{align}
-%\label{proposition:potts_equiv}}
-
-{\proposition The objective function $E({\bf y})$ of the LP relaxation (P-LP) is the Lovasz extension of $F_{Potts}(A) = \sum_{i = 1}^L F_i(A)$, where
-\begin{align}
-    F_i(A) &= \sum_a \phi_{a}(i) [|A \cap \{v_{ai}\}| = 1] + \nonumber \\
-           &\sum_{(a, b) \in {\mathcal N}} \frac{w_{ab}}{2} \cdot [|A \cap \{v_{ai}, v_{bi}\}| = 1]
-\end{align}
-%\label{proposition:a-potts_optimal}}
-
-\begin{proof}
-Since $F_{Potts}$ is sum of Ising models $F_i$, we first focus on a particular label $i$ and then generalize.
-Consider a graph with only two variables $X_a$ and $X_b$ with an edge between them.
-The ground set in this case is $\{v_{ai}, v_{bi}\}$.
-Let the corresponding relaxed indicator variables be ${\bf y} = \{y_{aj}, y_{bj}\}$, such that $y_{ai}, y_{bi} \in [0, 1]$ and assume $y_{ai} > y_{bi}$.
-The Lovasz extension in this case is:
-
-\begin{align}
-f({\bf y}) &= y_{ai} \cdot [F_i(\{v_{ai}\}) - F_i(\{\})] + y_{bi} \cdot [F_i(\{v_{ai}, v_{bi}\}) - F_i(\{v_{ai}\})] \nonumber \\
-&= y_{ai} \cdot [\left(\phi_{a}\left(j\right) + \frac{w_{ab}}{2}\right) - 0] + y_{bi} \cdot [\left(\phi_{a}\left(j\right) + \phi_{b}\left(j\right)\right) - \left(\phi_{a}\left(j\right) + \frac{w_{ab}}{2}\right)] \nonumber \\
-&=  \phi_{a}\left(j\right) \cdot y_{ai} + \phi_{b}\left(j\right) \cdot y_{bi} + \frac{w_{ab}}{2} \cdot \left(y_{ai} - y_{bi}\right)
-\end{align}
-
-In general for both orderings of $y_{ab}$ and $y_{bi}$, we can write
-%
-\begin{equation}
-f({\bf y}) = \phi_{a}(j) \cdot y_{ai} + \phi_{b}(j) \cdot y_{bi} + \frac{w_{ab}}{2} \cdot |y_{ai} - y_{bi}|
-%\label{eq:a-lovasz_ising}
-\end{equation}
-%
-Extending the Lovasz extension of equation \eqref{eq:a-lovasz_ising} to all variables and all labels gives us $E({\bf y})$ in (P-LP). 
 \end{proof}
@@
 \end{proof}
-
+\iffalse
 \section{Additional Experiment Results}
@@
 However, mean-field gives lower energy value solutions and runs faster than our method for each instance.
+\fi
 \clearpage

### 2018-12-04 tex/appendix.tex
Edited appendix - section on hierarchical Potts model.

@@
 \mysection{Proofs for Hierarchical Potts Model Extension}
-\myparagraph{Transformed Tightest LP Relaxation} We take (T-LP) and rewrite it using indicator variables for all labels and meta-labels.
-Let $\mathcal R$ denote the set of all labels and meta-labels, that is, all nodes in the tree apart from the root.
-Also, let $\mathcal L$ denote the set of labels, that is, the leaves of the tree.
-Let $T_i$ denote the subtree which is rooted at the $i$-th node.
-We introduce an indicator variable $z_{ai} \in \{0, 1\}$, where
-\begin{align}
-    \enskip  z_{ai} =  \begin{cases} 
-        y_{ai} \quad &\text{if} \enskip i \in {\mathcal L} \\
-        y_{a}(T_i) \quad &\text{if} \enskip i \in {\mathcal R - \mathcal L} \\
-    \end{cases}
-\end{align}
-
-We need to extend the definition of unary potentials to the expanded label space as follows:
-\begin{align}
-    \text{where} \enskip    \phi'_{a}(i) =  \begin{cases} 
-        \phi_{a}(i) \quad &\text{if} \enskip i \in {\mathcal L} \\
-        0  \quad &\text{if} \enskip i \in {\mathcal R - L} \\
-    \end{cases}
-\end{align}
-%
-We can now rewrite problem \eqref{eq:t-lp} in terms of new indicator variables $z_{ai}$:
-%
-\begin{align}
-    \text{(T-LP-FULL)} \quad &\min \widetilde{E}({\bf z}) = \sum_{i \in \mathcal{R}} \sum_{a \in {\mathcal X}} \phi_a'(i) \cdot z_{ai} + \nonumber \\
-            &\sum_{i \in \mathcal{R}} \sum_{(a, b) \in {\mathcal N}} w_{ab} \cdot l_{T_i} \cdot |z_{ai} - z_{bi}| \nonumber \\
-    \text{such that} \quad {\bf z} \in \Delta'
-%\label{eq:t-lp-full}
-\end{align}
-   where $\Delta'$ is the convex hull of the vectors satisfying
-\begin{align}
-    \enskip &\sum_{i \in \mathcal{L}} z_{ai} = 1, \enskip z_{ai} \in \{0, 1\} \enskip \forall a \in {\mathcal X}, i \in \mathcal{L} \\
-    \text{and} \enskip &z_{ai} = \sum_{j \in L(T_i) } z_{aj}.\enskip \forall a \in {\mathcal X}, i \in \mathcal{R - L} %\label{eq:consistency_constraint} 
-\end{align}
-%
-Constraint (\ref{eq:consistency_constraint}) ensures consistency among labels and meta-labels, that is, if a label is assigned then all the meta-labels which lie on the path from the root to the label should be assigned as well.
-We are now going to identify a suitable set encoding and the worst-case optimal submodular extension using (T-LP-FULL).
 
@@
     & = - \max_{{\bf s} \in EP(F)} \min_{{\bf z} \in \Delta} \langle {\bf z}, {\bf s'} \rangle
-    %\label{eq:metric_maxmin}
+    \label{eq:metric_maxmin}
 \end{align}
@@
 \begin{proof}
-    We observe that $F_{r-\textrm{HST}}$ is of exactly the same form as $F_{Potts}$, except that the Ising models $F_i$ are defined over not just labels, but meta-labels as well.
-Using the same logic as in the proof of proposition \ref{proposition:a-potts_optimal}, each $F_i$ is the Lovasz extension of
+    We observe that $F_{r-\textrm{HST}}$ is of exactly the same form as
+    $F_{Potts}$, except that the Ising models $F_i$ are defined over not just labels, but meta-labels as well.
+Using the same logic as in the proof of Potts extension, each $F_i$ is the Lovasz extension of
         \begin{equation}

### 2018-12-05 tex/paper.tex
Experiments section for higher-order Potts model completed.

@@
 \end{figure}
+
 \begin{figure*}
@@
 run our algorithm till 100 iterations, since the visual quality of the solution
-does not show much improvement beyond this point. We run mean-field up to convergence, with a threshold of 0.001 for change in KL-divergence.
+does not show much improvement beyond this point. We run mean-field up to
+convergence, with a threshold of 0.001 for change in KL-divergence.
 
@@
 
+\begin{figure*}
+    \centering
+\begin{tabular}{cccccc}
+        \includegraphics[scale = 0.22]{\imagePath/stereo/higher_potts/tsukuba_GT} &
+        \includegraphics[scale = 0.22]{\imagePath/stereo/higher_potts/tsukuba_hop} &
+        \includegraphics[scale = 0.34]{\imagePath/stereo/higher_potts/baby1_GT} &
+        \includegraphics[scale = 0.16]{\imagePath/stereo/higher_potts/baby1_hop} &
+        \includegraphics[scale = 0.34]{\imagePath/stereo/higher_potts/flowerpots_GT} &
+        \includegraphics[scale = 0.16]{\imagePath/stereo/higher_potts/flowerpots_hop} \\
+        \scriptsize(a) Tsukuba GT & \scriptsize(b) Submod solution
+        & \scriptsize(a) Baby1 GT & \scriptsize(b) Submod solution
+        & \scriptsize(a) Flowerpots GT & \scriptsize(b) Submod solution\\
+\end{tabular}
+%\vspace{2mm}
+\mycaption{\footnotesize \em Stereo matching using dense CRFs with higher-order
+  Potts term in addition to Potts compatibility and Gaussian pairwise potentials.}
+\label{fig:stereo_hop}
+\end{figure*}
+
 \mysubsection{Stereo Matching using Higher-order Potts Model}
 \label{subsec:stereo_hop}
-\myparagraph{Data} Next, we use the higher-order Potts model for stereo
-matching on the Middlebury dataset. For this
-purpose, we take a dense CRF model with Potts compatibility term and Gaussian pairwise
-potentials, and augment it with higher-order Potts term.
+\myparagraph{Data} Next, we use the higher-order Potts model for stereo matching on the Middlebury dataset.
+Higher-order Potts model is suitable to be used for only some images in the dataset, since others have gradually sloping surfaces and higher-order cliques will force all pixels on these surfaces to take similar disparity values.
+We take a dense CRF model with Potts compatibility term and Gaussian pairwise potentials, and augment it with higher-order Potts term.
+We obtain higher-order cliques by using super-pixels obtained from mean-shift algorithm \citep{comaniciu2002mean}.
 
@@
 \myparagraph{Results} Figure \ref{fig:stereo_hop} shows some example solutions obtained by picking the
-label with maximum marginal probability for each variable.
+label with maximum marginal probability for each variable. We observe that our method gives results that are reasonably close to the ground-truth.
 

### 2018-12-05 tex/abstract.tex
Version sent to Pawan & Chris for review.

@@
     worst-case optimal submodular extension for the more general class of
-    metric labeling; and (iii) efficiently compute the marginals for the widely
-    used dense CRF model with the help of a recently proposed Gaussian filtering method.
+    metric labeling; (iii) efficiently compute the marginals for the widely
+    used dense CRF model with the help of a recently proposed Gaussian
+    filtering method; and (iv) identify the submodular extension based on an LP
+    relaxation for the higher-order Potts model.
 Using synthetic and real data, we show that our approach provides comparable upper bounds on the log-partition function to those obtained using tree-reweighted message passing (TRW) in cases where the latter is computationally feasible.

### 2018-12-05 tex/paper.tex
Version sent to Pawan & Chris for review.

@@
 inference algorithm for dense CRFs, which computes a lower bound on the
-log-partition ~\citep{koltun2011efficient}. We show that the quality of our
+log-partition ~\citep{koltun2011efficient}. Finally, we also obtain the
+submodular extension based on the LP relaxation for a higher-order version of
+the Potts model that captures useful image statistics.  We show that the quality of our
 solution is comparable to tree reweighted message passing (TRW)
@@
 
-\begin{figure}
-\centering
-\includegraphics[scale = 0.30]{./figures/potts_chain_extension.png}
-\mycaption{\footnotesize \em An illustration of the worst-case optimal
-submodular extension for Potts model for a chain graph of 4 variables, each of
-which can take 3 labels. The figure shows the way to compute the extension
-values of the set $A = \{v_{1, 1}, v_{4, 1}, v_{3, 2}$\}.}
-\label{fig:alternate}
-\end{figure}
-\begin{figure}
-\centering
-\resizebox{150pt}{120pt}{
-\begin{tikzpicture}[-latex ,auto, node distance =1.5 cm and 1.5cm ,on grid
-  , semithick , state/.style ={ circle , color = black , draw,black , text=blue , minimum width =1 cm}, cross/.style={color=red, very thick, dashed}, unary/.style={color=blue, very thick}]
-
-  \node[state] (n11) at (1, 1) {\Large $1\mhyphen1$};
-  \node[state] (n12) at (3, 1) {\Large $1\mhyphen2$};
-  \node[state] (n13) at (5, 1) {\Large $1\mhyphen3$};
-  \node[state] (n21) at (7, 1) {\Large $2\mhyphen1$};
-  \node[state] (n22) at (9, 1) {\Large $2\mhyphen2$};
-  \node[state] (n23) at (11, 1) {\Large $2\mhyphen3$};
-  \node[state] (source) at (6, 3) {\Large s};
-  \node[state] (sink) at (6, -3) {\Large t};
-
-% \begin{scope}[ every node/.style={fill=white,circle}]
-\path[unary] (n11) edge [bend right =25] node [pos=0.20, anchor=west] {\Large$\phi_1(1)$} (sink);
-\path[unary] (n12) edge [bend right =25] node [pos=0.20, anchor=west]{\Large$\phi_1(2)$} (sink);
-\path[unary] (n13) edge [bend right =25] node [pos=0.20, anchor=west]{\Large$\phi_1(3)$} (sink);
-\path[unary] (n21) edge [bend left =25] node [pos=0.20, anchor=west]{\Large$\phi_2(1)$} (sink);
-\path[unary] (n22) edge [bend left =25] node [pos=0.20, anchor=west]{\Large$\phi_2(2)$} (sink);
-\path[unary] (n23) edge [bend left =25] node [pos=0.20, anchor=west]{\Large$\phi_2(3)$} (sink);
-%\end{scope}
-
-\path[cross] (n11) edge [bend left =25] node[above] {} (n21);
-\path[cross] (n21) edge [bend left =25] node[below] {} (n11);
-
-\path[cross] (n12) edge [bend left =25] node[above] {} (n22);
-\path[cross] (n22) edge [bend left =25] node[below] {} (n12);
-
-\path[cross] (n13) edge [bend left =25] node[above] {} (n23);
-\path[cross] (n23) edge [bend left =25] node[below] {} (n13);
-\end{tikzpicture}
-}
-\mycaption{\footnotesize \em An st-graph specifying the worst-case optimal submodular
-extension for Potts model for 2 variables with 3 labels each and connected to
-each other. There is a node for each variable and each label, that is, for all
-elements of the ground set. The nodes have been labeled as `variable-label',
-hence node 1-1 represents the element $v_{11}$ and so on. The solid blue arcs
-model the unary potentials, and the dotted red arcs represent the pairwise
-potentials. Each dotted red arc has weight $w_{12}/2$.}
-\label{fig:optimal_extension_graph}
-\end{figure}
-
 \mysection{Potts Model}
@@
 \end{equation}
+
+\begin{figure}
+\centering
+\includegraphics[scale = 0.30]{./figures/potts_chain_extension.png}
+\mycaption{\footnotesize \em An illustration of the worst-case optimal
+submodular extension for Potts model for a chain graph of 4 variables, each of
+which can take 3 labels. The figure shows the way to compute the extension
+values of the set $A = \{v_{1, 1}, v_{4, 1}, v_{3, 2}$\}.}
+\label{fig:alternate}
+\end{figure}
+
+\begin{figure}
+\centering
+\resizebox{180pt}{150pt}{
+\begin{tikzpicture}[-latex ,auto, node distance =1.5 cm and 1.5cm ,on grid
+  , semithick , state/.style ={ circle , color = black , draw,black , text=blue , minimum width =1 cm}, cross/.style={color=red, very thick, dashed}, unary/.style={color=blue, very thick}]
+
+  \node[state] (n11) at (1, 1) {\Large $1\mhyphen1$};
+  \node[state] (n12) at (3, 1) {\Large $1\mhyphen2$};
+  \node[state] (n13) at (5, 1) {\Large $1\mhyphen3$};
+  \node[state] (n21) at (7, 1) {\Large $2\mhyphen1$};
+  \node[state] (n22) at (9, 1) {\Large $2\mhyphen2$};
+  \node[state] (n23) at (11, 1) {\Large $2\mhyphen3$};
+  \node[state] (source) at (6, 3) {\Large s};
+  \node[state] (sink) at (6, -3) {\Large t};
+
+% \begin{scope}[ every node/.style={fill=white,circle}]
+\path[unary] (n11) edge [bend right =25] node [pos=0.20, anchor=west] {\Large$\phi_1(1)$} (sink);
+\path[unary] (n12) edge [bend right =25] node [pos=0.20, anchor=west]{\Large$\phi_1(2)$} (sink);
+\path[unary] (n13) edge [bend right =25] node [pos=0.20, anchor=west]{\Large$\phi_1(3)$} (sink);
+\path[unary] (n21) edge [bend left =25] node [pos=0.20, anchor=west]{\Large$\phi_2(1)$} (sink);
+\path[unary] (n22) edge [bend left =25] node [pos=0.20, anchor=west]{\Large$\phi_2(2)$} (sink);
+\path[unary] (n23) edge [bend left =25] node [pos=0.20, anchor=west]{\Large$\phi_2(3)$} (sink);
+%\end{scope}
+
+\path[cross] (n11) edge [bend left =25] node[above] {} (n21);
+\path[cross] (n21) edge [bend left =25] node[below] {} (n11);
+
+\path[cross] (n12) edge [bend left =25] node[above] {} (n22);
+\path[cross] (n22) edge [bend left =25] node[below] {} (n12);
+
+\path[cross] (n13) edge [bend left =25] node[above] {} (n23);
+\path[cross] (n23) edge [bend left =25] node[below] {} (n13);
+\end{tikzpicture}
+}
+\mycaption{\footnotesize \em An st-graph specifying the worst-case optimal submodular
+extension for Potts model for 2 variables with 3 labels each and connected to
+each other. There is a node for each variable and each label, that is, for all
+elements of the ground set. The nodes have been labeled as `variable-label',
+hence node 1-1 represents the element $v_{11}$ and so on. The solid blue arcs
+model the unary potentials, and the dotted red arcs represent the pairwise
+potentials. Each dotted red arc has weight $w_{12}/2$.}
+\label{fig:optimal_extension_graph}
+\end{figure}
+
 \myparagraph{Worst-case Optimal Submodular Extension} 
@@
 
-\begin{figure}
-\centering
-\includegraphics[scale = 0.30]{./figures/rhst_notation.png}
-\mycaption{\footnotesize \em A hierarchical Potts model instance illustrating notations with 2 meta-labels (blue) and 4 labels (yellow).
-All labels are at the same level.
-$r = 2$, that is, edge-length decreases by 2 at each level.
-Also, distance between labels 1 and 3 is $d(1, 3) = 0.5 + 1 + 1 + 0.5 = 3$.}
-\label{fig:rhst}
-\end{figure}
 Potts model imposes the same penalty for unequal assignment of labels to neighbouring variables, regardless of the label dissimilarity.
@@
 %
+\begin{figure}
+\centering
+\includegraphics[scale = 0.30]{./figures/rhst_notation.png}
+\mycaption{\footnotesize \em A hierarchical Potts model instance illustrating notations with 2 meta-labels (blue) and 4 labels (yellow).
+All labels are at the same level.
+$r = 2$, that is, edge-length decreases by 2 at each level.
+Also, distance between labels 1 and 3 is $d(1, 3) = 0.5 + 1 + 1 + 0.5 = 3$.}
+\label{fig:rhst}
+\end{figure}
+
 A subtree $T$ of an hierarchical Potts model is a tree comprising all the descendants of some node $v$ (not root).
@@
 considering the Potts model as a star-shaped tree with edge weights as 0.5.
-\begin{figure}
-\centering
-\begin{tabular}{cc}
-\includegraphics[scale = 0.16]{./figures/synthetic_tree.png} &
-\includegraphics[scale = 0.16]{./figures/synthetic_tree_2_new.png} \\
-  \scriptsize(a) Tree A & \scriptsize(b) Tree B \\
-\end{tabular}
-\mycaption{\footnotesize \em The hierarchical Potts models defining pairwise
-distance among 20 labels used for upper-bound comparison with TRW. Blue nodes
-are the meta-labels and yellow nodes are labels. All the edges at a particular
-level have the same edge weights. The sequence of weights from root level to
-leaf level is 1, 0.5 for tree A and 1, 1, 0.5 for tree B. The yellow node is
-shown to clump together 4 and 5 leaf nodes for tree A and B respectively.} 
-\label{fig:syn_rhst}
-\end{figure}
 
@@
 
+A similar observation can be made in case of hierarchical Potts model.
+Hence we have the first practical algorithm to compute upper bound of log-partition function of a dense CRF for Potts and metric energies.
+
+\mysection{Experiments}
+\label{sec:exp}
+Using synthetic data, we show that our upper-bound compares favorably with TRW
+for both Potts and hierarchical Potts models. For comparison, we restrict
+ourselves to sparse CRFs, as the code available for TRW does not scale well to
+dense CRFs. We also perform stereo matching using dense CRF models and compare
+our results with the mean-field-based approach of \citep{koltun2011efficient}.
+All experiments were run on a x86-64, 3.8GHz machine with 16GB RAM. In this
+section, we refer to our algorithm as \emph{Submod} and mean field as \emph{MF}.
+
+%\begin{figure}
+%       \Tree [.root [.{m0} 0-3 ] [.{m1} 4-7 ] [.{m2} 8-11 ] [.{m3} 12-15 ] [.{m4} 16-19 ]]
+%   %    \Tree [.root [.{m0} 0 1 2 3 ] [.{m1} 4 5 6 7 ] [.{m2} 8 9 10 11 ] [.{m3} 12 13 14 15 ] [.{m4} 16 17 18 19 ]]
+%%    \Tree[. [. [.1 .2 .3]]]
+%\mycaption{\footnotesize \em An $r$-HST defining pairwise distance among 20 labels used for upper-bound comparison with TRW. $m0$ stands for meta-label 0, and so on.}
+%\label{fig:syn_rhst}
+%\end{figure}
 \begin{figure}
 \centering
-\resizebox{150pt}{120pt}{
+\begin{tabular}{cc}
+\includegraphics[scale = 0.25]{./figures/synthetic_tree.png} &
+\includegraphics[scale = 0.25]{./figures/synthetic_tree_2_new.png} \\
+  \scriptsize(a) Tree A & \scriptsize(b) Tree B \\
+\end{tabular}
+\mycaption{\footnotesize \em The hierarchical Potts models defining pairwise
+distance among 20 labels used for upper-bound comparison with TRW. Blue nodes
+are the meta-labels and yellow nodes are labels. All the edges at a particular
+level have the same edge weights. The sequence of weights from root level to
+leaf level is 1, 0.5 for tree A and 1, 1, 0.5 for tree B. The yellow node is
+shown to clump together 4 and 5 leaf nodes for tree A and B respectively.} 
+\label{fig:syn_rhst}
+\end{figure}
+
+\begin{figure}
+\centering
+\resizebox{180pt}{150pt}{
 \begin{tikzpicture}[-latex ,auto ,node distance =2 cm and 2cm ,on grid
@@
 
-A similar observation can be made in case of hierarchical Potts model.
-Hence we have the first practical algorithm to compute upper bound of log-partition function of a dense CRF for Potts and metric energies.
-
-\mysection{Experiments}
-\label{sec:exp}
-Using synthetic data, we show that our upper-bound compares favorably with TRW
-for both Potts and hierarchical Potts models. For comparison, we restrict
-ourselves to sparse CRFs, as the code available for TRW does not scale well to
-dense CRFs. We also perform stereo matching using dense CRF models and compare
-our results with the mean-field-based approach of \citep{koltun2011efficient}.
-All experiments were run on a x86-64, 3.8GHz machine with 16GB RAM. In this
-section, we refer to our algorithm as \emph{Submod} and mean field as \emph{MF}.
-
-%\begin{figure}
-%       \Tree [.root [.{m0} 0-3 ] [.{m1} 4-7 ] [.{m2} 8-11 ] [.{m3} 12-15 ] [.{m4} 16-19 ]]
-%   %    \Tree [.root [.{m0} 0 1 2 3 ] [.{m1} 4 5 6 7 ] [.{m2} 8 9 10 11 ] [.{m3} 12 13 14 15 ] [.{m4} 16 17 18 19 ]]
-%%    \Tree[. [. [.1 .2 .3]]]
-%\mycaption{\footnotesize \em An $r$-HST defining pairwise distance among 20 labels used for upper-bound comparison with TRW. $m0$ stands for meta-label 0, and so on.}
-%\label{fig:syn_rhst}
-%\end{figure}
 \begin{figure}
@@
 \end{figure}
-
-\begin{figure*}
-    \centering
-\begin{tabular}{cccccc}
-        \includegraphics[scale = 0.14]{\imagePath/stereo/venus_GT} &
-        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/venus_left_mf} &
-        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/venus_left_submod} &
-    %
-        \includegraphics[scale = 0.19]{\imagePath/stereo/tsukuba_GT} &
-        \includegraphics[scale = 0.19]{\imagePath/stereo/potts/tsukuba_left_mf} &
-        \includegraphics[scale = 0.19]{\imagePath/stereo/potts/tsukuba_left_submod} \\
-        \scriptsize(a) Venus GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution & \scriptsize(a) Tsukuba GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
-        {} & \scriptsize  60.32s, 1.83e+07 &  \scriptsize 469.75s, {\bf 1.55e+07} & {} & \scriptsize  14.93s, 8.21e+06 & \scriptsize 215.22s, {\bf 4.12e+06} \\
-         \includegraphics[scale = 0.14]{\imagePath/stereo/cones_GT} &
-        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/cones_left_mf} &
-        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/cones_left_submod} &
-        %
-       %
-         \includegraphics[scale = 0.15]{\imagePath/stereo/teddy_GT} &
-        \includegraphics[scale = 0.15]{\imagePath/stereo/potts/teddy_left_mf} &
-        \includegraphics[scale = 0.15]{\imagePath/stereo/potts/teddy_left_submod} \\
-    \scriptsize(a) Cones GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution & \scriptsize(a) Teddy GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
-        {} & \scriptsize 239.14s, 2.68e+07 &  \scriptsize 1082.72s, {\bf 1.27e+07} & {} & \scriptsize 555.30s, 2.36e+07 & \scriptsize 1257.86s, {\bf 1.58e+07}\\
-\end{tabular}
-%\vspace{2mm}
-\mycaption{\footnotesize \em Stereo matching using dense CRFs with Potts
-compatibility and Gaussian pairwise potentials. We compare our solution with
-the mean-field algorithm of Koltun and Krahenbuhl \citep{koltun2011efficient}.
-We observe that our method gives better-looking solutions with lower energy value at the cost of higher computational time.}
-\label{fig:stereo}
-\end{figure*}
 
@@
 
+\begin{figure*}
+    \centering
+\begin{tabular}{cccccc}
+        \includegraphics[scale = 0.14]{\imagePath/stereo/venus_GT} &
+        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/venus_left_mf} &
+        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/venus_left_submod} &
+    %
+        \includegraphics[scale = 0.19]{\imagePath/stereo/tsukuba_GT} &
+        \includegraphics[scale = 0.19]{\imagePath/stereo/potts/tsukuba_left_mf} &
+        \includegraphics[scale = 0.19]{\imagePath/stereo/potts/tsukuba_left_submod} \\
+        \scriptsize(a) Venus GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution & \scriptsize(a) Tsukuba GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
+        {} & \scriptsize  60.32s, 1.83e+07 &  \scriptsize 469.75s, {\bf 1.55e+07} & {} & \scriptsize  14.93s, 8.21e+06 & \scriptsize 215.22s, {\bf 4.12e+06} \\
+         \includegraphics[scale = 0.14]{\imagePath/stereo/cones_GT} &
+        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/cones_left_mf} &
+        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/cones_left_submod} &
+        %
+       %
+         \includegraphics[scale = 0.15]{\imagePath/stereo/teddy_GT} &
+        \includegraphics[scale = 0.15]{\imagePath/stereo/potts/teddy_left_mf} &
+        \includegraphics[scale = 0.15]{\imagePath/stereo/potts/teddy_left_submod} \\
+    \scriptsize(a) Cones GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution & \scriptsize(a) Teddy GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
+        {} & \scriptsize 239.14s, 2.68e+07 &  \scriptsize 1082.72s, {\bf 1.27e+07} & {} & \scriptsize 555.30s, 2.36e+07 & \scriptsize 1257.86s, {\bf 1.58e+07}\\
+\end{tabular}
+%\vspace{2mm}
+\mycaption{\footnotesize \em Stereo matching using dense CRFs with Potts
+compatibility and Gaussian pairwise potentials. We compare our solution with
+the mean-field algorithm of Koltun and Krahenbuhl \citep{koltun2011efficient}.
+We observe that our method gives better-looking solutions with lower energy value at the cost of higher computational time.}
+\label{fig:stereo}
+\end{figure*}
+
 \mysubsection{Stereo Matching using Dense CRFs}
@@
 We observe that our algorithm results in more natural looking stereo matching results with lower energy values for all images.
-However, mean-field runs faster than our method for each instance.
+However, mean-field runs faster than our method for each instance.The set of hyperparameters that we used can be found in the
+appendix.
 
@@
         & \scriptsize(a) Flowerpots GT & \scriptsize(b) Submod solution\\
+        {} & {} 
+        & {} & \scriptsize 2476.82s
+        & {} & \scriptsize 2471.58s\\
+
 \end{tabular}
@@
 amount of time. In this case, computation of conditional gradient takes
-$\mathcal{O}(cNL)$ where $c$ is the number of hihger-order cliques.
+$\mathcal{O}(cNL)$ where $c$ is the number of higher-order cliques. We ran our
+algorithm for 200 iterations.
 
-\myparagraph{Results} Figure \ref{fig:stereo_hop} shows some example solutions obtained by picking the
-label with maximum marginal probability for each variable. We observe that our method gives results that are reasonably close to the ground-truth.
+\myparagraph{Results} Figure \ref{fig:stereo_hop} shows some example solutions
+obtained by picking the label with maximum marginal probability for each
+variable. We observe that our method gives results that are reasonably close to
+the ground-truth. The set of hyperparameters that we used can be found in the
+appendix.
 

### 2018-12-12 tex/abstract.tex
Abstract modified@

--- 2018-11-30 tex/abstract.tex
@@
     the marginals depends crucially on the quality of the submodular extension.
-    To identify the best possible extension, we show an equivalence between the
+    To identify accurate extensions for different classes of energy functions, we show a relation between the
     submodular extensions of the energy and the objective functions of linear
-    programming (LP) relaxations for the corresponding MAP estimation problem.
+    programming (LP) relaxations for the corresponding MAP estimation problem via the Lovasz extension. 
     This allows us to (i) establish the worst-case optimality of the submodular
@@
     worst-case optimal submodular extension for the more general class of
-    metric labeling; and (iii) efficiently compute the marginals for the widely
-    used dense CRF model with the help of a recently proposed Gaussian filtering method.
+    metric labeling; (iii) efficiently compute the marginals for the widely
+    used dense CRF model with the help of a recently proposed Gaussian
+    filtering method; and (iv) propose a good submodular extension based on an accurate LP
+    relaxation for the higher-order Potts model.
 Using synthetic and real data, we show that our approach provides comparable upper bounds on the log-partition function to those obtained using tree-reweighted message passing (TRW) in cases where the latter is computationally feasible.
-Importantly, unlike TRW, our approach provides the first practical algorithm to compute an upper bound on the dense CRF model.
+Importantly, unlike TRW, our approach provides the first practical algorithm to compute an upper bound on the dense CRF model with higher-order Potts potentials.
 \end{abstract}

### 2018-12-12 tex/paper.tex
Abstract modified@

@@
 
-\begin{figure*}
-    \centering
-\begin{tabular}{cccccc}
-        \includegraphics[scale = 0.14]{\imagePath/stereo/venus_GT} &
-        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/venus_left_mf} &
-        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/venus_left_submod} &
-    %
-        \includegraphics[scale = 0.19]{\imagePath/stereo/tsukuba_GT} &
-        \includegraphics[scale = 0.19]{\imagePath/stereo/potts/tsukuba_left_mf} &
-        \includegraphics[scale = 0.19]{\imagePath/stereo/potts/tsukuba_left_submod} \\
-        \scriptsize(a) Venus GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution & \scriptsize(a) Tsukuba GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
-        {} & \scriptsize  60.32s, 1.83e+07 &  \scriptsize 469.75s, {\bf 1.55e+07} & {} & \scriptsize  14.93s, 8.21e+06 & \scriptsize 215.22s, {\bf 4.12e+06} \\
-         \includegraphics[scale = 0.14]{\imagePath/stereo/cones_GT} &
-        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/cones_left_mf} &
-        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/cones_left_submod} &
-        %
-       %
-         \includegraphics[scale = 0.15]{\imagePath/stereo/teddy_GT} &
-        \includegraphics[scale = 0.15]{\imagePath/stereo/potts/teddy_left_mf} &
-        \includegraphics[scale = 0.15]{\imagePath/stereo/potts/teddy_left_submod} \\
-    \scriptsize(a) Cones GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution & \scriptsize(a) Teddy GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
-        {} & \scriptsize 239.14s, 2.68e+07 &  \scriptsize 1082.72s, {\bf 1.27e+07} & {} & \scriptsize 555.30s, 2.36e+07 & \scriptsize 1257.86s, {\bf 1.58e+07}\\
-\end{tabular}
-%\vspace{2mm}
-\mycaption{\footnotesize \em Stereo matching using dense CRFs with Potts
-compatibility and Gaussian pairwise potentials. We compare our solution with
-the mean-field algorithm of Koltun and Krahenbuhl \citep{koltun2011efficient}.
+%\begin{figure*}
+%    \centering
+%\begin{tabular}{cccccc}
+%        \includegraphics[scale = 0.14]{\imagePath/stereo/venus_GT} &
+%        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/venus_left_mf} &
+%        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/venus_left_submod} &
+%    %
+%        \includegraphics[scale = 0.19]{\imagePath/stereo/tsukuba_GT} &
+%        \includegraphics[scale = 0.19]{\imagePath/stereo/potts/tsukuba_left_mf} &
+%        \includegraphics[scale = 0.19]{\imagePath/stereo/potts/tsukuba_left_submod} \\
+%        \scriptsize(a) Venus GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution & \scriptsize(a) Tsukuba GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
+%        {} & \scriptsize  60.32s, 1.83e+07 &  \scriptsize 469.75s, {\bf 1.55e+07} & {} & \scriptsize  14.93s, 8.21e+06 & \scriptsize 215.22s, {\bf 4.12e+06} \\
+%         \includegraphics[scale = 0.14]{\imagePath/stereo/cones_GT} &
+%        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/cones_left_mf} &
+%        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/cones_left_submod} &
+%        %
+%       %
+%         \includegraphics[scale = 0.15]{\imagePath/stereo/teddy_GT} &
+%        \includegraphics[scale = 0.15]{\imagePath/stereo/potts/teddy_left_mf} &
+%        \includegraphics[scale = 0.15]{\imagePath/stereo/potts/teddy_left_submod} \\
+%    \scriptsize(a) Cones GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution & \scriptsize(a) Teddy GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
+%        {} & \scriptsize 239.14s, 2.68e+07 &  \scriptsize 1082.72s, {\bf 1.27e+07} & {} & \scriptsize 555.30s, 2.36e+07 & \scriptsize 1257.86s, {\bf 1.58e+07}\\
+%\end{tabular}
+%%\vspace{2mm}
+%\mycaption{\footnotesize \em Stereo matching using dense CRFs with Potts
+%compatibility and Gaussian pairwise potentials. We compare our solution with
+%the mean-field algorithm of Koltun and Krahenbuhl \citep{koltun2011efficient}.
 We observe that our method gives better-looking solutions with lower energy value at the cost of higher computational time.}
-\label{fig:stereo}
-\end{figure*}
+%\label{fig:stereo}
+%\end{figure*}
 
@@
 
-\begin{figure*}
-    \centering
-\begin{tabular}{cccccc}
-        \includegraphics[scale = 0.22]{\imagePath/stereo/higher_potts/tsukuba_GT} &
-        \includegraphics[scale = 0.22]{\imagePath/stereo/higher_potts/tsukuba_hop} &
-        \includegraphics[scale = 0.34]{\imagePath/stereo/higher_potts/baby1_GT} &
-        \includegraphics[scale = 0.16]{\imagePath/stereo/higher_potts/baby1_hop} &
-        \includegraphics[scale = 0.34]{\imagePath/stereo/higher_potts/flowerpots_GT} &
-        \includegraphics[scale = 0.16]{\imagePath/stereo/higher_potts/flowerpots_hop} \\
-        \scriptsize(a) Tsukuba GT & \scriptsize(b) Submod solution
-        & \scriptsize(a) Baby1 GT & \scriptsize(b) Submod solution
-        & \scriptsize(a) Flowerpots GT & \scriptsize(b) Submod solution\\
-        {} & {} 
-        & {} & \scriptsize 2476.82s
-        & {} & \scriptsize 2471.58s\\
-
-\end{tabular}
-%\vspace{2mm}
-\mycaption{\footnotesize \em Stereo matching using dense CRFs with higher-order
-  Potts term in addition to Potts compatibility and Gaussian pairwise potentials.}
-\label{fig:stereo_hop}
-\end{figure*}
+%\begin{figure*}
+%    \centering
+%\begin{tabular}{cccccc}
+%        \includegraphics[scale = 0.22]{\imagePath/stereo/higher_potts/tsukuba_GT} &
+%        \includegraphics[scale = 0.22]{\imagePath/stereo/higher_potts/tsukuba_hop} &
+%        \includegraphics[scale = 0.34]{\imagePath/stereo/higher_potts/baby1_GT} &
+%        \includegraphics[scale = 0.16]{\imagePath/stereo/higher_potts/baby1_hop} &
+%        \includegraphics[scale = 0.34]{\imagePath/stereo/higher_potts/flowerpots_GT} &
+%        \includegraphics[scale = 0.16]{\imagePath/stereo/higher_potts/flowerpots_hop} \\
+%        \scriptsize(a) Tsukuba GT & \scriptsize(b) Submod solution
+%        & \scriptsize(a) Baby1 GT & \scriptsize(b) Submod solution
+%        & \scriptsize(a) Flowerpots GT & \scriptsize(b) Submod solution\\
+%        {} & {} 
+%        & {} & \scriptsize 2476.82s
+%        & {} & \scriptsize 2471.58s\\
+%
+%\end{tabular}
+%%\vspace{2mm}
+%\mycaption{\footnotesize \em Stereo matching using dense CRFs with higher-order
+%  Potts term in addition to Potts compatibility and Gaussian pairwise potentials.}
+%\label{fig:stereo_hop}
+%\end{figure*}
+%
 

### 2018-12-13 tex/paper.tex
Modified introduction@

@@
 The accuracy of the variational inference algorithm depends crucially on the choice of the submodular extension.
-Yet, previous work has largely ignored the question of identifying the best extension.
+Yet, previous work has largely ignored the question of identifying accurate extensions for different energy classes.
 Indeed, the difficulty of identifying submodular extensions of general energy functions could be a major reason why the experiments of~\citep{zhang2015higher} were restricted to the special case of models specified by the Potts energy functions.
@@
 objective function of an accurate linear programming (LP) relaxation of the
-corresponding MAP estimation problem~\citep{kleinberg2002approximation}. This
-connection has three important practical consequences. First, it establishes
+corresponding MAP estimation problem~\citep{kleinberg2002approximation}.
+Specifically, the Lovasz extension of a submodular extension can be shown to be an objective function for the LP relaxation.
+This
+connection has four important practical consequences. First, it establishes
 the accuracy of the submodular extension of the Potts model, via the
@@
 inference algorithm for dense CRFs, which computes a lower bound on the
-log-partition ~\citep{koltun2011efficient}. Finally, we also obtain the
-submodular extension based on the LP relaxation for a higher-order version of
-the Potts model that captures useful image statistics.  We show that the quality of our
-solution is comparable to tree reweighted message passing (TRW)
+log-partition ~\citep{koltun2011efficient}. Fourthly, we obtain a good 
+submodular extension for higher-order Potts model based on an accurate LP relaxation.
+Higher-order Potts model is a generalisation of the pairwise Potts model, and helps to capture interesting properties of the image that cannot be done using pairwise models.
+We show that our upper-bounds on synthetic problems are comparable to those from tree reweighted message passing (TRW)
 ~\citep{wainwright2005new} for the case of sparse CRFs. Unlike our approach,
 TRW is computationally infeasible for dense CRFs, thereby limiting its use in
-practice. Using dense CRF models, we perform stereo matching on standard
-data sets and obtain better results than \citep{koltun2011efficient}. The complete code is available at \url{https://github.com/pankajpansari/denseCRF}.
+practice. Using pairwise dense CRF models, we perform stereo matching on standard
+data sets and obtain better results than \citep{koltun2011efficient}. We also perform semantic segmentation on the MSRC-21 dataset
+
+The complete code is available at \url{https://github.com/pankajpansari/denseCRF}.
 

### 2018-12-13 tex/paper.tex
Preliminaries section modified.

--- 2018-12-05 tex/paper.tex
@@
 inference algorithm for dense CRFs, which computes a lower bound on the
-log-partition ~\citep{koltun2011efficient}. Finally, we also obtain the
-submodular extension based on the LP relaxation for a higher-order version of
-the Potts model that captures useful image statistics.  We show that the quality of our
-solution is comparable to tree reweighted message passing (TRW)
+log-partition ~\citep{koltun2011efficient}. Fourthly, we obtain a good 
+submodular extension for higher-order Potts model based on an accurate LP
+relaxation. Higher-order Potts model is a generalisation of the pairwise Potts
+model, and helps to capture interesting properties of the image that cannot be
+done using pairwise models~\citep{vineet2014filter, kohli2009robust}. We show that our upper-bounds on synthetic problems are comparable to those from tree reweighted message passing (TRW)
 ~\citep{wainwright2005new} for the case of sparse CRFs. Unlike our approach,
 TRW is computationally infeasible for dense CRFs, thereby limiting its use in
-practice. Using dense CRF models, we perform stereo matching on standard
-data sets and obtain better results than \citep{koltun2011efficient}. The complete code is available at \url{https://github.com/pankajpansari/denseCRF}.
+practice. Using pairwise dense CRF models, we perform stereo matching on standard
+data sets and obtain better results than \citep{koltun2011efficient}. We also perform semantic segmentation on the MSRC-21 dataset
+
+The complete code is available at \url{https://github.com/pankajpansari/denseCRF}.
 
@@
 \end{align}
-where $\bf s$ denotes the modular function $s(.)$ considered as a vector.
+where $\bf s$ denotes the modular function $s(.)$ represented as a vector.
 
@@
 
+We can augment the above pairwise models with higher-order potentials.
+Let $\mathcal C$ be the set of cliques, which can, for instance, be the set of superpixels from an image segmentation method.
+Also, let $\bf x_c$ be the component of $\bf x$ formed by elements in clique $c$.
+The energy function now also contains higher-order potentials
+%
+\begin{equation}
+    E({\bf x}) = \sum_{a \in {\mathcal X}} \phi_a(x_a) + \sum_{(a, b) \in {\mathcal N}} \phi_{ab}(x_a, x_b) + \sum_{c \in {\mathcal C}} \phi_c({\bf x_c}),
+    \label{eq:energy_hop_min}
+  \end{equation}
+%
+
 The energy function also defines a probability distribution $P({\bf x})$ as follows:
@@
 a ground set such that some of its subsets correspond to valid CRF labelings.
-To such an extension, we need an encoding scheme which gives the sets corresponding to valid CRF labelings.
+To obtain such an extension, we need an encoding scheme which gives the sets corresponding to valid CRF labelings.
 

### 2018-12-13 tex/paper.tex
The section connecting lp relaxations and submodular extensions modified

@@
 
-\mysection{Worst-case Optimal Submodular Extensions via LP Relaxations}
+\mysection{Accurate Submodular Extensions from LP Relaxations}
 
-\myparagraph{Worst-case Optimal Submodular Extensions} Different choices of extensions $F$ change the domain in problem \eqref{eq:min_upper_bound}, leading to different upper bounds on the log-partition function.
+\myparagraph{Relation between Marginal Estimation and LP Relaxation}
+Different choices of extensions $F$ change the domain in problem \eqref{eq:min_upper_bound}, leading to different upper bounds on the log-partition function.
 How does one come up with an extension which yields the tightest bound?
 
+If we introduce a temperature parameter in $P({\bf x})$ (equation \eqref{eq:crf_prob_distrib}) by using $E({\bf x})/T$ and decrease $T$, the resulting distribution starts to peak more sharply around its mode.
+As $T \to 0$, marginal estimation becomes the same as MAP inference since the resulting distribution $P^0({\bf x})$ has mass 1 at its mode ${\bf x}^*$ and is 0 everywhere else.
+Given the MAP solution ${\bf x}^*$, one can compute the marginals as $P(X_i = j) = [x_i^* = j]$, where [.] is the Iverson bracket.
+
+  Motivated by this connection, we ask if one can introduce a temperature parameter to our problem \eqref{eq:min_upper_bound} and transform it to an LP relaxation in the limit $T \to 0$?
+We can then hope to use accurate LP relaxations of MAP problems known in literature to find good submodular extensionsfor different classes of energy functions.
+
+We answer this question in affirmative.
+Specifically, in the following two sections we show how one can select the set encoding and submodular extension to convert problem \eqref{eq:min_upper_bound} to accurate LP relaxations for Potts, hierarchical Potts and higher-order Potts models.
+In fact, when the LP relaxation has tightness guarantees, we obtain worst-case optimal submodular extensions - a notion we elucidate shortly.
+
+\myparagraph{Worst-case Optimal Submodular Extensions}
 In this paper, we focus on submodular extension families $\mathcal F(.)$ which
 for each instance of the energy function $E(.)$ belonging to a given class $\mathcal E$ gives a corresponding
-submodular extension ${\mathcal F}(E)$. We find the extension family ${\mathcal
-F}_{opt}$ that is \emph{worst-case optimal}. This implies that there does not
+submodular extension ${\mathcal F}(E)$. Our extensions are derived from LP relaxations. When the LP relaxation is the tightest known, the extension family ${\mathcal
+F}_{opt}$ we obtain is \emph{worst-case optimal}. This implies that there does not
 exist another submodular extension family ${\mathcal F}$ that gives a tighter
@@
 \end{equation}
-Note that our problem is different from taking a given energy model and
-obtaining a submodular extension which is optimal for that model. Also, we seek
-a closed-form analytical expression for $\mathcal F$. For the sake of clarity,
+Note that the notion of worst-case optimality does not guarantee that the extension is best for every instance of the energy function. 
+Also, we seek a closed-form analytical expression for $\mathcal F$. For the sake of clarity,
 in the analysis that follows we use $F$ to represent ${\mathcal F}(E)$ where
-the meaning is clear from context. The two classes of energy functions we
-consider in this paper are Potts and hierarchical Potts families.
-%
-\myparagraph{Using LP Relaxations} If we introduce a temperature parameter in $P({\bf x})$ (equation \eqref{eq:crf_prob_distrib}) by using $E({\bf x})/T$ and decrease $T$, the resulting distribution starts to peak more sharply around its mode.
-As $T \to 0$, marginal estimation becomes the same as MAP inference since the resulting distribution $P^0({\bf x})$ has mass 1 at its mode ${\bf x}^*$ and is 0 everywhere else.
-Given the MAP solution ${\bf x}^*$, one can compute the marginals as $P(X_i = j) = [x_i^* = j]$, where [.] is the Iverson bracket.
-
-Motivated by this connection, we ask if one can introduce a temperature parameter to our problem \eqref{eq:min_upper_bound} and transform it to an LP relaxation in the limit $T \to 0$?
-We can then hope to use the tightest LP relaxations of MAP problems known in literature to find worst-case optimal submodular extensions.
-We answer this question in affirmative.
-Specifically, in the following two sections we show how one can select the set encoding and submodular extension to convert problem \eqref{eq:min_upper_bound} to the tightest known LP relaxations for Potts and hierarchical Potts models.
-Importantly, we prove the worst-case optimality of the extensions thus obtained.
+the meaning is clear from context.
 

### 2018-12-13 tex/paper.tex
Rearranged some sections and subections

@@
 
-\myparagraph{Relation between Marginal Estimation and LP Relaxation}
 Different choices of extensions $F$ change the domain in problem \eqref{eq:min_upper_bound}, leading to different upper bounds on the log-partition function.
@@
 
-\myparagraph{Worst-case Optimal Submodular Extensions}
+\mysection{Worst-case Optimal Submodular Extensions}
+
+\myparagraph{Worst-case Optimality}
 In this paper, we focus on submodular extension families $\mathcal F(.)$ which
@@
 
-\mysection{Potts Model}
-
+\mysubsection{Potts Model}
 The Potts model, also known as the uniform metric, specifies the pairwise potentials $\phi_{ab}(x_a, x_b)$ in equation (\ref{eq:energy_min}) as follows:
@@
 
-\mysection{Hierarchical Potts Model}
+\mysubsection{Hierarchical Potts Model}
 \label{sec:rhst_worst-case optimal}
@@
 
-\mysection{Higher-order Potts Model} The pairwise Potts model often fails
+\mysubsection{Fast Conditional Gradient Computation for Dense CRFs}
+\label{sec:dense_crf}
+
+\myparagraph{Dense CRF Energy Function} A dense CRF is specified by the following energy function 
+\begin{equation}
+    E({\bf x}) = \sum_{a \in {\mathcal X}} \phi_a(x_a) + \sum_{a \in {\mathcal X}} \sum_{b \in {\mathcal X}, b \neq a} \phi_{ab}(x_a, x_b).
+\end{equation}
+Note that every random variable is a neighbour of every other random variable in a dense CRF.
+Similar to previous work \citep{koltun2011efficient}, we consider the pairwise potentials to be to be Gaussian, that is,
+\begin{align}
+    &\phi(i, j) = \mu(i, j) \sum_{m} w^{(m)} k({\bf f}_a^{(m)}, {\bf f}_b^{(m)}), \\
+    &k({\bf f}_a^{(m)}, {\bf f}_b^{(m)}) = \exp\left( \frac{- ||{\bf f}_a - {\bf f}_b ||^2}{2} \right).
+\end{align}
+The term $\mu(i,j)$ is known as {\it label compatibility} function between
+labels $i$ and $j$. Potts model and hierarchical Potts models are examples of
+$\mu(i, j)$. The other term is a mixture of Gaussian kernels $k(.,.)$ and is
+called the {\it pixel compatibility} function. The terms ${\bf f}_a^{(m)}$ are
+features that describe the random variable $X_a$. In practice, similar to
+\citep{koltun2011efficient}, we use $x, y$ coordinates and RGB values
+associated to a pixel as its features.
+
+Algorithm \ref{algo:fw_inference} assumes that the conditional gradient $\bf s^*$ in step {\ref{line:cond_grad}} can be computed efficiently.
+This is certainly not the case for dense CRFs, since computing $s^*$ involves $NL$ function evaluations of the submodular extension $F$, where $N$ is the number of variables, and $L$ is the number of labels.
+Each $F$ evaluation has complexity ${\mathcal O}(N)$ using the efficient Gaussian filtering algorithm of \citep{koltun2011efficient}.
+However, computation of $s^*$ would still be ${\mathcal O} (N^2)$ this way, which is clearly impractical for computer-vision applications where $N \sim 10^5 - 10^6$.
+
+However, using the equivalence of relaxed LP objectives and the Lovasz
+extension of submodular extensions in proposition
+\ref{proposition:potts_equiv}, we are able to compute $s^*$ in ${\mathcal
+O}(NL)$ time. Specifically, we use the algorithm of
+Ajanthan et al. \citep{ajanthan2017efficient}, which provides an efficient filtering procedure to compute the subgradient of the LP relaxation objective $E({\bf y})$ of (P-LP).\\
+%
+\vspace{-0.5cm}
+{\proposition Computing the subgradient of $E({\bf y})$ in (P-LP) is equivalent to computing the conditional gradient for the submodular function $F_{Potts}$. \label{proposition:subgrad}} 
+\newline(Proof in appendix).
+\begin{proof}
+  For the Potts model, we derived the worst-case optimal extension $F_{Potts}$ by
+  making its Lovasz extension $f({\bf y})$ same as the objective function
+    $E({\bf y})$ of the worst-case optimal LP relaxation. Hence, we have
+\begin{align}
+    E({\bf y}) &= f({\bf y})  \nonumber \\
+    {} &=  \max_{{\bf s} \in EP(F)} \langle {\bf y}, {\bf s} \rangle.\nonumber
+\end{align}
+
+The subdifferential of $E({\bf y})$ at $\bf y_0$ is the `active' linear function at $\bf y_0$. Hence,
+\begin{align}
+    \partial E({\bf y})|_{\bf y = y_0} &=  \argmax_{{\bf s} \in EP(F)} \langle {\bf y_0}, {\bf s} \rangle
+    \label{eq:subgrad}
+\end{align}
+Equation (\ref{eq:subgrad}) is equivalent to ${\bf s}^*$ computation in line \ref{line:cond_grad} of algorithm \ref{algo:fw_inference}, with ${\bf y_0} = - \nabla g({\bf s}_k)$.
+\end{proof}
+
+A similar observation can be made in case of hierarchical Potts model.
+Hence we have the first practical algorithm to compute upper bound of log-partition function of a dense CRF for Potts and metric energies.
+
+\mysection{Accurate Submodular Extension for Higher-order Potts Model} The pairwise Potts model often fails
 to capture useful image statistics because it restricts the order of the
@@
 \end{proof}
-
-\mysection{Fast Conditional Gradient Computation for Dense CRFs}
-\label{sec:dense_crf}
-
-\myparagraph{Dense CRF Energy Function} A dense CRF is specified by the following energy function 
-\begin{equation}
-    E({\bf x}) = \sum_{a \in {\mathcal X}} \phi_a(x_a) + \sum_{a \in {\mathcal X}} \sum_{b \in {\mathcal X}, b \neq a} \phi_{ab}(x_a, x_b).
-\end{equation}
-Note that every random variable is a neighbour of every other random variable in a dense CRF.
-Similar to previous work \citep{koltun2011efficient}, we consider the pairwise potentials to be to be Gaussian, that is,
-\begin{align}
-    &\phi(i, j) = \mu(i, j) \sum_{m} w^{(m)} k({\bf f}_a^{(m)}, {\bf f}_b^{(m)}), \\
-    &k({\bf f}_a^{(m)}, {\bf f}_b^{(m)}) = \exp\left( \frac{- ||{\bf f}_a - {\bf f}_b ||^2}{2} \right).
-\end{align}
-The term $\mu(i,j)$ is known as {\it label compatibility} function between
-labels $i$ and $j$. Potts model and hierarchical Potts models are examples of
-$\mu(i, j)$. The other term is a mixture of Gaussian kernels $k(.,.)$ and is
-called the {\it pixel compatibility} function. The terms ${\bf f}_a^{(m)}$ are
-features that describe the random variable $X_a$. In practice, similar to
-\citep{koltun2011efficient}, we use $x, y$ coordinates and RGB values
-associated to a pixel as its features.
-
-Algorithm \ref{algo:fw_inference} assumes that the conditional gradient $\bf s^*$ in step {\ref{line:cond_grad}} can be computed efficiently.
-This is certainly not the case for dense CRFs, since computing $s^*$ involves $NL$ function evaluations of the submodular extension $F$, where $N$ is the number of variables, and $L$ is the number of labels.
-Each $F$ evaluation has complexity ${\mathcal O}(N)$ using the efficient Gaussian filtering algorithm of \citep{koltun2011efficient}.
-However, computation of $s^*$ would still be ${\mathcal O} (N^2)$ this way, which is clearly impractical for computer-vision applications where $N \sim 10^5 - 10^6$.
-
-However, using the equivalence of relaxed LP objectives and the Lovasz
-extension of submodular extensions in proposition
-\ref{proposition:potts_equiv}, we are able to compute $s^*$ in ${\mathcal
-O}(NL)$ time. Specifically, we use the algorithm of
-Ajanthan et al. \citep{ajanthan2017efficient}, which provides an efficient filtering procedure to compute the subgradient of the LP relaxation objective $E({\bf y})$ of (P-LP).\\
-%
-\vspace{-0.5cm}
-{\proposition Computing the subgradient of $E({\bf y})$ in (P-LP) is equivalent to computing the conditional gradient for the submodular function $F_{Potts}$. \label{proposition:subgrad}} 
-\newline(Proof in appendix).
-\begin{proof}
-  For the Potts model, we derived the worst-case optimal extension $F_{Potts}$ by
-  making its Lovasz extension $f({\bf y})$ same as the objective function
-    $E({\bf y})$ of the worst-case optimal LP relaxation. Hence, we have
-\begin{align}
-    E({\bf y}) &= f({\bf y})  \nonumber \\
-    {} &=  \max_{{\bf s} \in EP(F)} \langle {\bf y}, {\bf s} \rangle.\nonumber
-\end{align}
-
-The subdifferential of $E({\bf y})$ at $\bf y_0$ is the `active' linear function at $\bf y_0$. Hence,
-\begin{align}
-    \partial E({\bf y})|_{\bf y = y_0} &=  \argmax_{{\bf s} \in EP(F)} \langle {\bf y_0}, {\bf s} \rangle
-    \label{eq:subgrad}
-\end{align}
-Equation (\ref{eq:subgrad}) is equivalent to ${\bf s}^*$ computation in line \ref{line:cond_grad} of algorithm \ref{algo:fw_inference}, with ${\bf y_0} = - \nabla g({\bf s}_k)$.
-\end{proof}
-
-A similar observation can be made in case of hierarchical Potts model.
-Hence we have the first practical algorithm to compute upper bound of log-partition function of a dense CRF for Potts and metric energies.
 

### 2018-12-13 tex/paper.tex
Pawan's corrections made for Potts model.

@@
 
-\mysubsection{Potts Model}
-The Potts model, also known as the uniform metric, specifies the pairwise potentials $\phi_{ab}(x_a, x_b)$ in equation (\ref{eq:energy_min}) as follows:
+\mysubsection{Pairwise Potts Model}
+The pairwise Potts model, also known as the uniform metric, specifies the pairwise potentials $\phi_{ab}(x_a, x_b)$ in equation (\ref{eq:energy_min}) as follows:
 \begin{equation}
@@
 \end{equation}
-where $w_{ab}$ is the weight associated with edge $(a, b)$. 
+where $w_{ab}$ is the weight associated with edge $(a, b)$. There are no higher-order potential terms in this model as in equation (\ref{eq:energy_min}).
 %
@@
 \end{align}
-$\Delta$ is the union of $N$ probability simplices:
-\begin{equation}
-    \Delta = \{{\bf y}_a \in \mathbb{R}^L | {\bf y}_a \succeq 0  \textrm{ and } \langle {\bf 1}, {\bf y}_a \rangle = 1\}
-    \label{eq:delta2}
-\end{equation}
-where ${\bf y}_a$ is the component of $\bf y$ corresponding to the $a$-th variable. By the minimax theorem for LP, we can reorder the terms:
+where $\Delta$ is the domain as defined as in equation (\ref{eq:delta}). By the minimax theorem for LP, we can reorder the terms:
 \begin{equation}
@@
 
-The question now becomes how to recover the worst-case optimal submodular extension using $E({\bf y})$. The following propositions answers this question.
+The question now becomes how to recover the worst-case optimal submodular extension using $E({\bf y})$. The following proposition answers this question.
 {\proposition The worst-case optimal  submodular extension for Potts model is given by $F_{Potts}(A) = \sum_{i = 1}^L F_i(A)$, where

### 2018-12-14 tex/paper.tex
Corrections for pairwise Potts model done.

@@
 values of the set $A = \{v_{1, 1}, v_{4, 1}, v_{3, 2}$\}.}
-\label{fig:alternate}
+\label{fig:potts_extension_example}
 \end{figure}
@@
 \vspace{-1mm}
-
-Proposition \ref{proposition:potts_worst-case optimal} paves the way for us to
-identify the worst-case optimal extension for hierarchical Potts model, which
-we discuss in the following section.
+Proposition \ref{proposition:potts_worst-case optimal} paves the way for us to identify the worst-case optimal extension for hierarchical Potts model, which we discuss in the following section.
+Figure (\ref{fig:potts_extension_example}) shows an example where we compute $F_{Potts}(A)$ for a small graph and a given set $A$ representing an invalid labeling.
+Figure (\ref{fig:optimal_extension_graph}) shows the $st$-graph representation of $F_{Potts}$ for a small instance.
 
@@
 {\proposition Computing the subgradient of $E({\bf y})$ in (P-LP) is equivalent to computing the conditional gradient for the submodular function $F_{Potts}$. \label{proposition:subgrad}} 
-\newline(Proof in appendix).
 \begin{proof}

### 2018-12-16 tex/paper.tex
Higher-Potts model adjusted. There is some problem with the proof.

@@
 model, and helps to capture interesting properties of the image that cannot be
-done using pairwise models~\citep{vineet2014filter, kohli2009robust}. We show that our upper-bounds on synthetic problems are comparable to those from tree reweighted message passing (TRW)
+done using pairwise models~\citep{vineet2014filter, kohli2007p3}. We show that our upper-bounds on synthetic problems are comparable to those from tree reweighted message passing (TRW)
 ~\citep{wainwright2005new} for the case of sparse CRFs. Unlike our approach,
@@
 To this end, we define indicator variables $y_{ai}$ which equal 1 if $X_a = i$, and 0 otherwise.
-The following LP relaxation is the tightest known for Potts model in the worst-case, assuming the Unique Games Conjecture to be true \citep{manokaran2008sdp}
+The following LP relaxation is the tightest known for Potts model in the worst-case, assuming the Unique Games Conjecture to be true \citep{manokaran2008sdp}:
 \begin{align}
@@
 
-The simplest generalisation of pairwise Potts model to higher-order results in
-\textit{higher-order Potts model}. The clique potentials for these models are
-defined as
+One of the higher-order models used in practice is the \textit{higher-order Potts} model \citep{kappes2016higher}. For this model, the clique potentials $\phi_c({\bf x})$ in 
+the energy function (equation (\ref{eq:energy_hop_min})) are defined as
 
@@
 
+The higher-order Potts model differs from the $P^n$ Potts model \citep{kohli2007p3} in that the penalty for homogeneous labelings $\gamma_0$ is invariant to the label, that is, does not depend upon the label.
+We found this invariance property important in deriving the submodular extension from the LP relaxation.
+
 \myparagraph{LP Relaxation}
+We state the LP relaxation of the MAP estimation problem for higher-order Potts model.
+Let us assume the pairwise potentials in equation (\ref{eq:energy_hop_min}) to be Potts model.
+Also, we use the same set of indicator variables as for the Potts model.
+The following is an accurate LP relaxation for the higher-order Potts model:
+
 \begin{align}
     \text{(HOP-LP)} \quad \min_{\bf y} \enskip & E({\bf y}) =  \sum_{a = 1}^N
-    \sum_{i = 1}^L \phi_a(i)y_{ai} + \nonumber\\
-    &\sum_{c \in {\mathcal C}} \max_{(a, b) \in c} \frac{w_c}{2} \sum_{i = 1}^L \cdot |y_{ai} - y_{bi}| \nonumber\\
+    \sum_{i = 1}^L \phi_a(i)y_{ai} + 
+    \sum_{(a, b) \in {\mathcal N}} \sum_{i} \frac{w_{ab}}{2} \cdot |y_{ai} - y_{bi}| \nonumber\\
+    & + \quad \sum_{c \in {\mathcal C}} \max_{(a, b) \in c} \frac{w_c}{2} \sum_{i = 1}^L \cdot |y_{ai} - y_{bi}| \nonumber\\
     \text{s.t} \quad &{\bf y} \in \Delta.
@@
 \end{align}
-The constraints for (HOP-LP) are the same as for (P-LP).
+
+The objective function is similar to that for the Potts model along with the addition of terms corresponding to higher-order potentials.The constraints for (HOP-LP) are the same as for (P-LP).
+To the best of our knowledge, the above LP relaxation does not have worst-case tightness guarantee, hence the submodular extension we derive using it does not have worst-case optimality property.
 
 \myparagraph{LP-based Submodular Extension} We use the same $1-of-L$ encoding
-scheme as for Potts model. In a similar way, problem can be rewritten as
+scheme as for Potts model. Problem \ref{eq:min_upper_bound} then factorises as:
 
@@
   \begin{align}
-      F_{HOPotts}(A) &= \sum_a \sum_i \phi_{a}(i) [|A \cap \{v_{ai}\}| = 1] + \nonumber \\
+      F_{HOPotts}(A) &= F_{Potts}(A) + \nonumber \\
         &\sum_{c \in {\mathcal C}} \max_{(a, b) \in c} \frac{w_c}{2} \sum_{i = 1}^L [|A \cap \{v_{ai}, v_{bi}\}| = 1]
@@
 \label{proposition:hop_extension}}
+\newline
+We note that the component of the submodular extension corresponding to cliques is of the form of maximum of cut function (that are submodular) over all pairs in a clique.
+In general, maximum of submodular functions is not submodular.
+However, in our particular case, the maximum turns out to be submodular as shown in the proof below.
+
 \begin{proof}

### 2018-12-16 tex/paper.tex
Modified Pn Potts model to be diversity model now. Proof to be corrected.

@@
 
-\mysection{Accurate Submodular Extension for Higher-order Potts Model} The pairwise Potts model often fails
+\mysection{Accurate Submodular Extension for Higher-order Diversity Model} The pairwise Potts model often fails
 to capture useful image statistics because it restricts the order of the
@@
 
-One of the higher-order models used in practice is the \textit{higher-order Potts} model \citep{kappes2016higher}. For this model, the clique potentials $\phi_c({\bf x})$ in 
-the energy function (equation (\ref{eq:energy_hop_min})) are defined as
+A higher-order model useful in real-world applications is the \textit{diversity} model, that favours labeling where superpixels have fewer number of unique labels.
+As an example from semantic segmentation, it is preferable to have pixels belonging to a superpixel to be incorrectly labeled with two class labels rather than three or more class labels.
+Let $\Gamma({\bf x_c})$ be the set of unique labels assigned to variables in the clique $c$, and $\omega_c$ be the weight associated with the clique.
+In our case, we take the clique potentials as proportional to the number of unique labels in the clique $c$:
 
 \begin{equation}
-  \phi_c({\bf x}_c) = 
-  \begin{cases}
-    \gamma_0, &{\text{if}} \; \forall i, j \in c: x_i = x_j, \\
-    \gamma_1, &{\text{else}}. 
-  \end{cases} 
+  \phi_c({\bf x}_c) = \omega_c|\Gamma({\bf x_c})|
+\end{equation}
+where the notation $|A|$ denotes the cardinality of a set A.
+
+\myparagraph{LP Relaxation}
+
+First, let us consider the IP formulation of the MAP problem for our diversity model.
+Using the same set of indicator variables $y_{ai} = \{0, 1\}$ as for the Potts model, which are binary now, the clique potential $\phi({\bf x_c})$ can be represented as
+
+\begin{equation}
+  \phi_c({\bf x}_c) = \omega_c \sum_{i = 1}^L \max_{(a, b) \in c} \cdot |y_{ai} - y_{bi}| 
+\label{eq:hod_ip}
 \end{equation}
 
-The higher-order Potts model differs from the $P^n$ Potts model \citep{kohli2007p3} in that the penalty for homogeneous labelings $\gamma_0$ is invariant to the label, that is, does not depend upon the label.
-We found this invariance property important in deriving the submodular extension from the LP relaxation.
+For example, let the set of unique labels in a clique be $\Gamma = \{l_1, l_2, l_3\}$.
+This implies that $y_{ai} = 1$ for some, but not all, variables in the clique for $i = \{1, 2, 3\}$.
+Hence, $|y_{ai} - y_{bi}| = 1$ for some pairs $(a, b) \in c$ for $i = \{1, 2, 3\}$.
+Also, for any other label $i$, $|y_{ai} - y_{bi}| = 0$ for all $(a, b) \in c$ since $y_{ai} = 0$ for all variables $a$ for these labels $i$.
+For binary variables, the maximum possible value of $|y_{ai} - y_{bi}|$ is 1.
+Hence,$\phi_c({\bf x}_c) = \omega_c \cdot (1 + 1 + 1) = 3 \cdot \omega_c = \omega_c |\{l_1, l_2, l_3\}$|.
 
-\myparagraph{LP Relaxation}
-We state the LP relaxation of the MAP estimation problem for higher-order Potts model.
-Let us assume the pairwise potentials in equation (\ref{eq:energy_hop_min}) to be Potts model.
-Also, we use the same set of indicator variables as for the Potts model.
-The following is an accurate LP relaxation for the higher-order Potts model:
+Equation (\ref{eq:hod_ip}) suggests that our diversity model can be interpreted as sum of $P^n$ Ising models over all labels.
+The $Pn$ Ising model corresponding to label $i$ is $\max_{(a, b) \in c} \cdot |y_{ai} - y_{bi}|$ and it imposes a penalty of 0 if all variables in clique $c$ take the same label $i$, else the penalty is 1.
+
+We now relax the indicator variables $y_{ai}$ to lie in $[0, 1]$.
+Assuming the pairwise potentials in equation (\ref{eq:energy_hop_min}) to be Potts model, an accurate LP relaxation for the higher-order diversity model is given by:
 
 \begin{align}
-    \text{(HOP-LP)} \quad \min_{\bf y} \enskip & E({\bf y}) =  \sum_{a = 1}^N
+    \text{(HOD-LP)} \quad \min_{\bf y} \enskip & E({\bf y}) =  \sum_{a = 1}^N
     \sum_{i = 1}^L \phi_a(i)y_{ai} + 
     \sum_{(a, b) \in {\mathcal N}} \sum_{i} \frac{w_{ab}}{2} \cdot |y_{ai} - y_{bi}| \nonumber\\
-    & + \quad \sum_{c \in {\mathcal C}} \max_{(a, b) \in c} \frac{w_c}{2} \sum_{i = 1}^L \cdot |y_{ai} - y_{bi}| \nonumber\\
+    & + \quad \sum_{c \, \in {\mathcal C}} \frac{w_c}{2} \sum_{i = 1}^L \max_{(a, b) \in c} \cdot |y_{ai} - y_{bi}| \nonumber\\
     \text{s.t} \quad &{\bf y} \in \Delta.
@@
 \end{align}
-
-The objective function is similar to that for the Potts model along with the addition of terms corresponding to higher-order potentials.The constraints for (HOP-LP) are the same as for (P-LP).
+where the label (HOD-LP) stands for LP relaxation for higher-order diversity model.
+The objective function is similar to that for the Potts model along with the addition of terms corresponding to higher-order potentials.
+The constraints for (HOD-LP) are the same as for (P-LP).
 To the best of our knowledge, the above LP relaxation does not have worst-case tightness guarantee, hence the submodular extension we derive using it does not have worst-case optimality property.
 
-\myparagraph{LP-based Submodular Extension} We use the same $1-of-L$ encoding
+\myparagraph{LP-based Submodular Extension} We use the same $1\text{-of-}L$ encoding
 scheme as for Potts model. Problem \ref{eq:min_upper_bound} then factorises as:
@@
 where $f(.)$ is the Lovasz extension of $F(.)$. We are interested in making
-this LP same as the (), thereby enabling us to recover the submodular extension
-for higher-order Potts. The following proposition gives us the submodular
+this LP same as (HOD-LP) of equation (\ref{eq:hop-lp}), thereby enabling us to recover the submodular extension
+for higher-order diversity model. The following proposition gives us the submodular
 extension.
 {\proposition The Lovasz extension of the following submodular extension for
-  higher-order Potts model:
+  our higher-order diversity model
   \begin{align}
-      F_{HOPotts}(A) &= F_{Potts}(A) + \nonumber \\
-        &\sum_{c \in {\mathcal C}} \max_{(a, b) \in c} \frac{w_c}{2} \sum_{i = 1}^L [|A \cap \{v_{ai}, v_{bi}\}| = 1]
+      F_{HOD}(A) &= F_{Potts}(A) + \nonumber \\
+        &\sum_{c \in {\mathcal C}} \frac{w_c}{2} \sum_{i = 1}^L \max_{(a, b) \in c} [|A \cap \{v_{ai}, v_{bi}\}| = 1]
 \end{align}
-is the objective function $E({\bf y})$ of the LP relaxation (HOP-LP).
-\label{proposition:hop_extension}}
+is the objective function $E({\bf y})$ of the LP relaxation (HOD-LP).
+\label{proposition:hod_extension}}
 \newline

### 2018-12-16 tex/paper.tex
Proof for higher-order diversity completed.

@@
   \phi_c({\bf x}_c) = \omega_c|\Gamma({\bf x_c})|
+\label{eq:diversity_potential}
 \end{equation}
@@
 \begin{equation}
-  \phi_c({\bf x}_c) = \omega_c \sum_{i = 1}^L \max_{(a, b) \in c} \cdot |y_{ai} - y_{bi}| 
+  \phi_c({\bf x}_c) = \omega_c \cdot \sum_{i = 1}^L \max_{(a, b) \in c} |y_{ai} - y_{bi}| 
 \label{eq:hod_ip}
@@
     \sum_{(a, b) \in {\mathcal N}} \sum_{i} \frac{w_{ab}}{2} \cdot |y_{ai} - y_{bi}| \nonumber\\
-    & + \quad \sum_{c \, \in {\mathcal C}} \frac{w_c}{2} \sum_{i = 1}^L \max_{(a, b) \in c} \cdot |y_{ai} - y_{bi}| \nonumber\\
+    & + \quad \sum_{c \, \in {\mathcal C}} w_c \cdot \sum_{i = 1}^L \max_{(a, b) \in c} |y_{ai} - y_{bi}| \nonumber\\
     \text{s.t} \quad &{\bf y} \in \Delta.
@@
 where $f(.)$ is the Lovasz extension of $F(.)$. We are interested in making
-this LP same as (HOD-LP) of equation (\ref{eq:hop-lp}), thereby enabling us to recover the submodular extension
-for higher-order diversity model. The following proposition gives us the submodular
-extension.
-{\proposition The Lovasz extension of the following submodular extension for
-  our higher-order diversity model
+this LP same as (HOD-LP) of equation (\ref{eq:hop-lp}), thereby enabling us to recover the submodular extension for higher-order diversity model.
+The following lemma will be of use to us in indentifying the extension.
+
+{\lemma The Lovasz extension of the following set function 
   \begin{align}
-      F_{HOD}(A) &= F_{Potts}(A) + \nonumber \\
-        &\sum_{c \in {\mathcal C}} \frac{w_c}{2} \sum_{i = 1}^L \max_{(a, b) \in c} [|A \cap \{v_{ai}, v_{bi}\}| = 1]
+      F_{HOD}(A) &= F_{Potts}(A) + \nonumber \sum_{c \in {\mathcal C}} w_c \cdot \sum_{i = 1}^L \max_{(a, b) \in c} [|A \cap \{v_{ai}, v_{bi}\}| = 1]
+\label{hod_submodular}
 \end{align}
 is the objective function $E({\bf y})$ of the LP relaxation (HOD-LP).
-\label{proposition:hod_extension}}
+\label{lemma:hod_extension}}
 \newline
@@
 However, in our particular case, the maximum turns out to be submodular as shown in the proof below.
+Our proof makes use of the following property:
 
 \begin{proof}
-  We first observe that $\forall {\bf y} \in {\mathcal L}^N$, that is, for all
-  valid labelings, $F_{HOPotts}(A_{\bf y})$ equals $E({\bf y}$.  Hence, $F_{HOPotts}$
-    is an extension.
+ Making use of proposition \ref{proposition:potts_worst-case optimal}, it will be sufficient for us to show that the Lovasz extension of the $i$-th Ising model $F_i(A) = \max_{(a, b) \in c} [|A \cap \{v_{ai}, v_{bi}\}| = 1]$ is $\max_{(a, b) \in c} |y_{ai} - y_{bi}|$.
 
-Consider a clique with 3 variables $X_a, X_b, X_c$
-each of which can take one of three labels $i = 1, 2, 3$. Then ${\bf x}
-= [x_{a1}, x_{a2},x_{a3},x_{b1},x_{b2},x_{b3},x_{c1},x_{c2},x_{c3}]$ is the
-vector of relaxed variables. To compute $E({\bf x})$ and $f({\bf x})$, it will
-suffice to specify per label ordering. So let us assume $x_{a1} > x_{b1}
-> x_{c1}$, $x_{a2} > x_{b2} > x_{c2}$ and $x_{a3} > x_{b3} > x_{c3}$.
-
-Then, the relaxed objective is
-
+Let the clique $c$ have $M$ variables.
+Given ${\bf y} \in [0, 1]^M$, let us order its components in decreasing order $y_{j_1} > \dots > y_{j_M}$ , where $(j_1,\dots, j_M)$ is a permutation.
+Then the Lovasz extension of $F_i()$ is:
 \begin{equation}
-  E({\bf x}) = \frac{w_c}{2} \cdot [(x_{a1} - x_{c1}) + (x_{a2} - x_{c2}) +(x_{a3} - x_{c3})  ]
+f_i({\bf y}) = \sum_{k = 1}^M y_{j_k} [F_i(\{j_1, \dots, j_k\}) - F_i(\{j_1, \dots, j_{k - 1}\})]
+\nonumber
 \end{equation}
 
-It can be shown that regardless of inter-labeling rankings, the L.E is given
-by:
+We note that when $A = \{ \}$ or $\{j_1, \dots, j_M\}$, that is, when $A$ represents homogeneous labelings, $F_i(A) = 0$.
+For any other $A \subset \{j_1, \dots, j_M\}$, $F_i(A) = 1$.
+Hence, the Lovasz extension simplifies to
+\begin{align}
+f_i({\bf y}) &= y_{j_1} - y_{j_M} \nonumber \\
+&= \max_{(a, b) \in c} |y_{ai} - y_{bi}| \nonumber
+\end{align}
+This proves our lemma.
+\end{proof}
 
-\begin{equation}
-  f({\bf x}) = \frac{w_c}{2} \cdot [(x_{a1} - x_{c1}) + (x_{a2} - x_{c2}) +(x_{a3} - x_{c3})  ]
-\end{equation}
+We are now in a position to state our main result:
 
-One can generalise this over more cliques, variables, and labels to show the equivalence of $E({\bf x})$ and $f({\bf x})$.
-The Lovasz extension of a set function is convex iff it is submodular. Since $E({\bf x})$
-is convex, this implies that $F_{HOPotts}$ is submodular. 
+{\proposition The set function $F_{HOD}$ of lemma \ref{lemma:hod_extension} is the submodular extension of our higher-order diversity model whose Lovasz extension is $E({\bf y})$ in problem (HOD-LP).}
+
+\begin{proof}
+We first observe that $\forall {\bf y} \in {\mathcal L}^N$, that is, for all valid labelings, $F_{HOD}(A_{\bf y})$ equals $E({\bf y})$.  Hence, $F_{HOD}$ is an extension.
+
+Also, we showed in lemma \ref{lemma:hod_extension} that the Lovasz extension of $F_{HOD}$ is $E({\bf y})$ in probem (HOD-LP).
+$E({\bf y})$ being a sum of terms corresponding to maximum of linear functions, is convex.
+The convexity of Lovasz extension implies submodularity of the set function \citep{bach2013learning}.
+Hence, $F_{HOD}$ is submodular.
 \end{proof}
+
+We now show the effectiveness of the accurate submodular extensions for different classes of models by means of experiments on synthetic and real-world datasets.
 
@@
 
-\mysubsection{Stereo Matching using Higher-order Potts Model}
-\label{subsec:stereo_hop}
-\myparagraph{Data} Next, we use the higher-order Potts model for stereo matching on the Middlebury dataset.
+%\mysubsection{Stereo Matching using Higher-order Potts Model}
+%\label{subsec:stereo_hop}
+%\myparagraph{Data} Next, we use the higher-order Potts model for stereo matching on the Middlebury dataset.
 Higher-order Potts model is suitable to be used for only some images in the dataset, since others have gradually sloping surfaces and higher-order cliques will force all pixels on these surfaces to take similar disparity values.
@@
 We obtain higher-order cliques by using super-pixels obtained from mean-shift algorithm \citep{comaniciu2002mean}.
-
-\myparagraph{Method} We again use the modified Gaussian filtering implementation for dense CRFs by
-\citep{ajanthan2017efficient} to compute the conditional gradient at each step.
-To compute the contribution of the higher-order Potts, we make use of an
-efficient strategy to reuse computation. Recall that computing the conditional
-gradient coordinates $s^*_{\sigma(i)}$ requires us to obtain differences of the
-nested sets $F(S_i) - F(S_{i - 1})$. We store the clique state at each $i$, and
-this enables us to compute the difference for each coordinate in a constant
-amount of time. In this case, computation of conditional gradient takes
-$\mathcal{O}(cNL)$ where $c$ is the number of higher-order cliques. We ran our
-algorithm for 200 iterations.
-
-\myparagraph{Results} Figure \ref{fig:stereo_hop} shows some example solutions
-obtained by picking the label with maximum marginal probability for each
-variable. We observe that our method gives results that are reasonably close to
-the ground-truth. The set of hyperparameters that we used can be found in the
-appendix.
+%
+%\myparagraph{Method} We again use the modified Gaussian filtering implementation for dense CRFs by
+%\citep{ajanthan2017efficient} to compute the conditional gradient at each step.
+%To compute the contribution of the higher-order Potts, we make use of an
+%efficient strategy to reuse computation. Recall that computing the conditional
+%gradient coordinates $s^*_{\sigma(i)}$ requires us to obtain differences of the
+%nested sets $F(S_i) - F(S_{i - 1})$. We store the clique state at each $i$, and
+%this enables us to compute the difference for each coordinate in a constant
+%amount of time. In this case, computation of conditional gradient takes
+%$\mathcal{O}(cNL)$ where $c$ is the number of higher-order cliques. We ran our
+%algorithm for 200 iterations.
+%
+%\myparagraph{Results} Figure \ref{fig:stereo_hop} shows some example solutions
+%obtained by picking the label with maximum marginal probability for each
+%variable. We observe that our method gives results that are reasonably close to
+%the ground-truth. The set of hyperparameters that we used can be found in the
+%appendix.
 

### 2018-12-21 tex/abstract.tex
Incorporating Pawan's changes. Abstract modified.

@@
     the marginals depends crucially on the quality of the submodular extension.
-    To identify accurate extensions for different classes of energy functions, we show a relation between the
-    submodular extensions of the energy and the objective functions of linear
-    programming (LP) relaxations for the corresponding MAP estimation problem via the Lovasz extension. 
+    To identify accurate extensions for different classes of energy functions, we establish a relationship between the submodular extensions of the energy and linear programming (LP) relaxations for the corresponding MAP estimation problem.
     This allows us to (i) establish the worst-case optimality of the submodular
@@
     used dense CRF model with the help of a recently proposed Gaussian
-    filtering method; and (iv) propose a good submodular extension based on an accurate LP
-    relaxation for the higher-order Potts model.
+    filtering method; and (iv) propose an accurate submodular extension based on an LP
+    relaxation for a higher-order diversity model.
 Using synthetic and real data, we show that our approach provides comparable upper bounds on the log-partition function to those obtained using tree-reweighted message passing (TRW) in cases where the latter is computationally feasible.

### 2018-12-21 tex/paper.tex
Introduction modified.

@@
 connection has four important practical consequences. First, it establishes
-the accuracy of the submodular extension of the Potts model, via the
-UGC-hardness worst-case optimality of the LP relaxation. Second, it provides an
+the optimality of the submodular extension of the Potts model, via the tightest LP relaxation
+under UGC-hardness assumptions. Second, it provides an
 accurate submodular extension of the hierarchical Potts model, via the LP
@@
 inference algorithm for dense CRFs, which computes a lower bound on the
-log-partition ~\citep{koltun2011efficient}. Fourthly, we obtain a good 
-submodular extension for higher-order Potts model based on an accurate LP
-relaxation. Higher-order Potts model is a generalisation of the pairwise Potts
-model, and helps to capture interesting properties of the image that cannot be
+log-partition ~\citep{koltun2011efficient}. Fourth, we obtain an accurate 
+submodular extension for a higher-order diversity model based on an LP
+relaxation. Higher-order models can capture interesting properties of the image that cannot be
 done using pairwise models~\citep{vineet2014filter, kohli2007p3}. We show that our upper-bounds on synthetic problems are comparable to those from tree reweighted message passing (TRW)
@@
 practice. Using pairwise dense CRF models, we perform stereo matching on standard
-data sets and obtain better results than \citep{koltun2011efficient}. We also perform semantic segmentation on the MSRC-21 dataset
-
+data sets and obtain better results than \citep{koltun2011efficient}.
+We also perform semantic segmentation on the MSRC-21 dataset using a combination of dense pairwise and higher-order diversity model.
 The complete code is available at \url{https://github.com/pankajpansari/denseCRF}.

### 2018-12-21 tex/paper.tex
Prelimiaries modified.

@@
 We can augment the above pairwise models with higher-order potentials.
-Let $\mathcal C$ be the set of cliques, which can, for instance, be the set of superpixels from an image segmentation method.
-Also, let $\bf x_c$ be the component of $\bf x$ formed by elements in clique $c$.
+Let $\mathcal C$ be the set of cliques, that is, fully-connected groups of variables.
+These cliques can, for instance, be the set of superpixels from an image segmentation method.
+Also, let ${\bf x}_c$ be the component of $\bf x$ formed by elements in clique $c$.
 The energy function now also contains higher-order potentials
@@
 \begin{equation}
-    E({\bf x}) = \sum_{a \in {\mathcal X}} \phi_a(x_a) + \sum_{(a, b) \in {\mathcal N}} \phi_{ab}(x_a, x_b) + \sum_{c \in {\mathcal C}} \phi_c({\bf x_c}),
+    E({\bf x}) = \sum_{a \in {\mathcal X}} \phi_a(x_a) + \sum_{(a, b) \in {\mathcal N}} \phi_{ab}(x_a, x_b) + \sum_{c \in {\mathcal C}} \phi_c({{\bf x}_c}),
     \label{eq:energy_hop_min}
@@
 As an example from semantic segmentation, it is preferable to have pixels belonging to a superpixel to be incorrectly labeled with two class labels rather than three or more class labels.
-Let $\Gamma({\bf x_c})$ be the set of unique labels assigned to variables in the clique $c$, and $\omega_c$ be the weight associated with the clique.
+Let $\Gamma({\bf x}_c)$ be the set of unique labels assigned to variables in the clique $c$, and $\omega_c$ be the weight associated with the clique.
 In our case, we take the clique potentials as proportional to the number of unique labels in the clique $c$:
@@
 \begin{equation}
-  \phi_c({\bf x}_c) = \omega_c|\Gamma({\bf x_c})|
+  \phi_c({\bf x}_c) = \omega_c|\Gamma({\bf x}_c)|
 \label{eq:diversity_potential}
@@
 First, let us consider the IP formulation of the MAP problem for our diversity model.
-Using the same set of indicator variables $y_{ai} = \{0, 1\}$ as for the Potts model, which are binary now, the clique potential $\phi({\bf x_c})$ can be represented as
+Using the same set of indicator variables $y_{ai} = \{0, 1\}$ as for the Potts model, which are binary now, the clique potential $\phi({\bf x}_c)$ can be represented as
 

### 2018-12-21 tex/paper.tex
Overview modified.

@@
 
-\mysection{Accurate Submodular Extensions from LP Relaxations}
+\mysection{Overview: Accurate Submodular Extensions from LP Relaxations}
 
 Different choices of extensions $F$ change the domain in problem \eqref{eq:min_upper_bound}, leading to different upper bounds on the log-partition function.
-How does one come up with an extension which yields the tightest bound?
+How does one come up with accurate extensions for different classes of CRF energy functions?
+Is it possible to identify optimal extensions which yields the tightest bound?
 
@@
   Motivated by this connection, we ask if one can introduce a temperature parameter to our problem \eqref{eq:min_upper_bound} and transform it to an LP relaxation in the limit $T \to 0$?
-We can then hope to use accurate LP relaxations of MAP problems known in literature to find good submodular extensionsfor different classes of energy functions.
+We can then hope to use accurate LP relaxations of MAP problems known in literature to find good submodular extensions for different classes of energy functions.
 
 We answer this question in affirmative.
-Specifically, in the following two sections we show how one can select the set encoding and submodular extension to convert problem \eqref{eq:min_upper_bound} to accurate LP relaxations for Potts, hierarchical Potts and higher-order Potts models.
+Specifically, in the following two sections we show how one can select the set encoding and submodular extension to convert problem \eqref{eq:min_upper_bound} to accurate LP relaxations for Potts, hierarchical Potts and higher-order diversity models.
 In fact, when the LP relaxation has tightness guarantees, we obtain worst-case optimal submodular extensions - a notion we elucidate shortly.

### 2018-12-21 tex/paper.tex
Section on worst-case optimal submodular extension modified.

@@
 
+In this paper, we focus on submodular extension families $\mathcal F(.)$ corresponding to different classes of energy functions $\mathcal E$.
+Given an instance of the energy function $E(.)$ from the class $\mathcal E$,  there is a corresponding submodular extension ${\mathcal F}(E)$.
+Also, we seek a closed-form analytical expression for $\mathcal F$.
+
 \mysection{Worst-case Optimal Submodular Extensions}
@@
 \myparagraph{Worst-case Optimality}
-In this paper, we focus on submodular extension families $\mathcal F(.)$ which
-for each instance of the energy function $E(.)$ belonging to a given class $\mathcal E$ gives a corresponding
-submodular extension ${\mathcal F}(E)$. Our extensions are derived from LP relaxations. When the LP relaxation is the tightest known, the extension family ${\mathcal
+
+Our extensions are derived from LP relaxations. When the LP relaxation is the tightest known, the extension family ${\mathcal
 F}_{opt}$ we obtain is \emph{worst-case optimal}. This implies that there does not
@@
 Note that the notion of worst-case optimality does not guarantee that the extension is best for every instance of the energy function.
-Also, we seek a closed-form analytical expression for $\mathcal F$. For the sake of clarity,
-in the analysis that follows we use $F$ to represent ${\mathcal F}(E)$ where
-the meaning is clear from context.
+It may be possible to solve an optimisation problem to obtain the best extension for every given instace of energy $E(.)$.
+However, in this paper, we do not take that approach, and instead provide a general solution for a given class of energy function.
+
+In subsection \ref{subsec:potts}, we prove the worst-case optimality of the submodular extension used in literature for the Potts model.
+In subsection \ref{subsec:hier_potts}, we obtain worst-case optimal submodular extension for the more general hierarchical Potts model.
+Finally, in subsection \ref{subsec:dense_crf}, we provide an approach to make our inference algorithm efficient for the dense CRF model.
+Using our approach, the inference algorithm has linear time-complexity in the number of variables and labels.
 
 \mysubsection{Pairwise Potts Model}
+\label{subsec:potts}
 The pairwise Potts model, also known as the uniform metric, specifies the pairwise potentials $\phi_{ab}(x_a, x_b)$ in equation (\ref{eq:energy_min}) as follows:
@@
 \mysubsection{Hierarchical Potts Model}
-\label{sec:rhst_worst-case optimal}
+\label{subsec:hier_potts}
 
@@
 \mysubsection{Fast Conditional Gradient Computation for Dense CRFs}
-\label{sec:dense_crf}
+\label{subsec:dense_crf}
 

### 2018-12-21 tex/paper.tex
Higher-order section modified.

@@
 
-A higher-order model useful in real-world applications is the \textit{diversity} model, that favours labeling where superpixels have fewer number of unique labels.
-As an example from semantic segmentation, it is preferable to have pixels belonging to a superpixel to be incorrectly labeled with two class labels rather than three or more class labels.
+A higher-order model useful in real-world applications is the \textit{diversity} model, which favours labeling where variables in a clique have fewer number of unique labels.
+For instance, in semantic segmentataion we often first obtain superpixels from an image segmentation method, which we consider as cliques in our model.
+We expect the labeling in a superpixel to be homogeneous.
+As a result, it is preferable to have pixels belonging to a superpixel to be incorrectly labeled with two class labels rather than three or more class labels.
 Let $\Gamma({\bf x}_c)$ be the set of unique labels assigned to variables in the clique $c$, and $\omega_c$ be the weight associated with the clique.
@@
 For binary variables, the maximum possible value of $|y_{ai} - y_{bi}|$ is 1.
-Hence,$\phi_c({\bf x}_c) = \omega_c \cdot (1 + 1 + 1) = 3 \cdot \omega_c = \omega_c |\{l_1, l_2, l_3\}$|.
-
-Equation (\ref{eq:hod_ip}) suggests that our diversity model can be interpreted as sum of $P^n$ Ising models over all labels.
-The $Pn$ Ising model corresponding to label $i$ is $\max_{(a, b) \in c} \cdot |y_{ai} - y_{bi}|$ and it imposes a penalty of 0 if all variables in clique $c$ take the same label $i$, else the penalty is 1.
+Hence, $\phi_c({\bf x}_c) = \omega_c \cdot (1 + 1 + 1) = 3 \cdot \omega_c = \omega_c |\{l_1, l_2, l_3\}$|.
 
@@
 The constraints for (HOD-LP) are the same as for (P-LP).
-To the best of our knowledge, the above LP relaxation does not have worst-case tightness guarantee, hence the submodular extension we derive using it does not have worst-case optimality property.
+The above LP has not been formally analysed in literature.
+Hence, we do not make any claims regarding its optimality.
+However, its similairty in form to (P-LP) is an indication of its accuracy.
+We establish its accuracy empirically in section \ref{sec:exp}.
 
@@
 
-Introducing a temperature parameter $T$ to problem (\ref{eq:hop_upper_bound}),
+We introduce a temperature parameter $T$ to problem (\ref{eq:hop_upper_bound}) to obtain the following new problem:
 
@@
 
-and considering the limit $T \to 0$, problem (\ref{eq:hop_temp}) becomes
+Considering the limit $T \to 0$, problem (\ref{eq:hop_temp}) becomes
 
@@
 where $f(.)$ is the Lovasz extension of $F(.)$. We are interested in making
-this LP same as (HOD-LP) of equation (\ref{eq:hop-lp}), thereby enabling us to recover the submodular extension for higher-order diversity model.
-The following lemma will be of use to us in indentifying the extension.
+this LP same as (HOD-LP) of equation (\ref{eq:hop-lp}), thereby enabling us to obtain the submodular extension for the higher-order diversity model.
+We will make use of the following lemma in our proof.
 
@@
 \newline
-We note that the component of the submodular extension corresponding to cliques is of the form of maximum of cut function (that are submodular) over all pairs in a clique.
-In general, maximum of submodular functions is not submodular.
-However, in our particular case, the maximum turns out to be submodular as shown in the proof below.
-Our proof makes use of the following property:
 

### 2018-12-21 tex/paper.tex
Higher-order diversity section modified.

@@
     \centering
-\begin{tabular}{cccccc}
-        \includegraphics[scale = 0.14]{\imagePath/stereo/venus_GT} &
-        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/venus_left_mf} &
-        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/venus_left_submod} &
+\begin{tabular}{ccc}
+        \includegraphics[scale = 0.24]{\imagePath/stereo/venus_GT} &
+        \includegraphics[scale = 0.24]{\imagePath/stereo/potts/venus_left_mf} &
+        \includegraphics[scale = 0.24]{\imagePath/stereo/potts/venus_left_submod} \\
     %
-        \includegraphics[scale = 0.19]{\imagePath/stereo/tsukuba_GT} &
-        \includegraphics[scale = 0.19]{\imagePath/stereo/potts/tsukuba_left_mf} &
-        \includegraphics[scale = 0.19]{\imagePath/stereo/potts/tsukuba_left_submod} \\
-        \scriptsize(a) Venus GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution & \scriptsize(a) Tsukuba GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
-        {} & \scriptsize  60.32s, 1.83e+07 &  \scriptsize 469.75s, {\bf 1.55e+07} & {} & \scriptsize  14.93s, 8.21e+06 & \scriptsize 215.22s, {\bf 4.12e+06} \\
-         \includegraphics[scale = 0.14]{\imagePath/stereo/cones_GT} &
-        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/cones_left_mf} &
-        \includegraphics[scale = 0.14]{\imagePath/stereo/potts/cones_left_submod} &
+        \scriptsize(a) Venus GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution \\ 
+        {} & \scriptsize  60.32s, 1.83e+07 &  \scriptsize 469.75s, {\bf 1.55e+07}  \\
+        \includegraphics[scale = 0.27]{\imagePath/stereo/tsukuba_GT} &
+        \includegraphics[scale = 0.27]{\imagePath/stereo/potts/tsukuba_left_mf} &
+        \includegraphics[scale = 0.27]{\imagePath/stereo/potts/tsukuba_left_submod} \\
+        \scriptsize(a) Tsukuba GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
+        {} & \scriptsize  14.93s, 8.21e+06 & \scriptsize 215.22s, {\bf 4.12e+06} \\
+
+        \includegraphics[scale = 0.24]{\imagePath/stereo/cones_GT} &
+        \includegraphics[scale = 0.24]{\imagePath/stereo/potts/cones_left_mf} &
+        \includegraphics[scale = 0.24]{\imagePath/stereo/potts/cones_left_submod} \\
+        \scriptsize(a) Cones GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution \\
+        {} & \scriptsize 239.14s, 2.68e+07 &  \scriptsize 1082.72s, {\bf 1.27e+07} \\ 
         %
        %
-         \includegraphics[scale = 0.15]{\imagePath/stereo/teddy_GT} &
-        \includegraphics[scale = 0.15]{\imagePath/stereo/potts/teddy_left_mf} &
-        \includegraphics[scale = 0.15]{\imagePath/stereo/potts/teddy_left_submod} \\
-    \scriptsize(a) Cones GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution & \scriptsize(a) Teddy GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
-        {} & \scriptsize 239.14s, 2.68e+07 &  \scriptsize 1082.72s, {\bf 1.27e+07} & {} & \scriptsize 555.30s, 2.36e+07 & \scriptsize 1257.86s, {\bf 1.58e+07}\\
+         \includegraphics[scale = 0.25]{\imagePath/stereo/teddy_GT} &
+        \includegraphics[scale = 0.25]{\imagePath/stereo/potts/teddy_left_mf} &
+        \includegraphics[scale = 0.25]{\imagePath/stereo/potts/teddy_left_submod} \\
+        \scriptsize(a) Teddy GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution\\
+        {} & \scriptsize 555.30s, 2.36e+07 & \scriptsize 1257.86s, {\bf 1.58e+07}\\
 \end{tabular}
@@
 
-\begin{figure*}
-    \centering
-\begin{tabular}{cccccc}
-        \includegraphics[scale = 0.22]{\imagePath/stereo/higher_potts/tsukuba_GT} &
-        \includegraphics[scale = 0.22]{\imagePath/stereo/higher_potts/tsukuba_hop} &
-        \includegraphics[scale = 0.34]{\imagePath/stereo/higher_potts/baby1_GT} &
-        \includegraphics[scale = 0.16]{\imagePath/stereo/higher_potts/baby1_hop} &
-        \includegraphics[scale = 0.34]{\imagePath/stereo/higher_potts/flowerpots_GT} &
-        \includegraphics[scale = 0.16]{\imagePath/stereo/higher_potts/flowerpots_hop} \\
-        \scriptsize(a) Tsukuba GT & \scriptsize(b) Submod solution
-        & \scriptsize(a) Baby1 GT & \scriptsize(b) Submod solution
-        & \scriptsize(a) Flowerpots GT & \scriptsize(b) Submod solution\\
-        {} & {} 
-        & {} & \scriptsize 2476.82s
-        & {} & \scriptsize 2471.58s\\
-
-\end{tabular}
-%\vspace{2mm}
-\mycaption{\footnotesize \em Stereo matching using dense CRFs with higher-order
-  Potts term in addition to Potts compatibility and Gaussian pairwise potentials.}
-\label{fig:stereo_hop}
-\end{figure*}
-
-%\mysubsection{Stereo Matching using Higher-order Potts Model}
+%\begin{figure*}
+%    \centering
+%\begin{tabular}{cccccc}
+%        \includegraphics[scale = 0.22]{\imagePath/stereo/higher_potts/tsukuba_GT} &
+%        \includegraphics[scale = 0.22]{\imagePath/stereo/higher_potts/tsukuba_hop} &
+%        \includegraphics[scale = 0.34]{\imagePath/stereo/higher_potts/baby1_GT} &
+%        \includegraphics[scale = 0.16]{\imagePath/stereo/higher_potts/baby1_hop} &
+%        \includegraphics[scale = 0.34]{\imagePath/stereo/higher_potts/flowerpots_GT} &
+%        \includegraphics[scale = 0.16]{\imagePath/stereo/higher_potts/flowerpots_hop} \\
+%        \scriptsize(a) Tsukuba GT & \scriptsize(b) Submod solution
+%        & \scriptsize(a) Baby1 GT & \scriptsize(b) Submod solution
+%        & \scriptsize(a) Flowerpots GT & \scriptsize(b) Submod solution\\
+%        {} & {} 
+%        & {} & \scriptsize 2476.82s
+%        & {} & \scriptsize 2471.58s\\
+%
+%\end{tabular}
+%%\vspace{2mm}
+%\mycaption{\footnotesize \em Stereo matching using dense CRFs with higher-order
+%  Potts term in addition to Potts compatibility and Gaussian pairwise potentials.}
+%\label{fig:stereo_hop}
+%\end{figure*}
+%
+%\mysubsection{Stereo Matching using Higher-order Diversity Model}
 %\label{subsec:stereo_hop}
-%\myparagraph{Data} Next, we use the higher-order Potts model for stereo matching on the Middlebury dataset.
+%\myparagraph{Data} Next, we use the higher-order diversity model for stereo matching on the Middlebury dataset.
 Higher-order Potts model is suitable to be used for only some images in the dataset, since others have gradually sloping surfaces and higher-order cliques will force all pixels on these surfaces to take similar disparity values.
@@
 
+\mysubsection{Stereo Matching using Higher-order Diversity Model}
+TO-DO
+
+\mysubsection{Stereo Matching using Higher-order Diversity Model}
+TO-DO
+
 \mysection{Discussion}

### 2019-01-03 tex/paper.tex
Version sent to Chris for review.

@@
 
-%\begin{figure*}
-%    \centering
-%\begin{tabular}{cccccc}
-%        \includegraphics[scale = 0.22]{\imagePath/stereo/higher_potts/tsukuba_GT} &
-%        \includegraphics[scale = 0.22]{\imagePath/stereo/higher_potts/tsukuba_hop} &
-%        \includegraphics[scale = 0.34]{\imagePath/stereo/higher_potts/baby1_GT} &
-%        \includegraphics[scale = 0.16]{\imagePath/stereo/higher_potts/baby1_hop} &
-%        \includegraphics[scale = 0.34]{\imagePath/stereo/higher_potts/flowerpots_GT} &
-%        \includegraphics[scale = 0.16]{\imagePath/stereo/higher_potts/flowerpots_hop} \\
-%        \scriptsize(a) Tsukuba GT & \scriptsize(b) Submod solution
-%        & \scriptsize(a) Baby1 GT & \scriptsize(b) Submod solution
-%        & \scriptsize(a) Flowerpots GT & \scriptsize(b) Submod solution\\
-%        {} & {} 
-%        & {} & \scriptsize 2476.82s
-%        & {} & \scriptsize 2471.58s\\
-%
-%\end{tabular}
-%%\vspace{2mm}
-%\mycaption{\footnotesize \em Stereo matching using dense CRFs with higher-order
-%  Potts term in addition to Potts compatibility and Gaussian pairwise potentials.}
-%\label{fig:stereo_hop}
-%\end{figure*}
-%
-%\mysubsection{Stereo Matching using Higher-order Diversity Model}
-%\label{subsec:stereo_hop}
-%\myparagraph{Data} Next, we use the higher-order diversity model for stereo matching on the Middlebury dataset.
-Higher-order Potts model is suitable to be used for only some images in the dataset, since others have gradually sloping surfaces and higher-order cliques will force all pixels on these surfaces to take similar disparity values.
-We take a dense CRF model with Potts compatibility term and Gaussian pairwise potentials, and augment it with higher-order Potts term.
-We obtain higher-order cliques by using super-pixels obtained from mean-shift algorithm \citep{comaniciu2002mean}.
-%
-%\myparagraph{Method} We again use the modified Gaussian filtering implementation for dense CRFs by
-%\citep{ajanthan2017efficient} to compute the conditional gradient at each step.
-%To compute the contribution of the higher-order Potts, we make use of an
-%efficient strategy to reuse computation. Recall that computing the conditional
-%gradient coordinates $s^*_{\sigma(i)}$ requires us to obtain differences of the
-%nested sets $F(S_i) - F(S_{i - 1})$. We store the clique state at each $i$, and
-%this enables us to compute the difference for each coordinate in a constant
-%amount of time. In this case, computation of conditional gradient takes
-%$\mathcal{O}(cNL)$ where $c$ is the number of higher-order cliques. We ran our
-%algorithm for 200 iterations.
-%
-%\myparagraph{Results} Figure \ref{fig:stereo_hop} shows some example solutions
-%obtained by picking the label with maximum marginal probability for each
-%variable. We observe that our method gives results that are reasonably close to
-%the ground-truth. The set of hyperparameters that we used can be found in the
-%appendix.
+\begin{figure*}
+    \centering
+\begin{tabular}{cc}
+        \includegraphics[scale = 0.30]{\imagePath/stereo/higher_diversity/tsukuba_GT} &
+        \includegraphics[scale = 0.30]{\imagePath/stereo/higher_diversity/tsukuba_hop} \\
+        \scriptsize(a) Tsukuba GT & \scriptsize(b) Submod solution \\
+        {} & {}  \\
+        \includegraphics[scale = 0.52]{\imagePath/stereo/higher_diversity/baby1_GT} & 
+        \includegraphics[scale = 0.26]{\imagePath/stereo/higher_diversity/baby1_hop} \\ 
+        \scriptsize(a) Baby1 GT & \scriptsize(b) Submod solution \\
+        {} & \scriptsize 2476.82s \\
+        \includegraphics[scale = 0.52]{\imagePath/stereo/higher_diversity/flowerpots_GT}&  
+        \includegraphics[scale = 0.26]{\imagePath/stereo/higher_diversity/flowerpots_hop} \\
+        \scriptsize(a) Flowerpots GT & \scriptsize(b) Submod solution\\
+        {} & \scriptsize 2471.58s\\
+
+\end{tabular}
+%\vspace{2mm}
+\mycaption{\footnotesize \em Stereo matching using dense CRFs with higher-order
+  diversity term in addition to Potts compatibility and Gaussian pairwise potentials.}
+\label{fig:stereo_hod}
+\end{figure*}
 
 \mysubsection{Stereo Matching using Higher-order Diversity Model}
-TO-DO
-
-\mysubsection{Stereo Matching using Higher-order Diversity Model}
-TO-DO
+\label{subsec:stereo_hod}
+\myparagraph{Data} Next, we use the higher-order diversity model for stereo
+matching on the Middlebury dataset. A higher-order model is suitable to be used
+for only some images in the dataset, since others have gradually sloping
+surfaces and higher-order cliques will force all pixels on these surfaces to
+take similar disparity values.  We take a dense CRF model with Potts
+compatibility term and Gaussian pairwise potentials, and augment it with
+higher-order diversity term. We obtain higher-order cliques by using super-pixels obtained from the mean-shift algorithm \citep{comaniciu2002mean}.
+
+\myparagraph{Method} We again use the modified Gaussian filtering implementation for dense CRFs by
+\citep{ajanthan2017efficient} to compute the conditional gradient at each step.
+To compute the contribution of the higher-order model, we make use of an
+efficient strategy to reuse computation. Recall that computing the conditional
+gradient coordinates $s^*_{\sigma(i)}$ requires us to obtain differences of the
+nested sets $F(S_i) - F(S_{i - 1})$. We store the clique state at each $i$, and
+this enables us to compute the difference for each coordinate in a constant
+amount of time. In this case, computation of conditional gradient takes
+$\mathcal{O}(cNL)$ where $c$ is the number of higher-order cliques. We ran our
+algorithm for 200 iterations.
+
+\myparagraph{Results} Figure \ref{fig:stereo_hod} shows some example solutions
+obtained by picking the label with maximum marginal probability for each
+variable. We observe that our method gives results that are reasonably close to
+the ground-truth. The set of hyperparameters that we used can be found in the
+appendix.
+
+\begin{figure*}
+    \centering
+\begin{tabular}{ccc}
+        \includegraphics[scale = 0.24]{\imagePath/stereo/venus_GT} &
+        \includegraphics[scale = 0.24]{\imagePath/stereo/potts/venus_left_mf} &
+        \includegraphics[scale = 0.24]{\imagePath/stereo/potts/venus_left_submod} \\
+        \scriptsize(a) Venus GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution \\ 
+\label{fig:stereo}
+\end{figure*}
+
+\mysubsection{Semantic Segmentation using Higher-order Diversity Model}
+\label{subsec:semantic_hod}
+\myparagraph{Data} We evaluate our approach on the task of semantic segmentation on the MSRC-21 dataset \citep{shotton2009textonboost}.
+It consists of 591 color images of size 320$\times$213 with corresponding ground truth labelings of 21 object classes.
+We made use of the unary features from the \textit{TextonBoost} classifier \citep{shotton2009textonboost}.
+As for stereo matching (subsection \ref{subsec:stereo_hod}), we augment the dense CRF with Gaussian pairwise potentials with our higher-order diveristy model.
+Superpixels for higher-order modeling were obtained from the mean-shift algorithm \citep{comaniciu2002mean}.
+
+\myparagraph{Method} As for stereo matching, we used the modified Gaussian filtering implementation \citep{ajanthan2017efficient} for the contribution of the dense pairwise terms to the conditional gradient.
+We used the same strategy as in subsection \ref{subsec:stereo_hod} to compute the higher-order component of the conditional gradient in $\mathcal{O}(cNL)$ time-complexity.
+We ran our
+algorithm for 100 iterations.
+
+\myparagraph{Results} The ground-truth labelings provided with the MSRC-21
+dataset are very coarse. We evaluate our algorithm on the set of 94 images for which fine-grain
+annotations are available \citep{koltun2011efficient}. On this subset, our
+method correctly labeled 81.18\% of pixels. Some representative results are
+shown in figure \ref{}. Our set of hyperparameters for this dataset can be found in the
+appendix.
 
@@
 \label{sec:conclude}
-We have established the relation between submodular extension for the Potts
-model and the LP relaxation for MAP estimation using Lovasz extension.
-This allowed us to identify the worst-case optimal submodular extension for Potts as well as the general metric labeling problems.
-It is worth noting that it might still be possible to obtain an improved submodular extension for a given problem instance.
-The design of a computationally feasible algorithm for this task is an interesting direction of future research.
-While our current work has focused on pairwise graphical models, it can be readily applied to high-order potentials by considering the corresponding LP relaxation objective as the Lovasz extension of a submodular extension.
-The identification of such extensions for popular high-order potentials such as the $P^n$ Potts model or its robust version could further improve the accuracy of important computer vision applications such as semantic segmentation.
-
+We have established the relation between submodular extension and the LP
+relaxation for MAP estimation using Lovasz extension for various CRF models. This
+allowed us to identify the worst-case optimal submodular extension for Potts as
+well as the general metric labeling problems. In addition, we obtained an
+accurate submodular extension for a higher-order model based on label-diversity
+in cliques. It is worth noting that it might still be possible to obtain an
+improved submodular extension for a given problem instance. The design of
+a computationally feasible algorithm for this task is an interesting direction
+of future research. While our work focused on one class of higher-order model,
+there is potential for our approach to be used to identify accurate submodular
+extensions for other energy families, such as truncated max-of-convex models \citep{pansari2017truncated}. 
 \newpage

### 2019-01-03 tex/abstract.tex
Incorporated Chris's suggestions.

@@
     filtering method; and (iv) propose an accurate submodular extension based on an LP
-    relaxation for a higher-order diversity model.
-Using synthetic and real data, we show that our approach provides comparable upper bounds on the log-partition function to those obtained using tree-reweighted message passing (TRW) in cases where the latter is computationally feasible.
-Importantly, unlike TRW, our approach provides the first practical algorithm to compute an upper bound on the dense CRF model with higher-order Potts potentials.
+    relaxation for a higher-order diversity model. Using synthetic and real
+    data, we show that our approach provides comparable upper bounds on the
+    log-partition function to those obtained using tree-reweighted message
+    passing (TRW) in cases where the latter is computationally feasible.
+    Importantly, unlike TRW, our approach provides the first computationally
+    tractable algorithm to compute an upper bound on dense CRF model with higher-order Potts potentials.
 \end{abstract}

### 2019-01-03 tex/paper.tex
Incorporated Chris's suggestions.

@@
 CRF by minimising an upper-bound on the log-partition function. In the initial
-work of \citep{djolonga2014map}, the energy of the CRF was restricted to be submodular.
-In a later work ~\citep{zhang2015higher}, the algorithm was extended to handle more general Potts energy functions.
-The key idea here is to define a large ground set such that its subsets represent valid labelings, sublabelings or even incorrect labelings (these may assign two separate labels to a random variable and hence be invalid).
+work of \citep{djolonga2014map}, the energy of the CRF was restricted to be
+submodular. In a later work ~\citep{zhang2015higher}, the algorithm was
+extended to handle more general Potts energy functions.
+The key idea there was to define a large ground set such that its subsets represent valid labelings, sublabelings or even incorrect labelings (these may assign two separate labels to a random variable and hence be invalid).
 Given the large ground set, it is possible to define a submodular set function whose value is equal to the energy of the CRF for subsets that specify a valid labeling of the model.
@@
 relaxation. Higher-order models can capture interesting properties of the image that cannot be
-done using pairwise models~\citep{vineet2014filter, kohli2007p3}. We show that our upper-bounds on synthetic problems are comparable to those from tree reweighted message passing (TRW)
+expressed using pairwise models~\citep{vineet2014filter, kohli2007p3}. We show that our upper-bounds on synthetic problems are comparable to those from tree reweighted message passing (TRW)
 ~\citep{wainwright2005new} for the case of sparse CRFs. Unlike our approach,
@@
 
-We now introduce the notation and definitions that we will make use of in the remainder of the paper.
+We now introduce the notation and definitions that we make use of in the remainder of the paper.
 
@@
 The function $f$ is an extension because it equals $F$ on the vertices of the
-unit cube. That is, for any $A \subseteq V$, $f({\bf 1}_A) = F(A)$ where ${\bf
+unit cube. That is, for any $A \subseteq V$, $f({\bf 1}_A) = F(A)$, where ${\bf
 1}_A$ is the $0\mhyphen{1}$ indicator vector corresponding to the elements of $A$.
@@
 
-We can augment the above pairwise models with higher-order potentials.
-Let $\mathcal C$ be the set of cliques, that is, fully-connected groups of variables.
-These cliques can, for instance, be the set of superpixels from an image segmentation method.
-Also, let ${\bf x}_c$ be the component of $\bf x$ formed by elements in clique $c$.
-The energy function now also contains higher-order potentials
+We can augment the above pairwise models with higher-order potentials. Let
+$\mathcal C$ be the set of cliques on subgroups of variables. These cliques
+can, for instance, be the set of superpixels from a clustering method, such as
+$k$-means or mean-shift \cite{comaniciu2002mean}. Also, let ${\bf x}_c$ be the component of $\bf x$ formed by elements in clique $c$. The energy function now also contains higher-order potentials 
 %
@@
 
-The energy function also defines a probability distribution $P({\bf x})$ as follows:
+The energy function can be interpreted as defining a probability distribution $P({\bf x})$ as follows:
 \begin{equation}
@@
 We now summarise the marginal inference method of Zhang et al. \citep{zhang2015higher} which
-uses submodular extensions.
+made use of a submodular extension.
 
@@
 a ground set such that some of its subsets correspond to valid CRF labelings.
-To obtain such an extension, we need an encoding scheme which gives the sets corresponding to valid CRF labelings.
+Note that not every subset needs to represent a valid labeling - some of them could
+correspond to incomplete or invalid labelings. To obtain such an extension, we need an encoding scheme which gives the sets corresponding to valid CRF labelings.
 
@@
 candidate for solving problem (\ref{eq:min_upper_bound}) due to two reasons.
-First, problem \eqref{eq:min_upper_bound} is convex.
-Second, as solving an LP over EP(F) is computationally tractable (property \ref{prop:greedy}), the conditional gradient can be found efficiently.
-The algorithm starts with an initial solution $\bf s_0$ (line \ref{line:initial}).
-At each iteration, we compute the conditional gradient $\bf s^*$ (line \ref{line:cond_grad}), which minimises the linear approximation $g({\bf s}_k) + \nabla g({\bf s}_k)^T ({\bf s} - {\bf s}_k)$ of the objective function.
-Finally, $\bf s$ is updated by either (i) fixed step size schedule, as in line \ref{line:step_size} of algorithm \ref{algo:fw_inference}, or (ii) by doing line search ${\bf s}_{k+1} = \min_{0 \leq \gamma \leq 1} g(\gamma {\bf s}^* + (1 - \gamma) {\bf s}_{k})$.
+First, problem \eqref{eq:min_upper_bound} is convex. Second, as solving an LP
+over EP(F) is computationally tractable (property \ref{prop:greedy}), the
+conditional gradient can be found efficiently. The algorithm starts with an
+initial solution $\bf s_0$ (line \ref{line:initial}). At each iteration, we
+compute the conditional gradient $\bf s^*$ (line \ref{line:cond_grad}), which
+minimises the linear approximation $g({\bf s}_k) + \nabla g({\bf s}_k)^T ({\bf
+s} - {\bf s}_k)$ of the objective function. Finally, $\bf s$ is updated by
+either (i) fixed step size schedule, as in line \ref{line:step_size} of
+algorithm \ref{algo:fw_inference}, or (ii) by doing a line search and setting ${\bf s}_{k+1} = \min_{0 \leq \gamma \leq 1} g(\gamma {\bf s}^* + (1 - \gamma) {\bf s}_{k})$.
 
@@
 How does one come up with accurate extensions for different classes of CRF energy functions?
-Is it possible to identify optimal extensions which yields the tightest bound?
-
-If we introduce a temperature parameter in $P({\bf x})$ (equation \eqref{eq:crf_prob_distrib}) by using $E({\bf x})/T$ and decrease $T$, the resulting distribution starts to peak more sharply around its mode.
-As $T \to 0$, marginal estimation becomes the same as MAP inference since the resulting distribution $P^0({\bf x})$ has mass 1 at its mode ${\bf x}^*$ and is 0 everywhere else.
-Given the MAP solution ${\bf x}^*$, one can compute the marginals as $P(X_i = j) = [x_i^* = j]$, where [.] is the Iverson bracket.
-
+Is it possible to identify optimal extensions which yield the tightest bound?
+
+If we introduce a temperature parameter in $P({\bf x})$ (equation
+\eqref{eq:crf_prob_distrib}) by using $E({\bf x})/T$ and decrease $T$, the
+resulting distribution starts to peak more sharply around its mode. This
+is assuming that $P({\bf x})$ is unimodal, and there aren't multiple solutions
+with the same minima. As $T \to 0$, marginal estimation becomes the same as MAP inference since the resulting
+distribution $P^0({\bf x})$ has mass 1 at its mode ${\bf x}^*$ and is
+0 everywhere else. Given the MAP solution ${\bf x}^*$, one can compute the
+marginals as $P(X_i = j) = [x_i^* = j]$, where [.] is the Iverson bracket.
 Motivated by this connection, we ask if one can introduce a temperature parameter to our problem \eqref{eq:min_upper_bound} and transform it to an LP relaxation in the limit $T \to 0$?
@@
 
-In this paper, we focus on submodular extension families $\mathcal F(.)$ corresponding to different classes of energy functions $\mathcal E$.
-Given an instance of the energy function $E(.)$ from the class $\mathcal E$,  there is a corresponding submodular extension ${\mathcal F}(E)$.
-Also, we seek a closed-form analytical expression for $\mathcal F$.
+In this work, we focus on submodular extension families $\mathcal F(.)$
+corresponding to different classes of energy functions $\mathcal E$. Given an
+instance of the energy function $E(.)$ from the class $\mathcal E$,  there is
+a corresponding submodular extension ${\mathcal F}(E)$. Moreover, we seek a closed-form analytical expression for $\mathcal F$.
 
@@
 
-Our extensions are derived from LP relaxations. When the LP relaxation is the tightest known, the extension family ${\mathcal
-F}_{opt}$ we obtain is \emph{worst-case optimal}. This implies that there does not
+Our extensions are derived from LP relaxations. When the LP relaxation is the
+tightest possible, the extension family ${\mathcal
+F}_{opt}$ we obtain is \emph{worst-case optimal}. This is to say that there does not
 exist another submodular extension family ${\mathcal F}$ that gives a tighter
@@
 \end{equation}
-Note that the notion of worst-case optimality does not guarantee that the extension is best for every instance of the energy function.
-It may be possible to solve an optimisation problem to obtain the best extension for every given instace of energy $E(.)$.
-However, in this paper, we do not take that approach, and instead provide a general solution for a given class of energy function.
+Note that the notion of worst-case optimality does not guarantee that the
+extension is best for every instance of the energy function. It may be possible
+to solve an optimisation problem to obtain the best extension for every given
+instance of energy $E(.)$. However, in this paper, we do not take that approach, and instead provide a general solution for a given class of energy function.
 
@@
 %
-\myparagraph{Tightest LP Relaxation} Before describing our set encoding and submodular extension, we briefly outline the LP relaxation of the corresponding MAP estimation problem.
-To this end, we define indicator variables $y_{ai}$ which equal 1 if $X_a = i$, and 0 otherwise.
-The following LP relaxation is the tightest known for Potts model in the worst-case, assuming the Unique Games Conjecture to be true \citep{manokaran2008sdp}:
+\myparagraph{Tightest LP Relaxation} Before describing our set encoding and
+submodular extension, we briefly outline the LP relaxation of the corresponding
+MAP estimation problem. To this end, we define indicator variables $y_{ai}$
+which equal 1 if $X_a = i$, and 0 otherwise. The following LP relaxation is the
+tightest possible for Potts model in the worst-case, assuming the Unique Games
+Conjecture is true \citep{manokaran2008sdp}:
 \begin{align}
@@
 \end{equation}
-(See Remark 1 in appendix)
+(See Remark 1 in appendix A for proof)
 \myparagraph{Marginal Estimation with Temperature} We now introduce
@@
 
-The above problem is equivalent to an LP relaxation of the corresponding MAP estimtation problem (see Remark 2 in appendix).
-We note that $g_T ({\bf s})$ in problem \eqref{eq:potts_temp} becomes the objective function of an LP relaxation in the limit $T \to 0$.
-We seek to obtain the worst-case optimal submodular extension by making $g_T ({\bf s})$ same as the objective of (P-LP) as $T \to 0$.
-Since at $T = 1$, problems \eqref{eq:potts_temp} and \eqref{eq:potts_upper_bound} are equivalent, this gives us the worst-case optimal extension for our problem \eqref{eq:potts_upper_bound}.
+The above problem is equivalent to an LP relaxation of the corresponding MAP
+estimation problem (see Remark 2 in appendix). We note that $g_T ({\bf s})$ in
+problem \eqref{eq:potts_temp} becomes the objective function of an LP
+relaxation in the limit $T \to 0$. We seek to obtain the worst-case optimal
+submodular extension by making $g_T ({\bf s})$ same as the objective of (P-LP)
+as $T \to 0$. We note that problems \eqref{eq:potts_temp} and
+\eqref{eq:potts_upper_bound} become equivalent at $T = 1$, and our worst-case optimality
+guarantee was derived in the limit $T \to 0$. However, it is possible to
+derive such a guarantee only in the limit case, and it is in this sense that we
+claim the worst-case optimality.
 
@@
 
-Potts model imposes the same penalty for unequal assignment of labels to neighbouring variables, regardless of the label dissimilarity.
-A more natural approach is to vary the penalty based on how different the labels are.
+Potts model imposes the same penalty for unequal assignment of labels to
+neighbouring variables, regardless of the label dissimilarity. In some
+scenarios, a more natural approach is to vary the penalty based on how different the labels are.
 A hierarchical Potts model permits this by specifying the distance between labels using a tree with the following properties:
@@
 %
-\myparagraph{Tightest LP Relaxation} We use the same indicator variables $y_{ai}$ that were employed in the LP relaxation of Potts model.
-Let $y_a(T) = \sum_{i \in L(T)} y_{ai}$.
-The following LP relaxation is the tightest known for hierarchical Potts model in the worst-case, assuming the Unique Games Conjecture to be true \citep{manokaran2008sdp}
+\myparagraph{Tightest LP Relaxation} We use the same indicator variables
+$y_{ai}$ that were employed in the LP relaxation of Potts model. Let $y_a(T)
+= \sum_{i \in L(T)} y_{ai}$. The following LP relaxation is the tightest known
+for hierarchical Potts model in the worst-case, assuming the Unique Games
+Conjecture is true \citep{manokaran2008sdp}
 \begin{align}
@@
 The above problem is equivalent to an LP relaxation of the corresponding MAP
-estimtation problem (see Remark 3 in appendix).
+estimation problem (see Remark 3 in appendix).
 Hence, $g_T ({\bf s})$ becomes the objective function of an LP relaxation in the limit $T \to 0$.
@@
 Each $F$ evaluation has complexity ${\mathcal O}(N)$ using the efficient Gaussian filtering algorithm of \citep{koltun2011efficient}.
-However, computation of $s^*$ would still be ${\mathcal O} (N^2)$ this way, which is clearly impractical for computer-vision applications where $N \sim 10^5 - 10^6$.
+However, computation of $s^*$ would still be ${\mathcal O} (N^2)$, which is clearly impractical for computer-vision applications where $N \sim 10^5 - 10^6$.
 
@@
 
-A higher-order model useful in real-world applications is the \textit{diversity} model, which favours labeling where variables in a clique have fewer number of unique labels.
-For instance, in semantic segmentataion we often first obtain superpixels from an image segmentation method, which we consider as cliques in our model.
+A higher-order model useful in real-world applications is the
+\textit{diversity} model, which favours labeling where variables in a clique
+have fewer number of unique labels. For instance, in semantic segmentation we
+often first obtain superpixels from a clustering method \cite{comaniciu2002mean}, which we consider as cliques in our model.
 We expect the labeling in a superpixel to be homogeneous.
@@
 We now relax the indicator variables $y_{ai}$ to lie in $[0, 1]$.
-Assuming the pairwise potentials in equation (\ref{eq:energy_hop_min}) to be Potts model, an accurate LP relaxation for the higher-order diversity model is given by:
+Assuming the pairwise potentials in equation (\ref{eq:energy_hop_min}) to be Potts, an accurate LP relaxation for the higher-order diversity model is given by:
 
@@
 \end{align}
-where the label (HOD-LP) stands for LP relaxation for higher-order diversity model.
-The objective function is similar to that for the Potts model along with the addition of terms corresponding to higher-order potentials.
-The constraints for (HOD-LP) are the same as for (P-LP).
-The above LP has not been formally analysed in literature.
-Hence, we do not make any claims regarding its optimality.
-However, its similairty in form to (P-LP) is an indication of its accuracy.
-We establish its accuracy empirically in section \ref{sec:exp}.
+where the label (HOD-LP) stands for LP relaxation for higher-order diversity
+model. The objective function is similar to that for the Potts model along with
+the addition of terms corresponding to higher-order potentials. The constraints
+for (HOD-LP) are the same as for (P-LP). The above LP has not been formally
+analysed in literature, and we do not make any claims regarding its
+optimality. However, its similarity in form to (P-LP) is an indication of its accuracy. We establish its accuracy empirically in section \ref{sec:exp}.
 
@@
 scheme as for Potts model. Problem \ref{eq:min_upper_bound} then factorises as: 
-
 %Same as eq 5.4
@@
 \end{equation}
-
 We introduce a temperature parameter $T$ to problem (\ref{eq:hop_upper_bound}) to obtain the following new problem:
-
 %Same as eq 5.5
@@
 \end{equation}
-
-Considering the limit $T \to 0$, problem (\ref{eq:hop_temp}) becomes
-
+In the limit $T \to 0$, problem (\ref{eq:hop_temp}) becomes
 %Same as eq 5.6
@@
 where $f(.)$ is the Lovasz extension of $F(.)$. We are interested in making
-this LP same as (HOD-LP) of equation (\ref{eq:hop-lp}), thereby enabling us to obtain the submodular extension for the higher-order diversity model.
+this LP the same as (HOD-LP) of equation (\ref{eq:hop-lp}), thereby enabling us to obtain the submodular extension for the higher-order diversity model.
 We will make use of the following lemma in our proof.
-
 {\lemma The Lovasz extension of the following set function 
@@
 
-Also, we showed in lemma \ref{lemma:hod_extension} that the Lovasz extension of $F_{HOD}$ is $E({\bf y})$ in probem (HOD-LP).
+Also, we showed in lemma \ref{lemma:hod_extension} that the Lovasz extension of
+$F_{HOD}$ is $E({\bf y})$ in problem (HOD-LP).
 $E({\bf y})$ being a sum of terms corresponding to maximum of linear functions, is convex.
@@
 
-\myparagraph{Method} For our algorithm, we use the standard schedule $\gamma = 2/(k + 2)$ to obtain step size $\gamma$ at iteration $k$.
-We run our algorithm till convergence - 100 iterations suffice for this.
+\myparagraph{Method} For our algorithm, we use the standard schedule $\gamma
+= 2/(k + 2)$ to obtain step size $\gamma$ at iteration $k$. We run our
+algorithm till convergence - 100 iterations suffices for this.
 The experiments are repeated for 100 randomly generated unaries for each model and each weight.
@@
 compared to mean-field is 0.943.
-The avearge time ratio is 10.66.
+The average time ratio is 10.66.
 We observe that our algorithm results in more natural looking stereo matching results with lower energy values for all images.
@@
         \includegraphics[scale = 0.24]{\imagePath/stereo/potts/venus_left_submod} \\
-        \scriptsize(a) Venus GT & \scriptsize(b) MF solution & \scriptsize(c) Submod solution \\ 
-\label{fig:stereo}
+\end{tabular}
+\mycaption{\footnotesize \em To fill}
+%\label{fig:stereo}
 \end{figure*}
@@
 \label{subsec:semantic_hod}
-\myparagraph{Data} We evaluate our approach on the task of semantic segmentation on the MSRC-21 dataset \citep{shotton2009textonboost}.
-It consists of 591 color images of size 320$\times$213 with corresponding ground truth labelings of 21 object classes.
-We made use of the unary features from the \textit{TextonBoost} classifier \citep{shotton2009textonboost}.
-As for stereo matching (subsection \ref{subsec:stereo_hod}), we augment the dense CRF with Gaussian pairwise potentials with our higher-order diveristy model.
-Superpixels for higher-order modeling were obtained from the mean-shift algorithm \citep{comaniciu2002mean}.
+\myparagraph{Data} We evaluate our approach on the task of semantic
+segmentation on the MSRC-21 dataset \citep{shotton2009textonboost}. It consists
+of 591 color images of size 320$\times$213 with corresponding ground truth
+labelings of 21 object classes. We made use of the unary features from the
+\textit{TextonBoost} classifier \citep{shotton2009textonboost}. As for stereo
+matching (subsection \ref{subsec:stereo_hod}), we augment the dense CRF with
+Gaussian pairwise potentials with our higher-order diversity model. Superpixels for higher-order modeling were obtained from the mean-shift algorithm \citep{comaniciu2002mean}.
 

### 2019-01-03 tex/appendix.tex
Semantic segmentation experiments added. List of hyperparameters added.

@@
 \end{proof}
+
+\mysection{List of Hyperparameters}
+
+For pairwise dense CRFs, following \cite{koltun2011efficient}, we use constrast-sensitive two-kernel potentials for both stereo matching and semantic segmentation.
+These are defined in terms of the color vectors ${I}_a$ and ${I}_b$ and positions $p_a$ and $p_b$ as:
+
+\begin{equation}
+k({\bf f}_a, {\bf f}_b) = w^{(1)}\exp\left(- \frac{|p_a - p_b|^2}{2 \theta_{\alpha}^2} - \frac{|{I}_a - {I}_b|^2}{2 \theta_{\beta}^2}\right) + w^{(2)}\exp\left(- \frac{|p_a - p_b|^2}{2 \theta_{\gamma}^2}\right)
+\end{equation}
+
+For the higher-order diversity model, we have a clique weight $w_c$ that is the same across all cliques. We give the set of hyperparameters for our experiments in table \ref{tab:hyper}.
+
+\begin{table}[]
+\begin{tabular}{|l|l|l|l|l|l|l|}
+\hline
+{} &  $w^{(1)}$ & $\theta_{\alpha}$ & $\theta_{\beta}$ &  $w^{(1)}$ & $\theta_{\gamma}$ & $w_c$\\
+\hline
+Stereo matching (dense pairwise, Submod) & 1.64 & 12.99 & 50.84 &  1.51 & 1.10 & 0  \\
+\hline
+Stereo matching (dense pairwise, MF) & 12.43 & 22.38 & 17.26 & 4.59 & 8.75 & 0  \\
+\hline
+Stereo matching (higher-order, Submod) & 12.43 & 22.38 & 17.26 & 4.59 & 8.75 & 10.00 \\
+\hline
+Semantic Segmentation (higher-order, Submod) & 1.00 & 1.00 & 50.84 & 1.51 & 1.10 & 10.00 \\
+\hline
+\end{tabular}
+\caption{List of hyperparameters used in experiments}
+\label{tab:hyper}
+\end{table}
+
 \iffalse

### 2019-01-03 tex/paper.tex
Semantic segmentation experiments added. List of hyperparameters added.

@@
     \centering
-\begin{tabular}{ccc}
-        \includegraphics[scale = 0.24]{\imagePath/stereo/venus_GT} &
-        \includegraphics[scale = 0.24]{\imagePath/stereo/potts/venus_left_mf} &
-        \includegraphics[scale = 0.24]{\imagePath/stereo/potts/venus_left_submod} \\
+\begin{tabular}{cc}
+  \includegraphics[scale = 1.3]{./figures/semantic/2_14_s} &
+  \includegraphics[scale = 0.5]{./figures/semantic/2_14_s_submod} \\
+        \scriptsize (a) & \scriptsize (e) \\
+  \includegraphics[scale = 1.3]{./figures/semantic/11_4_s} &
+  \includegraphics[scale = 0.5]{./figures/semantic/11_4_s_submod} \\
+        \scriptsize (b) & \scriptsize (f) \\
+  \includegraphics[scale = 1.3]{./figures/semantic/12_7_s} &
+  \includegraphics[scale = 0.5]{./figures/semantic/12_7_s_submod} \\
+        \scriptsize (c) & \scriptsize (g) \\
+  \includegraphics[scale = 1.3]{./figures/semantic/4_16_s} &
+  \includegraphics[scale = 0.5]{./figures/semantic/4_16_s_submod} \\
+        \scriptsize (d) & \scriptsize (h) \\
 \end{tabular}
-\mycaption{\footnotesize \em To fill}
-%\label{fig:stereo}
+\mycaption{\footnotesize \em Sample results from semantic segmentation on the
+MSRC-21 dataset using dense CRFs with higher-order diversity term in addition
+to Potts compatibility and Gaussian pairwise potentials. (a)-(d) show the input
+image, and (e)-(f) show the corresponding results using our method. We observe that our
+method is able to segment out objects with complex boundaries fairly
+accurately.}
+\label{fig:semantic}
 \end{figure*}
@@
 annotations are available \citep{koltun2011efficient}. On this subset, our
-method correctly labeled 81.18\% of pixels. Some representative results are
-shown in figure \ref{}. Our set of hyperparameters for this dataset can be found in the
+method correctly labeled 81.18\% of pixels. It took 275.02s on average per
+instance to run our method for 100 iterations. Some representative results are
+shown in figure \ref{fig:semantic} where we label each pixel with the label
+having the maximum marginal probability. Our set of hyperparameters for this dataset can be found in the
 appendix.

### 2019-01-03 tex/paper.tex
Enlarged figures. Improved placement.

@@
 \centering
-\includegraphics[scale = 0.30]{./figures/1_of_L.png}
+\includegraphics[scale = 0.35]{./figures/1_of_L.png}
 \mycaption{\footnotesize \em Illustration of $1$-of-$L$ encoding used in \citep{zhang2015higher} with 4 variables and 3 labels.
@@
 \centering
-\includegraphics[scale = 0.30]{./figures/potts_chain_extension.png}
+\includegraphics[scale = 0.35]{./figures/potts_chain_extension.png}
 \mycaption{\footnotesize \em An illustration of the worst-case optimal
@@
 \centering
-\includegraphics[scale = 0.30]{./figures/rhst_notation.png}
+\includegraphics[scale = 0.35]{./figures/rhst_notation.png}
 \mycaption{\footnotesize \em A hierarchical Potts model instance illustrating notations with 2 meta-labels (blue) and 4 labels (yellow).
@@
 \begin{tabular}{cc}
-\includegraphics[scale = 0.25]{./figures/synthetic_tree.png} &
-\includegraphics[scale = 0.25]{./figures/synthetic_tree_2_new.png} \\
+\includegraphics[scale = 0.30]{./figures/synthetic_tree.png} &
+\includegraphics[scale = 0.30]{./figures/synthetic_tree_2_new.png} \\
   \scriptsize(a) Tree A & \scriptsize(b) Tree B \\
 \end{tabular}
-\mycaption{\footnotesize \em The hierarchical Potts models defining pairwise
+\mycaption{\footnotesize \em {\bf Trees for synthetic experiments:} The hierarchical Potts models defining pairwise
 distance among 20 labels used for upper-bound comparison with TRW. Blue nodes
@@
 \centering
-\resizebox{180pt}{150pt}{
+\resizebox{190pt}{160pt}{
 \begin{tikzpicture}[-latex ,auto ,node distance =2 cm and 2cm ,on grid
@@
 }
-\mycaption{\footnotesize \em An st-graph specifying the alternate submodular
+\mycaption{\footnotesize \em {\bf Alternate extension for synthetic experiments}: An st-graph specifying the alternate submodular
 extension for Potts model for 2 variables with 3 labels each and connected to
@@
 \centering
-\includegraphics[scale = 0.50]{./figures/synthetic_upper_bound_comparison.pdf}
-\mycaption{\footnotesize \em Upper-bound comparison using synthetic data. The
+\includegraphics[scale = 0.80]{./figures/synthetic_upper_bound_comparison.pdf}
+\mycaption{\footnotesize \em {\bf Upper-bound comparison using synthetic data:} The
 plot shows the ratio (Submod bound - TRW bound)/|TRW bound| averaged over 100
@@
 
+\clearpage
 \mysubsection{Stereo Matching using Dense CRFs}
@@
         \scriptsize(a) Tsukuba GT & \scriptsize(b) Submod solution \\
-        {} & {}  \\
+        {} & \scriptsize 1582.04s  \\
         \includegraphics[scale = 0.52]{\imagePath/stereo/higher_diversity/baby1_GT} & 

### 2019-01-03 tex/appendix.tex
Corrected dangling words.

@@
 
-{\lemma Given any submodular extension $F(.)$ of an hierachical Potts energy function $E(.)$, its Lovasz extension $f(.)$ defines an LP relaxation of the corresponding MAP estimation problem as
+{\lemma Given any submodular extension $F(.)$ of an hierarchical Potts energy function $E(.)$, its Lovasz extension $f(.)$ defines an LP relaxation of the corresponding MAP estimation problem as
 \begin{align}

### 2019-01-03 tex/paper.tex
Corrected dangling words.

@@
 
-The energy function can be interpreted as defining a probability distribution $P({\bf x})$ as follows:
+The energy function can be interpreted as defining a probability distribution $P({\bf x})$ as:
 \begin{equation}
@@
 Different choices of extensions $F$ change the domain in problem \eqref{eq:min_upper_bound}, leading to different upper bounds on the log-partition function.
-How does one come up with accurate extensions for different classes of CRF energy functions?
+How does one come up with accurate extensions for different classes of CRF energy?
 Is it possible to identify optimal extensions which yield the tightest bound?
@@
 upper bound for problem \eqref{eq:min_upper_bound} than ${\mathcal F}_{opt}$
-for all instances of the energy function in $\mathcal E$. Formally, 
+for all instances of the energy function in $\mathcal E$. Formally, ${\mathcal F}_{opt}$ is worst-case optimal if
 \begin{equation}
@@
 In subsection \ref{subsec:hier_potts}, we obtain worst-case optimal submodular extension for the more general hierarchical Potts model.
-Finally, in subsection \ref{subsec:dense_crf}, we provide an approach to make our inference algorithm efficient for the dense CRF model.
+Finally, in subsection \ref{subsec:dense_crf}, we provide a way to make our inference algorithm efficient for the dense CRF model.
 Using our approach, the inference algorithm has linear time-complexity in the number of variables and labels.
@@
 submodular extension for Potts model for a chain graph of 4 variables, each of
-which can take 3 labels. The figure shows the way to compute the extension
-values of the set $A = \{v_{1, 1}, v_{4, 1}, v_{3, 2}$\}.}
+which can take 3 labels. The figure shows how to compute extension for $A = \{v_{1, 1}, v_{4, 1}, v_{3, 2}$\}.}
 \label{fig:potts_extension_example}

### 2019-01-05 siam_cover_letter.tex
Cover letter for siam submission.

% SIAM Article Template
\documentclass[review,onefignum,onetabnum]{siamonline171218}

% Information that is shared between the article and the supplement
% (title and author information, macros, packages, etc.) goes into
% ex_shared.tex. If there is no supplement, this file can be included
% directly.

\title{LP-based Submodular Extensions for Marginal Estimation\thanks{Submitted to the editors 5th January, 2019.}}

\author{Pankaj Pansari\thanks{University of Oxford
  (\email{pankaj@robots.ox.ac.uk}, \email{pawan@robots.ox.ac.uk}).}
\and Chris Russell\thanks{University of Surrey
  (\email{crussell@turing.ac.uk}).}
\and M. Pawan Kumar\footnotemark[2]}

\begin{document}
\maketitle

Dear editors,

We are submitting this paper as an extension of a paper published in the
proceedings of AISTATS 2018 titled ``Worst-case Optimal Submodular Extensions
for Marginal Estimation". Both papers focus on deriving submodular
extensions which affect the accuracy of a particular variational inference
method for a conditional random field (CRF). The same connection between this
variational method and linear programming (LP) relaxations of the corresponding
MAP inference problem is used in both versions. In the conference paper, we
used this relation to prove the worst-case optimality of a submodular
extension used in literature for Potts model by connecting it to the tightest
Potts relaxation. We also extended the method to hierarchical Potts model by
deriving its worst-case optimal extension. Finally, we provided an efficient
algorithm that made this variational inference method tractable for dense CRFs.

In this journal version, we extend the use of the variational method to
higher-order CRFs. We derive an accurate submodular extension for
a higher-order CRF with diversity-based clique potentials. This potential favours
labelings that use few number of diverse labels in a clique. For this
derivation, we make use of an LP relaxation of this higher-order model. The
tightness property of this relaxation has not been studied in literature, so we
cannot claim worst-case optimality of the submodular extension for this model.

To show the efficacy of our method for the higher-order diversity model, we
carried additional experiments for (a) stereo matching, and (b) semantic
segmentation on the MSRC-21 dataset. For both applications, we used a CRF with
dense pairwise potential plus the diversity-based higher-order potentials.
\end{document}
